### 2020-09-30 paper/paper.Rmd
Add skeleton of paper.

---
documentclass: jss
author:
  - name: Lauren Kennedy
    affiliation: Monash University
    email: \email{name@company.com}
    url: http://rstudio.com
  - name: Mitzi Morris
    affiliation: Columbia University
  - name: Jonah Gabry
    affiliation: Columbia University
  - name: Rohan Alexander
    affiliation: University of Toronto
title:
  formatted: "\\pkg{MRP-Kit} A 'grammar' of multilevel regression with post-stratification and implementation in \\proglang{R}"
  # If you use tex in the formatted title, also supply version without
  plain:     "A 'grammar' of multilevel regression with post-stratification and implementation in R"
  # For running headers, if needed
  short:     "\\pkg{foo}: A grammar of MRP"
abstract: >
  In this paper we define a grammar for multilevel regression with post-stratification (MRP) and implement this grammar in the R Package \pkg{MRPKit}. Our grammar is centered around the following verbs: add, remove, delete, and replace. These act on survey objects. Our grammar, and its implementation, provides a detailed workflow when conducting MRP that will be useful for researchers and in teaching.
keywords:
  # at least one keyword must be supplied
  formatted: [multilevel regression with post-stratification, "\\proglang{R}", reproducibility, statistics, political science]
  plain:     [multilevel regression with post-stratification, R, reproducibility, statistics, political science]
preamble: >
  \usepackage{amsmath}
output: rticles::jss_article
---

```{r, setup, include=FALSE}
options(prompt = 'R> ', continue = '+ ')
```

\section[Introduction]{Introduction} \label{sec:intro}

Multilevel regression with post-stratification (MRP) is a statistical approach in which surveys are related to each other using a statistical model. This is important because known biases in one survey, can be adjusted for by another in a statistically reasonable way. This enables better use of non-representative surveys, additional information, and propagation of uncertainty. However it can be difficult to use MRP due to this need to related two different datasets. This package defines a grammar, or list of underlying rules, of MRP and then describes an R package, \proglang{MRPKit}, that implements this grammar.

At its core, MRP is a mapping between a survey object and a population object. It is from this mapping that the power of MRP exists, but it also establishing this mapping that is the difficult part of implementing MRP models. Making this implementation easier and more reproducible is important as interest in, and the use of, MRP increases. It can be difficult even for those experienced with MRP to ensure there are no mistakes in this mapping and easing this is an important contribution to enhancing the reproducibility of MRP analysis.

We first define a grammar of MRP, which we define as the underlying rules and principles that are common to every analysis based on MRP. This grammar is based around the following verbs: 'add', 'remove', 'delete', and 'replace'. These verbs are applied to a survey object, a post-stratification object, and survey\_map object. These three objects come together to create a mapping object, which is what processes such as regression act on. Finally, common diagnostics and graphs are implemented. In this way, our grammar and package implement an entire statistical workflow for conducting MRP.

The survey object would typically be a regular survey, such as a political poll of 1,000 respondents, but it could also be a larger survey, such as the Canadian Election Survey, or similar. The post-stratification object would typically be a larger survey, such as, in the case of the US, the (INSERT THE USUAL ONE), or a census.

Our grammar and package complements existing packages such as \proglang{survey} (ADD CITATION), X and Y. These packages... Instead, ours is focused on what is needed for MRP.

The remainder of this paper is structured as follows: Section \ref{sec:review} reviews similar packages and contributions and places ours within that context. Section \ref{sec:components} discusses the grammar and the core aspects of MRP as implemented in \proglang{MRPKit}. Section \ref{sec:implementation} discusses some of the implementation issues and technical notes related to the decisions that were made. Section \ref{sec:vignette} provides two examples of the package in use, one using SOMETHING, and the other using SOMETHING ELSE. Finally, Section \ref{sec:summary} provides a summary discussion, some cautions and weaknesses, as well as notes about next steps.

\section{Review of the other packages} \label{sec:review}

The most common alternative at the moment to this package is for users to do all aspects themselves. While there is nothing conceptually difficult about MRP, the implementation can be difficult. In particular, preparing and matching different levels between surveys can be time consuming and potentially introduce errors.

HERE

Survey package

Declare design

DIY

MRP package - https://github.com/gelman/mrp

And how we're different

\section{Components and grammar} \label{sec:components}

The \proglang{MRPKit} package has the following key components: survey objects, which for most users will be a collection of two surveys where one is larger than the other; a survey map, that relates the survey objects and then an MRP object, which is created once the survey map is applied to the survey objects. Regression acts on the MRP object. These objects are subject to the following verbs: 'add', 'delete', 'new', and 'replace'. For instance: \proglang{SurveyMap\$add}, \proglang{SurveyOb\$delete}, \proglang{SurveyMap\$new}, and \proglang{SurveyMap\$replace}.

\subsection*{Survey objects}

Survey objects are the surveys that MRP will bring together. A user can \proglang{add} a survey object by providing a CSV file location. The user then needs to identify the data types of each column. Typically, there will need to be two survey objects, where one will be the survey that is of interest for its response variables, such as political opinion, and another will be a survey that is to be used for post-stratification.

New is when you have a new survey object and you are bringing it in. e.g.

% sampl <- surveyStr$new(...)
% pop <- surveyStr$new(...)
% map <- surveyMap$new(sampl, pop, ...)
% mrp1 <- MRP$new(sampl, pop, map)

% option 1
% mod1 <- mrp1$fit(y ~ x + (1|group))
% mod1 <- mrp1$fit(y ~ x + (1|group) + (1|group2))

% mod1$poststratify()
% mod2$poststratify("model1")

% or

% option 2
% mrp1$fit(y ~ x + (1|group), name = "model1")
% mrp1$fit(y ~ x + (1|group) + (1|group2), name = "model2")

% mrp1$poststartify("model1")
% mrp1$poststartify("model2")

What is the difference between add and new? What is replace doing? URGH, I just need to look at the repo.

\subsection*{Survey map}

A `SurveyMap` object holds the mapping between a set of items in a survey and a population dataset. The label is the item label in each dataset and the values is a list of all possible values.   The values for the survey and population must be aligned, i.e., the lists must have the same number of elements and the values at index i in each list are equivalent.  If there is a meaningful ordering over the values, they should be listed in that order, either descending or ascending.

One of the fundamental issues when conducting MRP is to ensure that

\subsection*{MRP object}

An  MRP object contains survey objects, and a survey map. At the point at which the survey objects are put into the MRP object they become immutable but the survey map object is.

The MRP object outputs an analysis. There would be a MRP model fit class that would

Many objects on the same class. There are two

\section{Implementation and technical notes} \label{sec:implementation}

\section{Vignette} \label{sec:vignette}

\subsection*{Using CCES data}

\subsection*{Using your own data}

\section{Summary and discussion} \label{sec:summary}

\subsection*{Next steps and cautions}

\section*{Acknowledgments}

### 2020-10-04 paper/paper.tex

--- 2020-09-30 paper/paper.Rmd
@@
----
-documentclass: jss
-author:
-  - name: Lauren Kennedy
-    affiliation: Monash University
-    email: \email{name@company.com}
-    url: http://rstudio.com
-  - name: Mitzi Morris
-    affiliation: Columbia University
-  - name: Jonah Gabry
-    affiliation: Columbia University
-  - name: Rohan Alexander
-    affiliation: University of Toronto
-title:
-  formatted: "\\pkg{MRP-Kit} A 'grammar' of multilevel regression with post-stratification and implementation in \\proglang{R}"
-  # If you use tex in the formatted title, also supply version without
-  plain:     "A 'grammar' of multilevel regression with post-stratification and implementation in R"
-  # For running headers, if needed
-  short:     "\\pkg{foo}: A grammar of MRP"
-abstract: >
-  In this paper we define a grammar for multilevel regression with post-stratification (MRP) and implement this grammar in the R Package \pkg{MRPKit}.
-Our grammar is centered around the following verbs: add, remove, delete, and replace.
-These act on survey objects.
-Our grammar, and its implementation, provides a detailed workflow when conducting MRP that will be useful for researchers and in teaching.
-keywords:
-  # at least one keyword must be supplied
-  formatted: [multilevel regression with post-stratification, "\\proglang{R}", reproducibility, statistics, political science]
-  plain:     [multilevel regression with post-stratification, R, reproducibility, statistics, political science]
-preamble: >
+\documentclass[
+]{jss}
+
+\usepackage[utf8]{inputenc}
+
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+\author{
+Lauren Kennedy\\Monash University \And Mitzi Morris\\Columbia University \And Jonah Gabry\\Columbia University \And Rohan Alexander\\University of Toronto
+}
+\title{\pkg{MRP-Kit} A `grammar' of multilevel regression with
+post-stratification and implementation in \proglang{R}}
+
+\Plainauthor{Lauren Kennedy, Mitzi Morris, Jonah Gabry, Rohan Alexander}
+\Plaintitle{A `grammar' of multilevel regression with post-stratification and
+implementation in R}
+\Shorttitle{\pkg{foo}: A grammar of MRP}
+
+\Abstract{
+In this paper we define a grammar for multilevel regression with
+post-stratification (MRP) and implement this grammar in the R Package
+\pkg{MRPKit}. Our grammar is centered around the following verbs: add,
+remove, delete, and replace. These act on survey objects. Our grammar,
+and its implementation, provides a detailed workflow when conducting MRP
+that will be useful for researchers and in teaching.
+}
+
+\Keywords{multilevel regression with post-stratification, \proglang{R}, reproducibility, statistics, political science}
+\Plainkeywords{multilevel regression with post-stratification, R, reproducibility, statistics, political science}
+
+%% publication information
+%% \Volume{50}
+%% \Issue{9}
+%% \Month{June}
+%% \Year{2012}
+%% \Submitdate{}
+%% \Acceptdate{2012-06-04}
+
+\Address{
+        }
+
+% Pandoc header
+
 \usepackage{amsmath}
-output: rticles::jss_article
----
 
-```{r, setup, include=FALSE}
-options(prompt = 'R> ', continue = '+ ')
-```
+\begin{document}
 
@@
 
-Multilevel regression with post-stratification (MRP) is a statistical approach in which surveys are related to each other using a statistical model.
-This is important because known biases in one survey, can be adjusted for by another in a statistically reasonable way.
-This enables better use of non-representative surveys, additional information, and propagation of uncertainty.
-However it can be difficult to use MRP due to this need to related two different datasets.
-This package defines a grammar, or list of underlying rules, of MRP and then describes an R package, \proglang{MRPKit}, that implements this grammar.
+Multilevel regression with post-stratification (MRP) is a statistical
+approach in which surveys are related to each other using a statistical
+model. This is important because known biases in one survey, can be
+adjusted for by another in a statistically reasonable way. This enables
+better use of non-representative surveys, additional information, and
+propagation of uncertainty. However it can be difficult to use MRP due
+to this need to related two different datasets. This package defines a
+grammar, or list of underlying rules, of MRP and then describes an R
+package, \proglang{MRPKit}, that implements this grammar.
 
-At its core, MRP is a mapping between a survey object and a population object.
-It is from this mapping that the power of MRP exists, but it also establishing this mapping that is the difficult part of implementing MRP models.
-Making this implementation easier and more reproducible is important as interest in, and the use of, MRP increases.
-It can be difficult even for those experienced with MRP to ensure there are no mistakes in this mapping and easing this is an important contribution to enhancing the reproducibility of MRP analysis.
+At its core, MRP is a mapping between a survey object and a population
+object. It is from this mapping that the power of MRP exists, but it
+also establishing this mapping that is the difficult part of
+implementing MRP models. Making this implementation easier and more
+reproducible is important as interest in, and the use of, MRP increases.
+It can be difficult even for those experienced with MRP to ensure there
+are no mistakes in this mapping and easing this is an important
+contribution to enhancing the reproducibility of MRP analysis.
 
-We first define a grammar of MRP, which we define as the underlying rules and principles that are common to every analysis based on MRP.
-This grammar is based around the following verbs: 'add', 'remove', 'delete', and 'replace'.
-These verbs are applied to a survey object, a post-stratification object, and survey\_map object.
-These three objects come together to create a mapping object, which is what processes such as regression act on.
-Finally, common diagnostics and graphs are implemented.
-In this way, our grammar and package implement an entire statistical workflow for conducting MRP.
+We first define a grammar of MRP, which we define as the underlying
+rules and principles that are common to every analysis based on MRP.
+This grammar is based around the following verbs: `add', `remove',
+`delete', and `replace'. These verbs are applied to a survey object, a
+post-stratification object, and survey\_map object. These three objects
+come together to create a mapping object, which is what processes such
+as regression act on. Finally, common diagnostics and graphs are
+implemented. In this way, our grammar and package implement an entire
+statistical workflow for conducting MRP.
 
-The survey object would typically be a regular survey, such as a political poll of 1,000 respondents, but it could also be a larger survey, such as the Canadian Election Survey, or similar.
-The post-stratification object would typically be a larger survey, such as, in the case of the US, the (INSERT THE USUAL ONE), or a census.
+The survey object would typically be a regular survey, such as a
+political poll of 1,000 respondents, but it could also be a larger
+survey, such as the Canadian Election Survey, or similar. The
+post-stratification object would typically be a larger survey, such as,
+in the case of the US, the (INSERT THE USUAL ONE), or a census.
 
-Our grammar and package complements existing packages such as \proglang{survey} (ADD CITATION), X and Y. These packages... Instead, ours is focused on what is needed for MRP.
+Our grammar and package complements existing packages such as
+\proglang{survey} (ADD CITATION), X and Y. These packages\ldots{}
+Instead, ours is focused on what is needed for MRP.
 
-The remainder of this paper is structured as follows: Section \ref{sec:review} reviews similar packages and contributions and places ours within that context.
-Section \ref{sec:components} discusses the grammar and the core aspects of MRP as implemented in \proglang{MRPKit}.
-Section \ref{sec:implementation} discusses some of the implementation issues and technical notes related to the decisions that were made.
-Section \ref{sec:vignette} provides two examples of the package in use, one using SOMETHING, and the other using SOMETHING ELSE.
-Finally, Section \ref{sec:summary} provides a summary discussion, some cautions and weaknesses, as well as notes about next steps.
+The remainder of this paper is structured as follows: Section
+\ref{sec:review} reviews similar packages and contributions and places
+ours within that context. Section \ref{sec:components} discusses the
+grammar and the core aspects of MRP as implemented in \proglang{MRPKit}.
+Section \ref{sec:implementation} discusses some of the implementation
+issues and technical notes related to the decisions that were made.
+Section \ref{sec:vignette} provides two examples of the package in use,
+one using SOMETHING, and the other using SOMETHING ELSE. Finally,
+Section \ref{sec:summary} provides a summary discussion, some cautions
+and weaknesses, as well as notes about next steps.
 
@@
 
-The most common alternative at the moment to this package is for users to do all aspects themselves.
-While there is nothing conceptually difficult about MRP, the implementation can be difficult.
-In particular, preparing and matching different levels between surveys can be time consuming and potentially introduce errors.
+The most common alternative at the moment to this package is for users
+to do all aspects themselves. While there is nothing conceptually
+difficult about MRP, the implementation can be difficult. In particular,
+preparing and matching different levels between surveys can be time
+consuming and potentially introduce errors.
 
@@
 
-MRP package - https://github.com/gelman/mrp
+MRP package - \url{https://github.com/gelman/mrp}
 
@@
 
-The \proglang{MRPKit} package has the following key components: survey objects, which for most users will be a collection of two surveys where one is larger than the other; a survey map, that relates the survey objects and then an MRP object, which is created once the survey map is applied to the survey objects.
-Regression acts on the MRP object.
-These objects are subject to the following verbs: 'add', 'delete', 'new', and 'replace'.
-For instance: \proglang{SurveyMap\$add}, \proglang{SurveyOb\$delete}, \proglang{SurveyMap\$new}, and \proglang{SurveyMap\$replace}.
+The \proglang{MRPKit} package has the following key components: survey
+objects, which for most users will be a collection of two surveys where
+one is larger than the other; a survey map, that relates the survey
+objects and then an MRP object, which is created once the survey map is
+applied to the survey objects. Regression acts on the MRP object. These
+objects are subject to the following verbs: `add', `delete', `new', and
+`replace'. For instance: \proglang{SurveyMap\$add},
+\proglang{SurveyOb\$delete}, \proglang{SurveyMap\$new}, and
+\proglang{SurveyMap\$replace}.
 
@@
 
-Survey objects are the surveys that MRP will bring together.
-A user can \proglang{add} a survey object by providing a CSV file location.
-The user then needs to identify the data types of each column.
-Typically, there will need to be two survey objects, where one will be the survey that is of interest for its response variables, such as political opinion, and another will be a survey that is to be used for post-stratification.
+Survey objects are the surveys that MRP will bring together. A user can
+\proglang{add} a survey object by providing a CSV file location. The
+user then needs to identify the data types of each column. Typically,
+there will need to be two survey objects, where one will be the survey
+that is of interest for its response variables, such as political
+opinion, and another will be a survey that is to be used for
+post-stratification.
 
-New is when you have a new survey object and you are bringing it in. e.g.
+New is when you have a new survey object and you are bringing it in.
+e.g.
 
-% sampl <- surveyStr$new(...)
-% pop <- surveyStr$new(...)
-% map <- surveyMap$new(sampl, pop, ...)
-% mrp1 <- MRP$new(sampl, pop, map)
+\% sampl \textless- surveyStr\(new(...) % pop <- surveyStr
+\)new(\ldots) \% map \textless-
+surveyMap\(new(sampl, pop, ...) % mrp1 <- MRP
+\)new(sampl, pop, map)
 
-% option 1
-% mod1 <- mrp1$fit(y ~ x + (1|group))
-% mod1 <- mrp1$fit(y ~ x + (1|group) + (1|group2))
+\% option 1 \% mod1 \textless-
+mrp1\(fit(y ~ x + (1|group)) % mod1 <- mrp1
+\)fit(y \textasciitilde{} x + (1\textbar group) + (1\textbar group2))
 
-% mod1$poststratify()
-% mod2$poststratify("model1")
+\% mod1\(poststratify() % mod2
+\)poststratify(``model1'')
 
-% or
+\% or
 
-% option 2
-% mrp1$fit(y ~ x + (1|group), name = "model1")
-% mrp1$fit(y ~ x + (1|group) + (1|group2), name = "model2")
+\% option 2 \% mrp1\(fit(y ~ x + (1|group), name = "model1") % mrp1
+\)fit(y \textasciitilde{} x + (1\textbar group) + (1\textbar group2),
+name = ``model2'')
 
-% mrp1$poststartify("model1")
-% mrp1$poststartify("model2")
+\% mrp1\(poststartify("model1") % mrp1
+\)poststartify(``model2'')
 
-What is the difference between add and new? What is replace doing? URGH, I just need to look at the repo.
+What is the difference between add and new? What is replace doing? URGH,
+I just need to look at the repo.
 
@@
 
-A `SurveyMap` object holds the mapping between a set of items in a survey and a population dataset.
-The label is the item label in each dataset and the values is a list of all possible values.
-The values for the survey and population must be aligned, i.e., the lists must have the same number of elements and the values at index i in each list are equivalent.
-If there is a meaningful ordering over the values, they should be listed in that order, either descending or ascending.
+A \texttt{SurveyMap} object holds the mapping between a set of items in
+a survey and a population dataset. The label is the item label in each
+dataset and the values is a list of all possible values. The values for
+the survey and population must be aligned, i.e., the lists must have the
+same number of elements and the values at index i in each list are
+equivalent. If there is a meaningful ordering over the values, they
+should be listed in that order, either descending or ascending.
 
@@
 
-An  MRP object contains survey objects, and a survey map. At the point at which the survey objects are put into the MRP object they become immutable but the survey map object is.
+An MRP object contains survey objects, and a survey map. At the point at
+which the survey objects are put into the MRP object they become
+immutable but the survey map object is.
 
-The MRP object outputs an analysis. There would be a MRP model fit class that would
+The MRP object outputs an analysis. There would be a MRP model fit class
+that would
 
@@
 \section*{Acknowledgments}
+
+\end{document}

### 2021-03-03 R/sim_posterior_epred.R
cleanup sim_posterior_epred

@@ -1,39 +1,34 @@
-
-#' sim_posterior_epred
+#' Generate approximate samples of predicted probabilities from a binomial
+#' `glmerMod` object.
 #'
-#' @name sim_posterior_epred
 #' @export
-#' @description Helper function to create predictive samples from a glmer class model
-#' Takes in model class and newdata to predict for
-#' Return a matrix with rows that represent posterior samples and columns that represent
-#' data that is being predicted for
-#' Only for binomial family models.
-#' @param object Model of type glmer class and binomial family
-#' @param newdata Data to predict for
-#' @param nsamples Number of samples
+#' @param object A binomial model fit using `lme4::glmer()`.
+#' @param newdata The data to use.
+#' @param nsamples Number of samples to generate.
+#' @return A matrix with `nrow(newdata)` rows and `nsamples` columns.
 #' @examples
-
-
-
-sim_posterior_epred  <- function(object, newdata, nsamples = 4000){
-  if(!class(object)=="glmerMod"){
-    stop("Object must a model type of glmerMod")
-  }else{
-    if(!family(object)$family=="binomial"){
-      stop("Model must be a binomial model type")
-    }else{
-      #Make predictions using merTools (summary stats not needed as we use this as a convenience to get
-      #the predicted samples)
-      predict_nd <- merTools::predictInterval(merMod = object, newdata = newdata,
-                                                         level = 0.95, n.sims = nsamples,
-                                                         stat = "median", type="probability",
-                                                         include.resid.var = TRUE,
-                                                         returnSims = TRUE,
-                                                         .parallel=FALSE)
-      #Obtain predictive samples
-      predicted_results <- plogis(attributes(predict_nd)$sim.results)
-
-      return(t(predicted_results))
-    }
+#'
+sim_posterior_epred  <- function(object, newdata, nsamples = 4000) {
+  if (!requireNamespace("merTools", quietly = TRUE)) {
+    stop("Please install the merTools package.", call. = FALSE)
   }
+  if (!inherits(object, "glmerMod")) {
+    stop("Object must have class 'glmerMod'.", call. = FALSE)
+  }
+  if (family(object)$family != "binomial") {
+    stop("Model family must be binomial.", call. = FALSE)
+  }
+  predict_nd <-
+    merTools::predictInterval(
+      merMod = object,
+      newdata = newdata,
+      level = 0.95,
+      n.sims = nsamples,
+      stat = "median",
[…]

### 2021-03-03 R/misc.R
create misc.R with some helper functions

@@ -1,0 +1,26 @@
+require_suggested_package <- function(pkg) {
+  if (!requireNamespace(pkg, quietly = TRUE)) {
+    stop("Please install the ", pkg, " package.", call. = FALSE)
+  }
+}
+
+is.family <- function(x) {
+  inherits(x, "family")
+}
+
+#' Check if family is binomial/bernoulli
+#' @noRd
+#' @param x User's specified `family` argument.
+#' @return `TRUE` or `FALSE` if no error.
+family_is_binomial <- function(x) {
+  if (is.null(x)) {
+    return(FALSE)
+  }
+  if (!is.character(x) && !is.family(x)) {
+    stop("Model family must be a string or family object", call. = FALSE)
+  }
+  if (is.family(x)) {
+    x <- x$family
+  }
+  x %in% c("binomial", "bernoulli")
+}

### 2021-03-03 R/SurveyMap.R
error if not binomial model (if applicable)

@@ -313,6 +313,11 @@ SurveyMap <- R6::R6Class(
         stop("The 'data' argument should not be specified.",
              call. = FALSE)
       }
+      if ("family" %in% names(formals(fun)) &&
+          !family_is_binomial(args$family)) {
+        stop("Currently only binomial and bernoulli models are supported.",
+             call. = FALSE)
+      }
       if (is.null(self$samp_obj$mapped_data)) {
         stop("Mapped data not found. ",
              "Please call the mapping() method before fitting a model.",

### 2021-03-03 R/SurveyMap.R
minor code cleaning

@@ -193,6 +193,7 @@ SurveyMap <- R6::R6Class(
     replace = function(old_question, new_question) {
       self$delete(old_question)
       self$add(new_question)
+      invisible(self)
     },
 
     validate = function() {
@@ -245,6 +246,7 @@ SurveyMap <- R6::R6Class(
       if (sum(!samp_dfnames %in% c(samp_mapnames,popn_dfnames)) == 0) {
         warning("At least one variable in the survey needs to be unknown in the population.",call. = FALSE)
       }
+      invisible(self)
     },
     mapping  = function() {
       ####Set up keys####
@@ -282,6 +284,7 @@ SurveyMap <- R6::R6Class(
         self$samp_obj$mapped_data[[new_varname]] <- forcats::fct_recode(self$samp_obj$survey_data[[samp_mapnames]], !!!new_levels_samp)
         self$popn_obj$mapped_data[[new_varname]] <- forcats::fct_recode(self$popn_obj$survey_data[[popn_mapnames]], !!!new_levels_popn)
       }
+      invisible(self)
     },
     tabulate  = function(...) {
       grouping_vars <- c(...)
@@ -298,13 +301,16 @@ SurveyMap <- R6::R6Class(
       invisible(self)
     },
 
-    #' @description Fit a model
+    #' @description Fit a model.
     #' @param fun The model fitting function to use. For example,
-    #'   `fun=rstanarm::stan_glmer` or `fun=brms::brm`.
+    #'   `fun=rstanarm::stan_glmer`, `fun=brms::brm`, `fun=lme4::glmer`.
+    #'   If using a custom `fun` it must have a `data` argument that accepts
+    #'   a data frame (like standard R modeling functions).
     #' @param formula The model formula. Can be either a string or a formula
     #'   object.
-    #' @param ... Arguments other than `formula` and `data` to pass to the model
-    #'   fitting function.
+    #' @param ... Arguments other than `formula` and `data` to pass to `fun`.
+    #'   The data argument will be automatically specified internally.
+    #' @return The fitted model object returned by `fun`.
     #'
     fit = function(fun, formula, ...) {
       fun <- match.fun(fun)
@@ -345,7 +351,7 @@ SurveyMap <- R6::R6Class(
     #'   function then the first argument should take in the fitted model object
     #'   and the second argument should take in the poststratification
     #'   (`newdata`) data frame. The function must return a matrix with rows
-    #'   corresponding to the columns of the poststrat data and columns
[…]

### 2021-03-03 R/sim_posterior_epred.R

@@ -9,9 +9,7 @@
 #' @examples
 #'
 sim_posterior_epred  <- function(object, newdata, nsamples = 4000) {
-  if (!requireNamespace("merTools", quietly = TRUE)) {
-    stop("Please install the merTools package.", call. = FALSE)
-  }
+  require_suggested_package("merTools")
   if (!inherits(object, "glmerMod")) {
     stop("Object must have class 'glmerMod'.", call. = FALSE)
   }

### 2021-03-03 R/SurveyObj.R
cleanup SurveyObject doc

@@ -2,31 +2,58 @@
 #'
 #' @name SurveyObject
 #' @export
-#' @description A `SurveyObject` represents a survey and its metadata.
-#' The survey itself is a dataframe.
-#' The survey metatdata consists of
-#'  - per-column questions - a list of strings
-#'  - per-column allowed response values - a list of lists
-#'  - per-column survey weights, - a vector of numeric weights
-#'  - a string that specifies the survey design using lmer forumula syntax
+#' @description
+#' A `SurveyObject` represents a survey and its metadata.
+#' The survey itself is a data frame.
+#' The survey metatdata consists of the following:
+#'  - per-column questions: a list of strings
+#'  - per-column allowed response values: a list of character vectors
+#'  - per-column survey weights: a vector of numeric weights
+#'  - survey design: a string that specifies the survey design using \pkg{lme4}
+#'  style forumula syntax.
 #'
-#' Examples of survey design include:
-#' ~. a random sample, ~ (1|cluster), a one stage cluster sample, ~ stratum, a stratified sample
+#' Examples of survey designs include:
+#'   - `~.`: a random sample
+#'   - `~ (1|cluster)`: a one stage cluster sample
+#'   - `~ stratum`: a stratified sample
 #'
 #' @examples
-#' feline_prefs = SurveyObj$new(feline_survey[,c("age1","gender","pet_own","y")],
-#' questions = c("Please identify your age group","Please select your gender","Which pet do you own?", "Response"),
-#' responses = list(levels(feline_survey$age1),levels(feline_survey$gender),levels(feline_survey$pet_own),c("no","yes")),
-#' weights = feline_survey$wt,
-#' design = formula("~."))
+#' feline_prefs2 <- SurveyObj$new(
+#'   feline_survey[,c("age1","gender","pet_own","y")],
+#'   questions = c(
+#'     "Please identify your age group",
+#'     "Please select your gender",
+#'     "Which pet do you own?",
+#'     "Response"
+#'   ),
+#'   responses = list(
+#'     levels(feline_survey$age1),
+#'     levels(feline_survey$gender),
+#'     levels(feline_survey$pet_own),
+#'     c("no","yes")
+#'   ),
+#'   weights = feline_survey$wt,
+#'   design = formula("~.")
+#' )
 #' feline_prefs$print()
-#'popn_obj = SurveyObj$new(approx_popn[,c("age2","gender","pet_pref")],
-#' questions = c("Which age group are you?","Gender?","Which pet would you like to own?"),
[…]

### 2021-03-03 R/SurveyQuestion.R

@@ -13,7 +13,6 @@
 #' The ordering should be sample first and then population
 #'
 #' @examples
-#'
 #' q1 <- question$new(
 #'   name = "age",
 #'   col_names = c("age1","age2"),

### 2021-03-03 R/SurveyMap.R
rename SurveyObj to SurveyData

@@ -13,14 +13,14 @@
 #' they should be listed in that order, either descending or ascending.
 #'
 #' @examples
-#' feline_prefs = SurveyObj$new(
+#' feline_prefs = SurveyData$new(
 #'   feline_survey[,c("age1","gender","pet_own","y")],
 #'   questions = c("Please identify your age group","Please select your gender","Which pet do you own?", "Response"),
 #'   responses = list(levels(feline_survey$age1),levels(feline_survey$gender),levels(feline_survey$pet_own),c("no","yes")),
 #'   weights = feline_survey$wt,
 #'   design = formula("~.")
 #' )
-#' popn_obj = SurveyObj$new(
+#' popn_obj = SurveyData$new(
 #'   approx_popn[,c("age2","gender","pet_pref")],
 #'   questions = c("Which age group are you?","Gender?","Which pet would you like to own?"),
 #'   responses = list(levels(approx_popn$age2),levels(approx_popn$gender),levels(approx_popn$pet_pref)),
@@ -117,11 +117,11 @@ SurveyMap <- R6::R6Class(
     popn_obj = NULL,
 
     initialize = function(samp_obj, popn_obj, ...) {
-      if (!inherits(samp_obj, "SurveyObj")) {
-        stop("samp_obj must be a SurveyObj object.", call. = FALSE)
+      if (!inherits(samp_obj, "Survey")) {
+        stop("samp_obj must be a Survey object.", call. = FALSE)
       }
-      if (!inherits(popn_obj, "SurveyObj")) {
-        stop("popn_obj must be a SurveyObj object.", call. = FALSE)
+      if (!inherits(popn_obj, "Survey")) {
+        stop("popn_obj must be a SurveyData object.", call. = FALSE)
       }
 
       self$item_map <- list(...)

### 2021-03-03 R/SurveyObj.R

@@ -1,16 +1,16 @@
-#' SurveyObject
+#' SurveyData objects
 #'
-#' @name SurveyObject
+#' @name SurveyData
 #' @export
 #' @description
-#' A `SurveyObject` represents a survey and its metadata.
+#' A `SurveyData` object represents a survey and its metadata.
 #' The survey itself is a data frame.
 #' The survey metatdata consists of the following:
 #'  - per-column questions: a list of strings
 #'  - per-column allowed response values: a list of character vectors
 #'  - per-column survey weights: a vector of numeric weights
 #'  - survey design: a string that specifies the survey design using \pkg{lme4}
-#'  style forumula syntax.
+#'  style formula syntax.
 #'
 #' Examples of survey designs include:
 #'   - `~.`: a random sample
@@ -18,7 +18,7 @@
 #'   - `~ stratum`: a stratified sample
 #'
 #' @examples
-#' feline_prefs2 <- SurveyObj$new(
+#' feline_prefs2 <- SurveyData$new(
 #'   feline_survey[,c("age1","gender","pet_own","y")],
 #'   questions = c(
 #'     "Please identify your age group",
@@ -37,7 +37,7 @@
 #' )
 #' feline_prefs$print()
 #'
-#' popn_obj <- SurveyObj$new(
+#' popn_obj <- SurveyData$new(
 #'   approx_popn[,c("age2","gender","pet_pref")],
 #'   questions = c(
 #'     "Which age group are you?",
@@ -54,8 +54,8 @@
 #' )
 #' popn_obj$print()
 #'
-SurveyObj <- R6::R6Class(
-    classname = "SurveyObj",
+SurveyData <- R6::R6Class(
+    classname = "SurveyData",
     public = list(
         survey_data = data.frame(NULL),
         poststrat = data.frame(NULL),

### 2021-03-03 R/SurveyMap.R
question -> SurveyQuestion (plus doc)

@@ -27,7 +27,7 @@
 #'   weights = approx_popn$wt,
 #'   design = formula("~.")
 #' )
-#' q1 <- question$new(
+#' q1 <- SurveyQuestion$new(
 #'   name = "age",
 #'   col_names = c("age1","age2"),
 #'   values_map = list(
@@ -35,12 +35,12 @@
 #'     "46-55" = "36-55", "56-65" = "56-65", "66-75" = "66+", "76-90" = "66+"
 #'   )
 #' )
-#' q2 <- question$new(
+#' q2 <- SurveyQuestion$new(
 #'   name = "pet",
 #'   col_names = c("pet_own","pet_pref"),
 #'   values_map = list("cat" = "cat", "kitten" = "cat","dog" = "dog","puppy" = "dog")
 #' )
-#' q3 <- question$new(
+#' q3 <- SurveyQuestion$new(
 #'   name = "gender",
 #'   col_names = c("gender","gender"),
 #'   values_map = data.frame("male" = "m","female" = "f", "nonbinary" = "nb")

### 2021-03-03 R/SurveyQuestion.R

@@ -6,14 +6,14 @@
 #' @description A `SurveyQuestion` object holds the mapping for one question
 #' or demographic between the survey and population dataset.
 #' The name is the name of the underlying construct.
-#' The col_names maps the survey column name to the population column name.
-#' The values map the survey response values to the population values.
+#' The `col_names` map the survey column name to the population column name.
+#' The `values` map the survey response values to the population values.
 #' If there is a meaningful ordering over the values,
 #' they should be listed in that order, either descending or ascending.
 #' The ordering should be sample first and then population
 #'
 #' @examples
-#' q1 <- question$new(
+#' q1 <- SurveyQuestion$new(
 #'   name = "age",
 #'   col_names = c("age1","age2"),
 #'   values_map = list(
@@ -21,23 +21,37 @@
 #'     "46-55" = "36-55", "56-65" = "56-65", "66-75" = "66+", "76-90" = "66+"
 #'   )
 #' )
-#' q2 <- question$new(
+#' q2 <- SurveyQuestion$new(
 #'   name = "pet",
 #'   col_names = c("pet_pref","pet_own"),
 #'   values_map = list("cat" = "cat", "kitten" = "cat","dog" = "dog","dog" = "puppy")
 #' )
-#' q3 <- question$new(
+#' q3 <- SurveyQuestion$new(
 #'   name = "gender",
 #'   col_names = c("gender","gender"),
-#'   values_map = data.frame("male" = "m","female" = "f", "nonbinary" = "nb")
+#'   values_map = list("male" = "m","female" = "f", "nonbinary" = "nb")
 #' )
 #'
-question <- R6::R6Class(
+SurveyQuestion <- R6::R6Class(
   classname = "SurveyQuestion",
   public = list(
     name = character(),
     col_names = character(),
-    values = data.frame(),
+    values = list(),
+
+    #' @description Create a new `SurveyQuestion`
+    #' @param name The name of the underlying construct. For example if the
+    #'   sample data uses the name `age1` and the population data uses
+    #'   the name `age2` then `name` could be `age`.
+    #' @param col_names A character vector of length two giving the column names
+    #'   of the question in the sample data (first element) and the population
+    #'   data (second element).
+    #' @param values_map A named list where the names correspond to the
+    #'   responses to the question in the sample data and the values correspond
+    #'   to the responses in the population data.
+    #' @return A `SurveyQuestion` object that can be added to a [`SurveyData`]
+    #'   object.
+    #'
[…]

### 2021-03-03 paper/paper_joss/paper_joss.Rmd
Start updating paper.

--- 2020-09-30 paper/paper.Rmd
@@
     affiliation: Monash University
-    email: \email{name@company.com}
-    url: http://rstudio.com
+    email: \email{lauren.kennedy729@gmail.com}
+    url: https://jazzystats.com
   - name: Mitzi Morris
@@
   # For running headers, if needed
-  short:     "\\pkg{foo}: A grammar of MRP"
+  short:     "\\pkg{MRP-Kit}: A grammar of MRP"
 abstract: >
-  In this paper we define a grammar for multilevel regression with post-stratification (MRP) and implement this grammar in the R Package \pkg{MRPKit}.
+  In this paper we define a 'grammar' for multilevel regression with post-stratification (MRP) and implement this grammar in the R Package \pkg{MRPKit}.
 Our grammar is centered around the following verbs: add, remove, delete, and replace.
@@
 output: rticles::jss_article
+bibliography: references.bib
 ---
@@
 
-Our grammar and package complements existing packages such as \proglang{survey} (ADD CITATION), X and Y. These packages... Instead, ours is focused on what is needed for MRP.
+Our grammar and package complements existing packages such as \proglang{survey} [@lumleyone, @lumleytwo, @lumleythree], and \proglang{DeclareDesign} [@citeDeclareDesign].
+These packages are focused on the designing, implementing, and simulating from, surveys.
+Instead, ours is focused on what is needed for MRP.
 
@@
 The most common alternative at the moment to this package is for users to do all aspects themselves.
-While there is nothing conceptually difficult about MRP, the implementation can be difficult.
-In particular, preparing and matching different levels between surveys can be time consuming and potentially introduce errors.
+While there is nothing inherently difficult about MRP, the implementation can be difficult.
+In particular, preparing and matching different levels between surveys can be time consuming and potentially introduce undocumented errors.
 
-HERE
+The \proglang{survey} package [@lumleyone, @lumleytwo, @lumleythree] ...
 
-Survey package
+The \proglang{DeclareDesign} package [@citeDeclareDesign]...
 
-Declare design
+There is also an existing package called \proglang{MRP} available here: https://github.com/gelman/mrp. However it does not appear to have been updated in some time.
 
-DIY
-
-MRP package - https://github.com/gelman/mrp
-
-And how we're different
+The main alternative approach is for MRP to be conducted on a case-specific basis.
+That is a researcher interested in MRP estimates, writes the code needed to obtain, clean, prepare, analyze and ultimately interpret the estimates.
+Again, while there is nothing inherently wrong with this approach, successfully conducting MRP requires dealing with a large number of small issues.
+Each of these is small in their own right, but getting them wrong can have large effects that are potentially unnoticed on the estimates.
 
@@
 For instance: \proglang{SurveyMap\$add}, \proglang{SurveyOb\$delete}, \proglang{SurveyMap\$new}, and \proglang{SurveyMap\$replace}.
+
+new()
+validate() 
+add() 
+delete() 
+replace()
+validate()
+mapping()
+tabulate()
+fit() 
+predictify() 
+collapsify() 
+visify()
 
@@
 
-% sampl <- surveyStr$new(...)
-% pop <- surveyStr$new(...)
-% map <- surveyMap$new(sampl, pop, ...)
-% mrp1 <- MRP$new(sampl, pop, map)
-
-% option 1
-% mod1 <- mrp1$fit(y ~ x + (1|group))
-% mod1 <- mrp1$fit(y ~ x + (1|group) + (1|group2))
-
-% mod1$poststratify()
-% mod2$poststratify("model1")
-
-% or
-
-% option 2
-% mrp1$fit(y ~ x + (1|group), name = "model1")
-% mrp1$fit(y ~ x + (1|group) + (1|group2), name = "model2")
-
-% mrp1$poststartify("model1")
-% mrp1$poststartify("model2")
-
-What is the difference between add and new? What is replace doing? URGH, I just need to look at the repo.
+<!-- #' feline_prefs = SurveyObj$new( -->
+<!-- #'   feline_survey[,c("age1","gender","pet_own","y")], -->
+<!-- #'   questions = c("Please identify your age group","Please select your gender","Which pet do you own?", "Response"), -->
+<!-- #'   responses = list(levels(feline_survey$age1),levels(feline_survey$gender),levels(feline_survey$pet_own),c("no","yes")), -->
+<!-- #'   weights = feline_survey$wt, -->
+<!-- #'   design = formula("~.") -->
+<!-- #' ) -->
+<!-- #' popn_obj = SurveyObj$new( -->
+<!-- #'   approx_popn[,c("age2","gender","pet_pref")], -->
+<!-- #'   questions = c("Which age group are you?","Gender?","Which pet would you like to own?"), -->
+<!-- #'   responses = list(levels(approx_popn$age2),levels(approx_popn$gender),levels(approx_popn$pet_pref)), -->
+<!-- #'   weights = approx_popn$wt, -->
+<!-- #'   design = formula("~.") -->
+<!-- #' ) -->
+<!-- #' q1 <- question$new( -->
+<!-- #'   name = "age", -->
+<!-- #'   col_names = c("age1","age2"), -->
+<!-- #'   values_map = list( -->
+<!-- #'     "18-25" = "18-35", "26-35" = "18-35","36-45" = "36-55", -->
+<!-- #'     "46-55" = "36-55", "56-65" = "56-65", "66-75" = "66+", "76-90" = "66+" -->
+<!-- #'   ) -->
+<!-- #' ) -->
+<!-- #' q2 <- question$new( -->
+<!-- #'   name = "pet", -->
+<!-- #'   col_names = c("pet_own","pet_pref"), -->
+<!-- #'   values_map = list("cat" = "cat", "kitten" = "cat","dog" = "dog","puppy" = "dog") -->
+<!-- #' ) -->
+<!-- #' q3 <- question$new( -->
+<!-- #'   name = "gender", -->
+<!-- #'   col_names = c("gender","gender"), -->
+<!-- #'   values_map = data.frame("male" = "m","female" = "f", "nonbinary" = "nb") -->
+<!-- #' ) -->
+<!-- #' -->
+<!-- #' tmp_map <- SurveyMap$new(samp_obj = feline_prefs, popn_obj = popn_obj, q1) -->
+<!-- #' print(tmp_map) -->
+<!-- #' tmp_map$validate() -->
+<!-- #' tmp_map$add(q3) -->
+<!-- #' print(tmp_map) -->
+<!-- #' tmp_map$delete(q3) -->
+<!-- #' print(tmp_map) -->
+<!-- #' tmp_map$add(q3) -->
+<!-- #' tmp_map$delete("gender") -->
+<!-- #' print(tmp_map) -->
+<!-- #' tmp_map$add(q2) -->
+<!-- #' print(tmp_map) -->
+<!-- #' tmp_map$replace(q1,q3) -->
+<!-- #' print(tmp_map) -->
+<!-- #' tmp_map$add(q1) -->
+<!-- #' print(tmp_map) -->
+<!-- #' tmp_map$validate() # TODO: figure out why this errors -->
+<!-- #' tmp_map$mapping() -->
+<!-- #' tmp_map$tabulate("age") #Just use age in the poststrat matrix -->
+<!-- #' tmp_map$tabulate() #Use all variables in the map -->
+<!-- #' -->
+<!-- #' -->
+<!-- #' mod_fit_1 <- tmp_map$fit( -->
+<!-- #'   fun = rstanarm::stan_glmer, -->
+<!-- #'   formula = y ~ (1|age) + (1|gender), -->
+<!-- #'   family = "binomial", -->
+<!-- #'   refresh = 100, -->
+<!-- #'   cores = 2 -->
+<!-- #' ) -->
+<!-- #' -->
+<!-- #' mod_fit_2 <- tmp_map$fit( -->
+<!-- #'   fun = brms::brm, -->
+<!-- #'   formula = y ~ (1|age) + (1|gender), -->
+<!-- #'   family = "bernoulli", -->
+<!-- #'   refresh = 100, -->
+<!-- #'   cores = 2 -->
+<!-- #' ) -->
+<!-- #' -->
+<!-- #' class(mod_fit_1) -->
+<!-- #' -->
+<!-- #' # predict in postrat matrix - returns a matrix -->
+<!-- #' # with rows as poststrat rows, cols as posterior samples. -->
+<!-- #' poststrat_fit <- tmp_map$predictify(mod_fit_2) -->
+<!-- #' -->
+<!-- #' # get an estimate for a particular variable level or population -->
+<!-- #' # arguments: -->
+<!-- #' # - output of "predictify" -->
+<!-- #' # - (optional) variable name -->
+<!-- #' # body: -->
+<!-- #' # - if no variable name then compute population estimate -->
+<!-- #' # - if variable name specified compute weighted mean (N_j * theta)/sum(N_j) -->
+<!-- #' # return type: -->
+<!-- #' # - data.frame -->
+<!-- #' #   - one column if popn estimate, otherwise one column per level of variable -->
+<!-- #' #   - one row per posterior sample -->
+<!-- #' sae_preds <- tmp_map$collapsify(poststrat_fit, variable_aggr = "age") -->
+<!-- #' popn_preds <- tmp_map$collapsify(poststrat_fit) -->
+<!-- #' plot1 <- tmp_map$visify(sae_preds) -->
+<!-- #' plot1 -->
+<!-- #' plot2 <- tmp_map$visify(popn_preds) -->
+<!-- #' plot2 -->
 
@@
 
+\subsection*{Simulated data}
+
 \subsection*{Using CCES data}

### 2021-03-03 paper/paper_joss/paper_joss.tex
Start updating paper.

--- 2020-10-04 paper/paper.tex
@@
 \author{
-Lauren Kennedy\\Monash University \And Mitzi Morris\\Columbia University \And Jonah Gabry\\Columbia University \And Rohan Alexander\\University of Toronto
+Lauren Kennedy\\Monash University \And Mitzi Morris\\Columbia
+University \And Jonah Gabry\\Columbia University \And Rohan
+Alexander\\University of Toronto
 }
@@
 \Plainauthor{Lauren Kennedy, Mitzi Morris, Jonah Gabry, Rohan Alexander}
-\Plaintitle{A `grammar' of multilevel regression with post-stratification and
-implementation in R}
-\Shorttitle{\pkg{foo}: A grammar of MRP}
+\Plaintitle{A `grammar' of multilevel regression with
+post-stratification and implementation in R}
+\Shorttitle{\pkg{MRP-Kit}: A grammar of MRP}
 
 \Abstract{
-In this paper we define a grammar for multilevel regression with
+In this paper we define a `grammar' for multilevel regression with
 post-stratification (MRP) and implement this grammar in the R Package
@@
 
-\Keywords{multilevel regression with post-stratification, \proglang{R}, reproducibility, statistics, political science}
-\Plainkeywords{multilevel regression with post-stratification, R, reproducibility, statistics, political science}
+\Keywords{multilevel regression with
+post-stratification, \proglang{R}, reproducibility, statistics, political
+science}
+\Plainkeywords{multilevel regression with
+post-stratification, R, reproducibility, statistics, political science}
 
@@
         }
+
+% Pandoc citation processing
 
@@
 Our grammar and package complements existing packages such as
-\proglang{survey} (ADD CITATION), X and Y. These packages\ldots{}
-Instead, ours is focused on what is needed for MRP.
+\proglang{survey} \citep[\citet{lumleytwo},
+\citet{lumleythree}]{lumleyone}, and \proglang{DeclareDesign}
+\citep{citeDeclareDesign}. These packages are focused on the designing,
+implementing, and simulating from, surveys. Instead, ours is focused on
+what is needed for MRP.
 
@@
 The most common alternative at the moment to this package is for users
-to do all aspects themselves. While there is nothing conceptually
+to do all aspects themselves. While there is nothing inherently
 difficult about MRP, the implementation can be difficult. In particular,
 preparing and matching different levels between surveys can be time
-consuming and potentially introduce errors.
+consuming and potentially introduce undocumented errors.
 
-HERE
+The \proglang{survey} package \citep[\citet{lumleytwo},
+\citet{lumleythree}]{lumleyone} \ldots{}
 
-Survey package
+The \proglang{DeclareDesign} package \citep{citeDeclareDesign}\ldots{}
 
-Declare design
+There is also an existing package called \proglang{MRP} available here:
+\url{https://github.com/gelman/mrp}. However it does not appear to have
+been updated in some time.
 
-DIY
-
-MRP package - \url{https://github.com/gelman/mrp}
-
-And how we're different
+The main alternative approach is for MRP to be conducted on a
+case-specific basis. That is a researcher interested in MRP estimates,
+writes the code needed to obtain, clean, prepare, analyze and ultimately
+interpret the estimates. Again, while there is nothing inherently wrong
+with this approach, successfully conducting MRP requires dealing with a
+large number of small issues. Each of these is small in their own right,
+but getting them wrong can have large effects that are potentially
+unnoticed on the estimates.
 
@@
 e.g.
-
-\% sampl \textless- surveyStr\(new(...) % pop <- surveyStr
-\)new(\ldots) \% map \textless-
-surveyMap\(new(sampl, pop, ...) % mrp1 <- MRP
-\)new(sampl, pop, map)
-
-\% option 1 \% mod1 \textless-
-mrp1\(fit(y ~ x + (1|group)) % mod1 <- mrp1
-\)fit(y \textasciitilde{} x + (1\textbar group) + (1\textbar group2))
-
-\% mod1\(poststratify() % mod2
-\)poststratify(``model1'')
-
-\% or
-
-\% option 2 \% mrp1\(fit(y ~ x + (1|group), name = "model1") % mrp1
-\)fit(y \textasciitilde{} x + (1\textbar group) + (1\textbar group2),
-name = ``model2'')
-
-\% mrp1\(poststartify("model1") % mrp1
-\)poststartify(``model2'')
 
@@
 
+\bibliography{references.bib}
+
 \end{document}

### 2021-03-03 R/SurveyQuestion.R
more doc for SurveyQuestion

@@ -3,14 +3,8 @@
 #' @name SurveyQuestion
 #' @export
 #'
-#' @description A `SurveyQuestion` object holds the mapping for one question
-#' or demographic between the survey and population dataset.
-#' The name is the name of the underlying construct.
-#' The `col_names` map the survey column name to the population column name.
-#' The `values` map the survey response values to the population values.
-#' If there is a meaningful ordering over the values,
-#' they should be listed in that order, either descending or ascending.
-#' The ordering should be sample first and then population
+#' @description A `SurveyQuestion` object holds the mapping for one question or
+#'   demographic between the survey and population dataset.
 #'
 #' @examples
 #' q1 <- SurveyQuestion$new(
@@ -21,6 +15,8 @@
 #'     "46-55" = "36-55", "56-65" = "56-65", "66-75" = "66+", "76-90" = "66+"
 #'   )
 #' )
+#' print(q1)
+#'
 #' q2 <- SurveyQuestion$new(
 #'   name = "pet",
 #'   col_names = c("pet_pref","pet_own"),
@@ -28,7 +24,7 @@
 #' )
 #' q3 <- SurveyQuestion$new(
 #'   name = "gender",
-#'   col_names = c("gender","gender"),
+#'   col_names = c("gender", "gender"),
 #'   values_map = list("male" = "m","female" = "f", "nonbinary" = "nb")
 #' )
 #'
@@ -48,7 +44,9 @@ SurveyQuestion <- R6::R6Class(
     #'   data (second element).
     #' @param values_map A named list where the names correspond to the
     #'   responses to the question in the sample data and the values correspond
-    #'   to the responses in the population data.
+    #'   to the responses in the population data. If there is a meaningful
+    #'   ordering over the values, they should be listed in that order, either
+    #'   descending or ascending.
     #' @return A `SurveyQuestion` object that can be added to a [`SurveyData`]
     #'   object.
     #'
@@ -57,7 +55,10 @@ SurveyQuestion <- R6::R6Class(
         stop("Please specify a general name for this construct", call. = FALSE)
       }
       if (anyNA(col_names)) {
-        stop("Please specify the sample and population column names", call. = FALSE)
+        stop("Please specify the sample and population column names.", call. = FALSE)
+      }
+      if (length(col_names) != 2) {
+        stop("'col_names' must have length 2.", call. = FALSE)
       }
       if (anyNA(names(values_map))) {
[…]

### 2021-03-03 R/SurveyMap.R
fix class name in check

@@ -117,10 +117,10 @@ SurveyMap <- R6::R6Class(
     popn_obj = NULL,
 
     initialize = function(samp_obj, popn_obj, ...) {
-      if (!inherits(samp_obj, "Survey")) {
-        stop("samp_obj must be a Survey object.", call. = FALSE)
+      if (!inherits(samp_obj, "SurveyData")) {
+        stop("samp_obj must be a SurveyData object.", call. = FALSE)
       }
-      if (!inherits(popn_obj, "Survey")) {
+      if (!inherits(popn_obj, "SurveyData")) {
         stop("popn_obj must be a SurveyData object.", call. = FALSE)
       }
 

### 2021-03-03 R/misc.R
Update misc.R

@@ -4,7 +4,7 @@ require_suggested_package <- function(pkg) {
   }
 }
 
-is.family <- function(x) {
+is_family <- function(x) {
   inherits(x, "family")
 }
 
@@ -16,10 +16,10 @@ family_is_binomial <- function(x) {
   if (is.null(x)) {
     return(FALSE)
   }
-  if (!is.character(x) && !is.family(x)) {
+  if (!is.character(x) && !is_family(x)) {
     stop("Model family must be a string or family object", call. = FALSE)
   }
-  if (is.family(x)) {
+  if (is_family(x)) {
     x <- x$family
   }
   x %in% c("binomial", "bernoulli")

### 2021-03-03 R/SurveyMap.R
regenerate doc

@@ -62,7 +62,7 @@
 #' print(tmp_map)
 #' tmp_map$add(q1)
 #' print(tmp_map)
-#' tmp_map$validate() # TODO: figure out why this errors
+#' tmp_map$validate()
 #' tmp_map$mapping()
 #' tmp_map$tabulate("age") #Just use age in the poststrat matrix
 #' tmp_map$tabulate() #Use all variables in the map
@@ -83,8 +83,6 @@
 #'   refresh = 100,
 #'   cores = 2
 #' )
-#'
-#' class(mod_fit_1)
 #'
 #' # predict in postrat matrix - returns a matrix
 #' # with rows as poststrat rows, cols as posterior samples.

### 2021-03-04 R/SurveyMap.R
also allow checking version of suggested package

@@ -13,14 +13,14 @@
 #' they should be listed in that order, either descending or ascending.
 #'
 #' @examples
-#' feline_prefs = SurveyData$new(
+#' feline_prefs <- SurveyData$new(
 #'   feline_survey[,c("age1","gender","pet_own","y")],
 #'   questions = c("Please identify your age group","Please select your gender","Which pet do you own?", "Response"),
 #'   responses = list(levels(feline_survey$age1),levels(feline_survey$gender),levels(feline_survey$pet_own),c("no","yes")),
 #'   weights = feline_survey$wt,
 #'   design = formula("~.")
 #' )
-#' popn_obj = SurveyData$new(
+#' popn_obj <- SurveyData$new(
 #'   approx_popn[,c("age2","gender","pet_pref")],
 #'   questions = c("Which age group are you?","Gender?","Which pet would you like to own?"),
 #'   responses = list(levels(approx_popn$age2),levels(approx_popn$gender),levels(approx_popn$pet_pref)),
@@ -64,8 +64,8 @@
 #' print(tmp_map)
 #' tmp_map$validate()
 #' tmp_map$mapping()
-#' tmp_map$tabulate("age") #Just use age in the poststrat matrix
-#' tmp_map$tabulate() #Use all variables in the map
+#' tmp_map$tabulate("age") # Just use age in the poststrat matrix
+#' tmp_map$tabulate() # Use all variables in the map
 #'
 #'
 #' mod_fit_1 <- tmp_map$fit(
@@ -371,7 +371,7 @@ SurveyMap <- R6::R6Class(
 
       if (is.null(args$fun)) {
         if ("stanreg" %in% class(fitted_model)){
-          require_suggested_package("rstanarm")
+          require_suggested_package("rstanarm", "2.21.0")
           return(
             t(rstanarm::posterior_epred(
               object = fitted_model,

### 2021-03-04 R/misc.R

@@ -1,6 +1,9 @@
-require_suggested_package <- function(pkg) {
-  if (!requireNamespace(pkg, quietly = TRUE)) {
-    stop("Please install the ", pkg, " package.", call. = FALSE)
+require_suggested_package <- function(pkg, ver = NULL) {
+  if (!requireNamespace(pkg, quietly = TRUE) ||
+      (!is.null(ver) && utils::packageVersion(pkg) < ver)) {
+    stop("Please install ",
+         if (!is.null(ver)) paste("at least version", ver, "of "),
+         "the ", pkg, " package.", call. = FALSE)
   }
 }
 

### 2021-03-10 paper/paper_joss/paper_joss.Rmd

--- 2020-09-30 paper/paper.Rmd
@@
     affiliation: University of Toronto
+  - name: Dewi Amaliah
+    affiliation: Monash University
 title:
@@
   # For running headers, if needed
-  short:     "\\pkg{foo}: A grammar of MRP"
+  short:     "\\pkg{MRP-Kit}: A grammar of MRP"
 abstract: >
-  In this paper we define a grammar for multilevel regression with post-stratification (MRP) and implement this grammar in the R Package \pkg{MRPKit}.
-Our grammar is centered around the following verbs: add, remove, delete, and replace.
+  In this paper we define a 'grammar' for multilevel regression with post-stratification (MRP) and implement this grammar in the R Package \pkg{MRPKit}.
+Our grammar is centered around the following verbs: new, validate, add, delete, replace, mapping, tabulate, fit, predictify, collapsify, visify.
 These act on survey objects.
@@
   # at least one keyword must be supplied
-  formatted: [multilevel regression with post-stratification, "\\proglang{R}", reproducibility, statistics, political science]
-  plain:     [multilevel regression with post-stratification, R, reproducibility, statistics, political science]
+  formatted: [multilevel regression with post-stratification, "\\proglang{R}", reproducibility, statistics]
+  plain:     [multilevel regression with post-stratification, R, reproducibility, statistics]
 preamble: >
@@
 Multilevel regression with post-stratification (MRP) is a statistical approach in which surveys are related to each other using a statistical model.
-This is important because known biases in one survey, can be adjusted for by another in a statistically reasonable way.
+This is important because known biases in one survey can be adjusted for by another in a statistically reasonable way.
 This enables better use of non-representative surveys, additional information, and propagation of uncertainty.
-However it can be difficult to use MRP due to this need to related two different datasets.
-This package defines a grammar, or list of underlying rules, of MRP and then describes an R package, \proglang{MRPKit}, that implements this grammar.
+However, it can be difficult to use MRP due to this need to relate (at least) two different datasets.
+This package defines a grammar, or list of underlying rules, for MRP and then describes an R package, \proglang{MRPKit}, that implements this grammar.
 
 At its core, MRP is a mapping between a survey object and a population object.
-It is from this mapping that the power of MRP exists, but it also establishing this mapping that is the difficult part of implementing MRP models.
+It is from this mapping that the power of MRP exists, but it can be difficult to establish this mapping in a reproducible way that fits into a typical applied statistics workflow.
 Making this implementation easier and more reproducible is important as interest in, and the use of, MRP increases.
-It can be difficult even for those experienced with MRP to ensure there are no mistakes in this mapping and easing this is an important contribution to enhancing the reproducibility of MRP analysis.
+It can be difficult even for those experienced with MRP to ensure there are no mistakes in this mapping and creating a well-defined workflow for MRP is an important contribution to enhancing the reproducibility of MRP analysis.
 
 We first define a grammar of MRP, which we define as the underlying rules and principles that are common to every analysis based on MRP.
-This grammar is based around the following verbs: 'add', 'remove', 'delete', and 'replace'.
-These verbs are applied to a survey object, a post-stratification object, and survey\_map object.
-These three objects come together to create a mapping object, which is what processes such as regression act on.
+This grammar is based around the following verbs: 'new', 'validate', 'add', 'delete', 'replace', 'mapping', 'tabulate', 'fit', 'predictify', 'collapsify', and 'visify'.
+These verbs are applied to a survey object, a post-stratification object, and survey-map object.
+These three objects come together to create a mapping object, which is what estimation processes such as regression act on.
 Finally, common diagnostics and graphs are implemented.
-In this way, our grammar and package implement an entire statistical workflow for conducting MRP.
+In this way, our grammar and package implement an entire statistical workflow for MRP.
 
 The survey object would typically be a regular survey, such as a political poll of 1,000 respondents, but it could also be a larger survey, such as the Canadian Election Survey, or similar.
-The post-stratification object would typically be a larger survey, such as, in the case of the US, the (INSERT THE USUAL ONE), or a census.
+The post-stratification object would typically be a larger survey, such as, in the case of the US, the American Community Surveys, or a census.
 
-Our grammar and package complements existing packages such as \proglang{survey} (ADD CITATION), X and Y. These packages... Instead, ours is focused on what is needed for MRP.
+Our grammar and package complements existing packages such as \proglang{survey} [@lumleyone, @lumleytwo, @lumleythree], and \proglang{DeclareDesign} [@citeDeclareDesign].
+These packages are focused on the designing, implementing, and simulating from, surveys.
+Instead, ours is focused on what is needed for MRP, and we draw on their packages where possible.
 
@@
 
-\section{Review of the other packages} \label{sec:review}
+\section{Review of related packages} \label{sec:review}
 
-The most common alternative at the moment to this package is for users to do all aspects themselves.
-While there is nothing conceptually difficult about MRP, the implementation can be difficult.
-In particular, preparing and matching different levels between surveys can be time consuming and potentially introduce errors.
+The most common alternative at the moment to this package is for users to do all aspects of the MRP workflow themselves.
+While there is nothing inherently difficult about MRP, the implementation can be difficult.
+In particular, preparing and matching different levels between surveys can be time consuming and potentially introduce undocumented errors.
 
-HERE
+The \proglang{survey} package [@lumleyone, @lumleytwo, @lumleythree] is well-established and allows a user to specify a survey design, such as clustered, stratified, etc, generate summary statistics, and conduct sub-population analysis.
+The \proglang{DeclareDesign} package [@citeDeclareDesign] is a newer package that is based around a grammar of research design aspects including: data generation, creating treatment and control groups, and sampling.
+There is also an existing package called \proglang{MRP} available here: https://github.com/gelman/mrp.
+However, it does not appear to have been updated in some time.
 
-Survey package
-
-Declare design
-
-DIY
-
-MRP package - https://github.com/gelman/mrp
-
-And how we're different
+The main alternative approach is for MRP to be conducted on a case-specific basis.
+This means that a researcher who is interested in MRP estimates writes the code needed to obtain, clean, prepare, analyze and ultimately interpret the estimates.
+Again, while there is nothing inherently wrong with this approach, successfully conducting MRP requires dealing with a large number of small issues.
+Each of these is small in their own right but getting them wrong can have large effects that are potentially unnoticed on the estimates.
+It is also easy to overlook steps, or to not appropriately document them.
+The \proglang{MRPKit} package is prescriptive about the steps that must be taken, and ensures that all aspects are documented.
 
@@
 
-The \proglang{MRPKit} package has the following key components: survey objects, which for most users will be a collection of two surveys where one is larger than the other; a survey map, that relates the survey objects and then an MRP object, which is created once the survey map is applied to the survey objects.
-Regression acts on the MRP object.
-These objects are subject to the following verbs: 'add', 'delete', 'new', and 'replace'.
-For instance: \proglang{SurveyMap\$add}, \proglang{SurveyOb\$delete}, \proglang{SurveyMap\$new}, and \proglang{SurveyMap\$replace}.
+The \proglang{MRPKit} package has the following key components: \proglang{SurveyData} objects, which for most users will be a collection of two surveys where one is larger than the other; a \proglang{SurveyQuestion} object, that holds the mapping for one question or demographic between the survey and post-stratification dataset; and a \proglang{SurveyMap} object, which holds the mapping between a set of items in a survey and a post-stratification dataset.
+These are expanded on in the following sub-sections before being brought together.
 
-\subsection*{Survey objects}
+\subsection{SurveyData objects}
 
-Survey objects are the surveys that MRP will bring together.
-A user can \proglang{add} a survey object by providing a CSV file location.
+A \proglang{SurveyData} object represents a survey and its metadata. The survey itself is a data frame and importantly has a series of metadata. The survey metadata consists of the following:
+
+- per-column questions: a list of strings
+- per-column allowed response values: a list of character vectors
+- per-column survey weights: a vector of numeric weights
+- survey design: a string that specifies the survey design using \proglang{lme4} style formula syntax.
+
+Some examples of survey designs include:
+
+- `~.`: a random sample
+- `~ (1|cluster)`: a one stage cluster sample
+- `~ stratum`: a stratified sample
+
+These \proglang{SurveyData} objects are the surveys that MRP will bring together.
+A user can use the verb `add` to create a survey object by providing a CSV file location.
 The user then needs to identify the data types of each column.
@@
 
-New is when you have a new survey object and you are bringing it in. e.g.
+\subsection{SurveyQuestion objects}
 
-% sampl <- surveyStr$new(...)
-% pop <- surveyStr$new(...)
-% map <- surveyMap$new(sampl, pop, ...)
-% mrp1 <- MRP$new(sampl, pop, map)
+A \proglang{SurveyQuestion} object holds the mapping for one question or demographic between the survey and post-stratification dataset.
+For instance, if both the survey and post-stratification dataset contain a question about age, it may be that answers are expressed as a single integer representing years and so mapping them is relatively straight-forward.
+This is done with the verb `new`, for instance:
 
-% option 1
-% mod1 <- mrp1$fit(y ~ x + (1|group))
-% mod1 <- mrp1$fit(y ~ x + (1|group) + (1|group2))
+```{r, eval = FALSE}
+SurveyQuestion$new(
+  name = "age",
+  col_names = c("age1","age2"),
+  values_map = list(
+    "18" = "18", 
+    "19" = "19",
+    "20" = "20"
+    )
+)
+```
 
-% mod1$poststratify()
-% mod2$poststratify("model1")
+If ages have been aggregating into age-groups then it may not be the case that both the survey and the post-stratification dataset have groups that directly correspond.
+In that case, a user needs to specify how the age-groups fit together.
+This is also done with the verb `new`, for instance:
 
-% or
+```{r, eval = FALSE}
+SurveyQuestion$new(
+  name = "age",
+  col_names = c("age1","age2"),
+  values_map = list(
+    "18-25" = "18-35", 
+    "26-35" = "18-35",
+    "36-45" = "36-55",
+    "46-55" = "36-55",
+    "56-65" = "56-65",
+    "66-75" = "66+",
+    "76-90" = "66+"
+    )
+)
+```
 
-% option 2
-% mrp1$fit(y ~ x + (1|group), name = "model1")
-% mrp1$fit(y ~ x + (1|group) + (1|group2), name = "model2")
+In this example, the age-groups do not neatly match, however they can be made to match.
+For instance, the 36-45 and 46-55 age-groups in the survey correspond to the 36-55 age-group in the post-stratification dataset.
 
-% mrp1$poststartify("model1")
-% mrp1$poststartify("model2")
+If there is not a clean correspondence between the survey and post-stratification datasets, then the user needs to decide which age-group to assign to.
 
-What is the difference between add and new? What is replace doing? URGH, I just need to look at the repo.
+\subsection{SurveyMap objects}
 
-\subsection*{Survey map}
+A \proglang{SurveyMap} object holds the mapping between a set of items in a survey and a population dataset.
+That is, it draws on two \proglang{SurveyData} objects, and a \proglang{SurveyQuestion} object, and brings the two \proglang{SurveyData} objects together following the correspondences provided in the \proglang{SurveyQuestion} object.
 
-A `SurveyMap` object holds the mapping between a set of items in a survey and a population dataset.
 The label is the item label in each dataset and the values is a list of all possible values.
-The values for the survey and population must be aligned, i.e., the lists must have the same number of elements and the values at index i in each list are equivalent.
+The values for the survey and post-stratification must be aligned, i.e., the lists must have the same number of elements and the values at index i in each list are equivalent.
 If there is a meaningful ordering over the values, they should be listed in that order, either descending or ascending.
 
-One of the fundamental issues when conducting MRP is to ensure that
+One of the fundamental issues when conducting MRP is to ensure that the survey and post-stratification datasets are aligned. This object ensures that is the case before a user conducts their analysis.
 
-\subsection*{MRP object}
+For instance, if a user decides to use age-group and gender, then the map would specify the correspondences in age-group and also in gender.
+If they then also decided to use education, then the map would also specify that correspondence.
+They would then use the verb `add` to include that third question in the map.
 
-An  MRP object contains survey objects, and a survey map. At the point at which the survey objects are put into the MRP object they become immutable but the survey map object is.
+The verb `validate` is used on \proglang{SurveyMap} objects to assess whether the proposed map is appropriate.
+This means that an error occurs if there is a mismatch, and a warning occurs if there are variables that are available in either the survey or post-stratification datasets that have not been mapped and hence would be unavailable to model.
 
-The MRP object outputs an analysis. There would be a MRP model fit class that would
+After the \proglang{SurveyMap} object has been validated then the verb `mapping` is used to prepare it to be modelled.
 
-Many objects on the same class. There are two
+A model is fit to a \proglang{SurveyMap} object using the verb `fit`.
+After a model has been fit it can be used to get MRP-based estimates using `predictify()`.
+And finally the estimates can be visualized using `visify()`.
+
+\subsection{Putting them together}
+
+We provide examples of implementation in Section \ref{sec:vignette}, however, we want to explain putting the three objects together here and to briefly bring the three objects together.
+
+To begin with, one needs two datasets. These would be created using `SurveyObj$new`, specifying the questions, responses, weights, and design.
+
+Having established both a survey dataset and a post-stratification dataset, a user need to specify how the questions in the survey dataset correspond to the questions in the post-stratification dataset.
+This is done on a question-by-question basis using `question$new` assigning  a name for that question, the labels that are used in the survey and post-stratification datasets, and the values.
+A user does this in such a way that the values are matched between the two datasets.
+
+The user now need to put these questions together and does this using `SurveyMap$new` and specifying the survey object, the post-stratification object, and the questions are to be used.
+A user can validate this map with `validate`, and potentially add additional questions using `add`, delete some existing ones using `delete`, or potentially replace a question using `replace`.
+Once a user is happy with the correspondence to be used they can establish a mapping using `mapping`.
+
+The relationship can then be modelled using `fit`.
+There are a variety of options that are available as built-into the package, including `rstanarm::stan_glmer`, and `brms::brm`.
+A user could also specify their own modelling approach if they wanted to.
+In any case, once a model has been fit to the survey dataset, it can be used to create estimates based on the post-stratification dataset with `predictify`, which takes a fitted model as an argument.
+Typically a user is interested in some type of high-level estimates, for instance on a province/state basis, and `collapsify` provides the functionality to do this.
+Finally, the estimates on that collapsed basis can be visualized with `visify`
 
@@
 \section{Vignette} \label{sec:vignette}
+
+\subsection*{Simulated data}
 

### 2021-03-10 paper/paper_joss/paper_joss.tex

@@
 University \And Jonah Gabry\\Columbia University \And Rohan
-Alexander\\University of Toronto
+Alexander\\University of Toronto \And Dewi Amaliah\\Monash University
 }
@@
 
-\Plainauthor{Lauren Kennedy, Mitzi Morris, Jonah Gabry, Rohan Alexander}
+\Plainauthor{Lauren Kennedy, Mitzi Morris, Jonah Gabry, Rohan
+Alexander, Dewi Amaliah}
 \Plaintitle{A `grammar' of multilevel regression with
@@
 post-stratification (MRP) and implement this grammar in the R Package
-\pkg{MRPKit}. Our grammar is centered around the following verbs: add,
-remove, delete, and replace. These act on survey objects. Our grammar,
-and its implementation, provides a detailed workflow when conducting MRP
-that will be useful for researchers and in teaching.
+\pkg{MRPKit}. Our grammar is centered around the following verbs: new,
+validate, add, delete, replace, mapping, tabulate, fit, predictify,
+collapsify, visify. These act on survey objects. Our grammar, and its
+implementation, provides a detailed workflow when conducting MRP that
+will be useful for researchers and in teaching.
 }
@@
 \Keywords{multilevel regression with
-post-stratification, \proglang{R}, reproducibility, statistics, political
-science}
+post-stratification, \proglang{R}, reproducibility, statistics}
 \Plainkeywords{multilevel regression with
-post-stratification, R, reproducibility, statistics, political science}
+post-stratification, R, reproducibility, statistics}
 
@@
 approach in which surveys are related to each other using a statistical
-model. This is important because known biases in one survey, can be
+model. This is important because known biases in one survey can be
 adjusted for by another in a statistically reasonable way. This enables
 better use of non-representative surveys, additional information, and
-propagation of uncertainty. However it can be difficult to use MRP due
-to this need to related two different datasets. This package defines a
-grammar, or list of underlying rules, of MRP and then describes an R
-package, \proglang{MRPKit}, that implements this grammar.
+propagation of uncertainty. However, it can be difficult to use MRP due
+to this need to relate (at least) two different datasets. This package
+defines a grammar, or list of underlying rules, for MRP and then
+describes an R package, \proglang{MRPKit}, that implements this grammar.
 
 At its core, MRP is a mapping between a survey object and a population
-object. It is from this mapping that the power of MRP exists, but it
-also establishing this mapping that is the difficult part of
-implementing MRP models. Making this implementation easier and more
-reproducible is important as interest in, and the use of, MRP increases.
-It can be difficult even for those experienced with MRP to ensure there
-are no mistakes in this mapping and easing this is an important
-contribution to enhancing the reproducibility of MRP analysis.
+object. It is from this mapping that the power of MRP exists, but it can
+be difficult to establish this mapping in a reproducible way that fits
+into a typical applied statistics workflow. Making this implementation
+easier and more reproducible is important as interest in, and the use
+of, MRP increases. It can be difficult even for those experienced with
+MRP to ensure there are no mistakes in this mapping and creating a
+well-defined workflow for MRP is an important contribution to enhancing
+the reproducibility of MRP analysis.
 
@@
 rules and principles that are common to every analysis based on MRP.
-This grammar is based around the following verbs: `add', `remove',
-`delete', and `replace'. These verbs are applied to a survey object, a
-post-stratification object, and survey\_map object. These three objects
-come together to create a mapping object, which is what processes such
-as regression act on. Finally, common diagnostics and graphs are
-implemented. In this way, our grammar and package implement an entire
-statistical workflow for conducting MRP.
+This grammar is based around the following verbs: `new', `validate',
+`add', `delete', `replace', `mapping', `tabulate', `fit', `predictify',
+`collapsify', and `visify'. These verbs are applied to a survey object,
+a post-stratification object, and survey-map object. These three objects
+come together to create a mapping object, which is what estimation
+processes such as regression act on. Finally, common diagnostics and
+graphs are implemented. In this way, our grammar and package implement
+an entire statistical workflow for MRP.
 
@@
 post-stratification object would typically be a larger survey, such as,
-in the case of the US, the (INSERT THE USUAL ONE), or a census.
+in the case of the US, the American Community Surveys, or a census.
 
@@
 implementing, and simulating from, surveys. Instead, ours is focused on
-what is needed for MRP.
+what is needed for MRP, and we draw on their packages where possible.
 
@@
 
-\section{Review of the other packages} \label{sec:review}
+\section{Review of related packages} \label{sec:review}
 
 The most common alternative at the moment to this package is for users
-to do all aspects themselves. While there is nothing inherently
-difficult about MRP, the implementation can be difficult. In particular,
-preparing and matching different levels between surveys can be time
-consuming and potentially introduce undocumented errors.
+to do all aspects of the MRP workflow themselves. While there is nothing
+inherently difficult about MRP, the implementation can be difficult. In
+particular, preparing and matching different levels between surveys can
+be time consuming and potentially introduce undocumented errors.
 
 The \proglang{survey} package \citep[\citet{lumleytwo},
-\citet{lumleythree}]{lumleyone} \ldots{}
-
-The \proglang{DeclareDesign} package \citep{citeDeclareDesign}\ldots{}
-
-There is also an existing package called \proglang{MRP} available here:
-\url{https://github.com/gelman/mrp}. However it does not appear to have
-been updated in some time.
+\citet{lumleythree}]{lumleyone} is well-established and allows a user to
+specify a survey design, such as clustered, stratified, etc, generate
+summary statistics, and conduct sub-population analysis. The
+\proglang{DeclareDesign} package \citep{citeDeclareDesign} is a newer
+package that is based around a grammar of research design aspects
+including: data generation, creating treatment and control groups, and
+sampling. There is also an existing package called \proglang{MRP}
+available here: \url{https://github.com/gelman/mrp}. However, it does
+not appear to have been updated in some time.
 
 The main alternative approach is for MRP to be conducted on a
-case-specific basis. That is a researcher interested in MRP estimates,
-writes the code needed to obtain, clean, prepare, analyze and ultimately
-interpret the estimates. Again, while there is nothing inherently wrong
-with this approach, successfully conducting MRP requires dealing with a
-large number of small issues. Each of these is small in their own right,
-but getting them wrong can have large effects that are potentially
-unnoticed on the estimates.
+case-specific basis. This means that a researcher who is interested in
+MRP estimates writes the code needed to obtain, clean, prepare, analyze
+and ultimately interpret the estimates. Again, while there is nothing
+inherently wrong with this approach, successfully conducting MRP
+requires dealing with a large number of small issues. Each of these is
+small in their own right but getting them wrong can have large effects
+that are potentially unnoticed on the estimates. It is also easy to
+overlook steps, or to not appropriately document them. The
+\proglang{MRPKit} package is prescriptive about the steps that must be
+taken, and ensures that all aspects are documented.
 
@@
 
-The \proglang{MRPKit} package has the following key components: survey
-objects, which for most users will be a collection of two surveys where
-one is larger than the other; a survey map, that relates the survey
-objects and then an MRP object, which is created once the survey map is
-applied to the survey objects. Regression acts on the MRP object. These
-objects are subject to the following verbs: `add', `delete', `new', and
-`replace'. For instance: \proglang{SurveyMap\$add},
-\proglang{SurveyOb\$delete}, \proglang{SurveyMap\$new}, and
-\proglang{SurveyMap\$replace}.
+The \proglang{MRPKit} package has the following key components:
+\proglang{SurveyData} objects, which for most users will be a collection
+of two surveys where one is larger than the other; a
+\proglang{SurveyQuestion} object, that holds the mapping for one
+question or demographic between the survey and post-stratification
+dataset; and a \proglang{SurveyMap} object, which holds the mapping
+between a set of items in a survey and a post-stratification dataset.
+These are expanded on in the following sub-sections before being brought
+together.
 
-\subsection*{Survey objects}
+\subsection{SurveyData objects}
 
-Survey objects are the surveys that MRP will bring together. A user can
-\proglang{add} a survey object by providing a CSV file location. The
-user then needs to identify the data types of each column. Typically,
-there will need to be two survey objects, where one will be the survey
-that is of interest for its response variables, such as political
-opinion, and another will be a survey that is to be used for
-post-stratification.
+A \proglang{SurveyData} object represents a survey and its metadata. The
+survey itself is a data frame and importantly has a series of metadata.
+The survey metadata consists of the following:
 
-New is when you have a new survey object and you are bringing it in.
-e.g.
+\begin{itemize}
+\tightlist
+\item
+  per-column questions: a list of strings
+\item
+  per-column allowed response values: a list of character vectors
+\item
+  per-column survey weights: a vector of numeric weights
+\item
+  survey design: a string that specifies the survey design using
+  \proglang{lme4} style formula syntax.
+\end{itemize}
 
-What is the difference between add and new? What is replace doing? URGH,
-I just need to look at the repo.
+Some examples of survey designs include:
 
-\subsection*{Survey map}
+\begin{itemize}
+\tightlist
+\item
+  \texttt{\textasciitilde{}.}: a random sample
+\item
+  \texttt{\textasciitilde{}\ (1\textbar{}cluster)}: a one stage cluster
+  sample
+\item
+  \texttt{\textasciitilde{}\ stratum}: a stratified sample
+\end{itemize}
 
-A \texttt{SurveyMap} object holds the mapping between a set of items in
-a survey and a population dataset. The label is the item label in each
-dataset and the values is a list of all possible values. The values for
-the survey and population must be aligned, i.e., the lists must have the
-same number of elements and the values at index i in each list are
-equivalent. If there is a meaningful ordering over the values, they
-should be listed in that order, either descending or ascending.
+These \proglang{SurveyData} objects are the surveys that MRP will bring
+together. A user can use the verb \texttt{add} to create a survey object
+by providing a CSV file location. The user then needs to identify the
+data types of each column. Typically, there will need to be two survey
+objects, where one will be the survey that is of interest for its
+response variables, such as political opinion, and another will be a
+survey that is to be used for post-stratification.
 
-One of the fundamental issues when conducting MRP is to ensure that
+\subsection{SurveyQuestion objects}
 
-\subsection*{MRP object}
+A \proglang{SurveyQuestion} object holds the mapping for one question or
+demographic between the survey and post-stratification dataset. For
+instance, if both the survey and post-stratification dataset contain a
+question about age, it may be that answers are expressed as a single
+integer representing years and so mapping them is relatively
+straight-forward. This is done with the verb \texttt{new}, for instance:
 
-An MRP object contains survey objects, and a survey map. At the point at
-which the survey objects are put into the MRP object they become
-immutable but the survey map object is.
+\begin{CodeChunk}
+\begin{CodeInput}
+R> SurveyQuestion$new(
++   name = "age",
++   col_names = c("age1","age2"),
++   values_map = list(
++     "18" = "18", 
++     "19" = "19",
++     "20" = "20"
++     )
++ )
+\end{CodeInput}
+\end{CodeChunk}
 
-The MRP object outputs an analysis. There would be a MRP model fit class
-that would
+If ages have been aggregating into age-groups then it may not be the
+case that both the survey and the post-stratification dataset have
+groups that directly correspond. In that case, a user needs to specify
+how the age-groups fit together. This is also done with the verb
+\texttt{new}, for instance:
 
-Many objects on the same class. There are two
+\begin{CodeChunk}
+\begin{CodeInput}
+R> SurveyQuestion$new(
++   name = "age",
++   col_names = c("age1","age2"),
++   values_map = list(
++     "18-25" = "18-35", 
++     "26-35" = "18-35",
++     "36-45" = "36-55",
++     "46-55" = "36-55",
++     "56-65" = "56-65",
++     "66-75" = "66+",
++     "76-90" = "66+"
++     )
++ )
+\end{CodeInput}
+\end{CodeChunk}
+
+In this example, the age-groups do not neatly match, however they can be
+made to match. For instance, the 36-45 and 46-55 age-groups in the
+survey correspond to the 36-55 age-group in the post-stratification
+dataset.
+
+If there is not a clean correspondence between the survey and
+post-stratification datasets, then the user needs to decide which
+age-group to assign to.
+
+\subsection{SurveyMap objects}
+
+A \proglang{SurveyMap} object holds the mapping between a set of items
+in a survey and a population dataset. That is, it draws on two
+\proglang{SurveyData} objects, and a \proglang{SurveyQuestion} object,
+and brings the two \proglang{SurveyData} objects together following the
+correspondences provided in the \proglang{SurveyQuestion} object.
+
+The label is the item label in each dataset and the values is a list of
+all possible values. The values for the survey and post-stratification
+must be aligned, i.e., the lists must have the same number of elements
+and the values at index i in each list are equivalent. If there is a
+meaningful ordering over the values, they should be listed in that
+order, either descending or ascending.
+
+One of the fundamental issues when conducting MRP is to ensure that the
+survey and post-stratification datasets are aligned. This object ensures
+that is the case before a user conducts their analysis.
+
+For instance, if a user decides to use age-group and gender, then the
+map would specify the correspondences in age-group and also in gender.
+If they then also decided to use education, then the map would also
+specify that correspondence. They would then use the verb \texttt{add}
+to include that third question in the map.
+
+The verb \texttt{validate} is used on \proglang{SurveyMap} objects to
+assess whether the proposed map is appropriate. This means that an error
+occurs if there is a mismatch, and a warning occurs if there are
+variables that are available in either the survey or post-stratification
+datasets that have not been mapped and hence would be unavailable to
+model.
+
+After the \proglang{SurveyMap} object has been validated then the verb
+\texttt{mapping} is used to prepare it to be modelled.
+
+A model is fit to a \proglang{SurveyMap} object using the verb
+\texttt{fit}. After a model has been fit it can be used to get MRP-based
+estimates using \texttt{predictify()}. And finally the estimates can be
+visualized using \texttt{visify()}.
+
+\subsection{Putting them together}
+
+We provide examples of implementation in Section \ref{sec:vignette},
+however, we want to explain putting the three objects together here and
+to briefly bring the three objects together.
+
+To begin with, one needs two datasets. These would be created using
+\texttt{SurveyObj\$new}, specifying the questions, responses, weights,
+and design.
+
+Having established both a survey dataset and a post-stratification
+dataset, a user need to specify how the questions in the survey dataset
+correspond to the questions in the post-stratification dataset. This is
+done on a question-by-question basis using \texttt{question\$new}
+assigning a name for that question, the labels that are used in the
+survey and post-stratification datasets, and the values. A user does
+this in such a way that the values are matched between the two datasets.
+
+The user now need to put these questions together and does this using
+\texttt{SurveyMap\$new} and specifying the survey object, the
+post-stratification object, and the questions are to be used. A user can
+validate this map with \texttt{validate}, and potentially add additional
+questions using \texttt{add}, delete some existing ones using
+\texttt{delete}, or potentially replace a question using
+\texttt{replace}. Once a user is happy with the correspondence to be
+used they can establish a mapping using \texttt{mapping}.
+
+The relationship can then be modelled using \texttt{fit}. There are a
+variety of options that are available as built-into the package,
+including \texttt{rstanarm::stan\_glmer}, and \texttt{brms::brm}. A user
+could also specify their own modelling approach if they wanted to. In
+any case, once a model has been fit to the survey dataset, it can be
+used to create estimates based on the post-stratification dataset with
+\texttt{predictify}, which takes a fitted model as an argument.
+Typically a user is interested in some type of high-level estimates, for
+instance on a province/state basis, and \texttt{collapsify} provides the
+functionality to do this. Finally, the estimates on that collapsed basis
+can be visualized with \texttt{visify}
 

### 2021-03-15 R/sim_data_approx_voters_popn.R
Add documentation to the example data.

@@ -1,0 +1,21 @@
+#' A Simulated Data of NLP Voters Approximate Population (Modelled after ACS)
+#'
+#' This is a generated data shows the approximate population of the
+#' Neverland Labor Party (NLP) Survey to demonstrate the usage of the package.
+#'
+#' @source Created in NeverlandVoters.R
+#'
+#' @format A data frame with columns:
+#'
+#' \describe{
+#'  \item{age2}{Age group}
+#'  \item{gender}{Gender}
+#'  \item{vote_for}{The voted party in the 2016 election}
+#'  \item{y}{Preference to NLP}
+#'  \item{wt}{Raked and trimmed survey weight}
+#' }
+#' @examples
+#'
+#'  approx_voters_popn
+#'
+"approx_voters_popn"

### 2021-03-15 R/sim_data_nlp_survey.R

@@ -1,0 +1,22 @@
+#' A Simulated Data of NLP Voters Survey
+#'
+#' This is a generated data shows the survey result of the preference
+#' to Neverland Labor Party (NLP)
+#' along with several demographic variables to demonstrate the usage of the package.
+#'
+#' @source Created in NeverlandVoters.R
+#'
+#' @format A data frame with columns:
+#'
+#' \describe{
+#'  \item{age1}{Age group}
+#'  \item{gender}{Gender}
+#'  \item{vote_for}{The party that the respondent vote for in the 2016 election}
+#'  \item{y}{Preference to NLP}
+#'  \item{wt}{Raked and trimmed survey weight}
+#' }
+#' @examples
+#'
+#'  nlp_survey
+#'
+"nlp_survey"

### 2021-03-16 paper/paper.Rmd

@@
 ---
-documentclass: jss
+title: "MRP-Kit: A 'grammar' of multilevel regression with post-stratification and implementation in R"
 author:
-  - name: Lauren Kennedy
-    affiliation: Monash University
-    email: \email{name@company.com}
-    url: http://rstudio.com
-  - name: Mitzi Morris
-    affiliation: Columbia University
-  - name: Jonah Gabry
-    affiliation: Columbia University
-  - name: Rohan Alexander
-    affiliation: University of Toronto
-title:
-  formatted: "\\pkg{MRP-Kit} A 'grammar' of multilevel regression with post-stratification and implementation in \\proglang{R}"
-  # If you use tex in the formatted title, also supply version without
-  plain:     "A 'grammar' of multilevel regression with post-stratification and implementation in R"
-  # For running headers, if needed
-  short:     "\\pkg{foo}: A grammar of MRP"
-abstract: >
-  In this paper we define a grammar for multilevel regression with post-stratification (MRP) and implement this grammar in the R Package \pkg{MRPKit}.
-Our grammar is centered around the following verbs: add, remove, delete, and replace.
-These act on survey objects.
+  - Lauren Kennedy^[Monash University, lauren.kennedy729@gmail.com]
+  - Mitzi Morris^[Columbia University]
+  - Jonah Gabry^[Columbia University]
+  - Rohan Alexander^[University of Toronto]
+  - Dewi Amaliah^[Monash University]
+thanks: "Code and data are available at: https://github.com/mitzimorris/mrp-kit."
+date: "`r format(Sys.time(), '%d %B %Y')`"
+abstract: |
+    In this paper we define a 'grammar' for multilevel regression with post-stratification (MRP) and implement this grammar in the R Package `MRPKit`.
+Our grammar is centered around the following components: datasets; correspondences; and models.
+Survey datasets are linked by correspondences and models act on these correspondences.
 Our grammar, and its implementation, provides a detailed workflow when conducting MRP that will be useful for researchers and in teaching.
-keywords:
-  # at least one keyword must be supplied
-  formatted: [multilevel regression with post-stratification, "\\proglang{R}", reproducibility, statistics, political science]
-  plain:     [multilevel regression with post-stratification, R, reproducibility, statistics, political science]
-preamble: >
-  \usepackage{amsmath}
-output: rticles::jss_article
+
+    Keywords: multilevel regression with post-stratification, R, reproducibility, statistics.
+output:
+  bookdown::pdf_document2:
+toc: FALSE
+bibliography: references.bib
 ---
 
-```{r, setup, include=FALSE}
-options(prompt = 'R> ', continue = '+ ')
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
 ```
 
-\section[Introduction]{Introduction} \label{sec:intro}
+# Introduction
 
 Multilevel regression with post-stratification (MRP) is a statistical approach in which surveys are related to each other using a statistical model.
-This is important because known biases in one survey, can be adjusted for by another in a statistically reasonable way.
+This is important because known biases in one survey can be adjusted by another in a statistically reasonable way.
 This enables better use of non-representative surveys, additional information, and propagation of uncertainty.
-However it can be difficult to use MRP due to this need to related two different datasets.
-This package defines a grammar, or list of underlying rules, of MRP and then describes an R package, \proglang{MRPKit}, that implements this grammar.
+However, it can be difficult to use MRP due to this need to relate (at least) two different datasets.
+This package defines a grammar, or list of underlying rules, for MRP and then describes an R package, `MRPKit`, that implements this grammar.
 
 At its core, MRP is a mapping between a survey object and a population object.
-It is from this mapping that the power of MRP exists, but it also establishing this mapping that is the difficult part of implementing MRP models.
+It is from this mapping that the power of MRP exists, but it can be difficult to establish this mapping in a reproducible way that fits into a typical applied statistics workflow.
 Making this implementation easier and more reproducible is important as interest in, and the use of, MRP increases.
-It can be difficult even for those experienced with MRP to ensure there are no mistakes in this mapping and easing this is an important contribution to enhancing the reproducibility of MRP analysis.
+It can be difficult even for those experienced with MRP to ensure there are no mistakes in this mapping and creating a well-defined workflow for MRP is an important contribution to enhancing the reproducibility of MRP analysis.
 
 We first define a grammar of MRP, which we define as the underlying rules and principles that are common to every analysis based on MRP.
-This grammar is based around the following verbs: 'add', 'remove', 'delete', and 'replace'.
-These verbs are applied to a survey object, a post-stratification object, and survey\_map object.
-These three objects come together to create a mapping object, which is what processes such as regression act on.
-Finally, common diagnostics and graphs are implemented.
-In this way, our grammar and package implement an entire statistical workflow for conducting MRP.
+This grammar is based around the following components: 'datasets'; 'correspondences'; and 'models'.
+Typically MRP requires at least two datasets - usually a survey object and a post-stratification object, which we implement using the `SurveyData` method.
+The questions in each datasets are then aligned by establishing a correspondence between each possible answer to a particular question, and we implement this using the `SurveyQuestion` method.
+Having linked each question, the surveys themselves can be  linked and we implement this using the `SurveyMap` method.
+Finally, that correspondence can be modelled, and used to generate estimates.
+In this way, our grammar and package implement an entire statistical workflow for MRP that is largely agnostic of the specifics of each component.
 
-The survey object would typically be a regular survey, such as a political poll of 1,000 respondents, but it could also be a larger survey, such as the Canadian Election Survey, or similar.
-The post-stratification object would typically be a larger survey, such as, in the case of the US, the (INSERT THE USUAL ONE), or a census.
+The survey object would typically be a regular survey, such as a political poll of, say, 1,000 respondents, but it could also be a larger survey, such as the Canadian Election Survey, or similar.
+The post-stratification object would typically be a larger survey, such as, in the case of the US, the American Community Surveys, or a census.
+The correspondences need to be established for each possible answer for each question, for instance, the age-groups used in both surveys need to be homogenized.
+Finally, the model would typically be a Bayesian hierarchical model, but in principle any variety could be used.
 
-Our grammar and package complements existing packages such as \proglang{survey} (ADD CITATION), X and Y. These packages... Instead, ours is focused on what is needed for MRP.
+Our grammar and package complements existing packages such as `survey` [@lumleyone, @lumleytwo, @lumleythree], and `DeclareDesign` [@citeDeclareDesign].
+These packages are focused on the designing, implementing, and simulating from, surveys.
+Instead, ours is focused on what is needed for MRP, and we draw on their packages where possible.
 
-The remainder of this paper is structured as follows: Section \ref{sec:review} reviews similar packages and contributions and places ours within that context.
-Section \ref{sec:components} discusses the grammar and the core aspects of MRP as implemented in \proglang{MRPKit}.
-Section \ref{sec:implementation} discusses some of the implementation issues and technical notes related to the decisions that were made.
-Section \ref{sec:vignette} provides two examples of the package in use, one using SOMETHING, and the other using SOMETHING ELSE.
-Finally, Section \ref{sec:summary} provides a summary discussion, some cautions and weaknesses, as well as notes about next steps.
+The remainder of this paper is structured as follows: Section \@ref(review) reviews similar packages and contributions and places ours within that context.
+Section \@ref(components) discusses the grammar and the core aspects of MRP as implemented in `MRPKit`.
+Section \@ref(implementation) discusses some of the implementation issues and technical notes related to the decisions that were made.
+Section \@ref(vignette) provides two examples of the package in use, one using SOMETHING, and the other using SOMETHING ELSE.
+Finally, Section \@ref(summary) provides a summary discussion, some cautions and weaknesses, as well as notes about next steps.
 
-\section{Review of the other packages} \label{sec:review}
+# Review of related packages {#review}
 
-The most common alternative at the moment to this package is for users to do all aspects themselves.
-While there is nothing conceptually difficult about MRP, the implementation can be difficult.
-In particular, preparing and matching different levels between surveys can be time consuming and potentially introduce errors.
+The most common alternative at the moment to this package is for users to do all aspects of the MRP workflow themselves.
+While there is nothing inherently difficult about MRP, the implementation can be difficult.
+In particular, preparing and matching different levels between surveys can be time consuming and potentially introduce undocumented errors.
 
-HERE
+The `survey` package [@lumleyone, @lumleytwo, @lumleythree] is well-established and allows a user to specify a survey design, such as clustered, stratified, etc, generate summary statistics, and conduct sub-population analysis.
+The `DeclareDesign` package [@citeDeclareDesign] is a newer package that is based around a grammar of research design aspects including: data generation, creating treatment and control groups, and sampling.
+There is also an existing package called `MRP` available here: https://github.com/gelman/mrp.
+However, it does not appear to have been updated in some time.
 
-Survey package
+The main alternative approach is for MRP to be conducted on a case-specific basis.
+This means that a researcher who is interested in MRP estimates writes the code needed to obtain, clean, prepare, analyze and ultimately interpret the estimates.
+Again, while there is nothing inherently wrong with this approach, successfully conducting MRP requires dealing with a large number of small issues.
+Each of these is small in their own right but getting them wrong can have large effects that are potentially unnoticed on the estimates.
+It is also easy to overlook steps, or to not appropriately document them.
+The `MRPKit` package is prescriptive about the steps that must be taken, and ensures that all aspects are documented.
 
-Declare design
+# Components and grammar {#components}
 
-DIY
+The `MRPKit` package has the following key components: `SurveyData` objects, which for most users will be a collection of two surveys where one is larger than the other; a `SurveyQuestion` object, that holds the mapping for one question or demographic between the survey and post-stratification dataset; and a `SurveyMap` object, which holds the mapping between a set of items in a survey and a post-stratification dataset.
+These are expanded on in the following sub-sections before being brought together.
 
-MRP package - https://github.com/gelman/mrp
+## SurveyData objects
 
-And how we're different
+A `SurveyData` object represents a survey and its metadata. The survey itself is a data frame and importantly has a series of metadata. The survey metadata consists of the following:
 
-\section{Components and grammar} \label{sec:components}
+- per-column questions: a list of strings
+- per-column allowed response values: a list of character vectors
+- per-column survey weights: a vector of numeric weights
+- survey design: a string that specifies the survey design using `lme4` style formula syntax.
 
-The \proglang{MRPKit} package has the following key components: survey objects, which for most users will be a collection of two surveys where one is larger than the other; a survey map, that relates the survey objects and then an MRP object, which is created once the survey map is applied to the survey objects.
-Regression acts on the MRP object.
-These objects are subject to the following verbs: 'add', 'delete', 'new', and 'replace'.
-For instance: \proglang{SurveyMap\$add}, \proglang{SurveyOb\$delete}, \proglang{SurveyMap\$new}, and \proglang{SurveyMap\$replace}.
+Some examples of survey designs include:
 
-\subsection*{Survey objects}
+- `~.`: a random sample
+- `~ (1|cluster)`: a one stage cluster sample
+- `~ stratum`: a stratified sample
 
-Survey objects are the surveys that MRP will bring together.
-A user can \proglang{add} a survey object by providing a CSV file location.
+These `SurveyData` objects are the surveys that MRP will bring together.
+A user can use the function `add` to create a survey object by providing a CSV file location.
 The user then needs to identify the data types of each column.
@@
 
-New is when you have a new survey object and you are bringing it in. e.g.
+## SurveyQuestion objects
 
-% sampl <- surveyStr$new(...)
-% pop <- surveyStr$new(...)
-% map <- surveyMap$new(sampl, pop, ...)
-% mrp1 <- MRP$new(sampl, pop, map)
+A `SurveyQuestion` object holds the mapping for one question or demographic between the survey and post-stratification dataset.
+For instance, if both the survey and post-stratification dataset contain a question about age, it may be that answers are expressed as a single integer representing years and so mapping them is relatively straight-forward.
+This is done with the function `new`, for instance:
 
-% option 1
-% mod1 <- mrp1$fit(y ~ x + (1|group))
-% mod1 <- mrp1$fit(y ~ x + (1|group) + (1|group2))
+```{r, eval = FALSE}
+SurveyQuestion$new(
+  name = "age",
+  col_names = c("age1","age2"),
+  values_map = list(
+    "18" = "18", 
+    "19" = "19",
+    "20" = "20"
+    )
+)
+```
 
-% mod1$poststratify()
-% mod2$poststratify("model1")
+If ages have been aggregating into age-groups then it may not be the case that both the survey and the post-stratification dataset have groups that directly correspond.
+In that case, a user needs to specify how the age-groups fit together.
+This is also done with the function `new`, for instance:
 
-% or
+```{r, eval = FALSE}
+SurveyQuestion$new(
+  name = "age",
+  col_names = c("age1","age2"),
+  values_map = list(
+    "18-25" = "18-35", 
+    "26-35" = "18-35",
+    "36-45" = "36-55",
+    "46-55" = "36-55",
+    "56-65" = "56-65",
+    "66-75" = "66+",
+    "76-90" = "66+"
+    )
+)
+```
 
-% option 2
-% mrp1$fit(y ~ x + (1|group), name = "model1")
-% mrp1$fit(y ~ x + (1|group) + (1|group2), name = "model2")
+In this example, the age-groups do not neatly match, however they can be made to match.
+For instance, the 36-45 and 46-55 age-groups in the survey correspond to the 36-55 age-group in the post-stratification dataset.
 
-% mrp1$poststartify("model1")
-% mrp1$poststartify("model2")
+If there is not a clean correspondence between the survey and post-stratification datasets, then the user needs to decide which age-group to assign to.
 
-What is the difference between add and new? What is replace doing? URGH, I just need to look at the repo.
-
-\subsection*{Survey map}
+## SurveyMap objects
 
 A `SurveyMap` object holds the mapping between a set of items in a survey and a population dataset.
+That is, it draws on two `SurveyData` objects, and a `SurveyQuestion` object, and brings the two `SurveyData` objects together following the correspondences provided in the `SurveyQuestion` object.
+
 The label is the item label in each dataset and the values is a list of all possible values.
-The values for the survey and population must be aligned, i.e., the lists must have the same number of elements and the values at index i in each list are equivalent.
+The values for the survey and post-stratification must be aligned, i.e., the lists must have the same number of elements and the values at index i in each list are equivalent.
 If there is a meaningful ordering over the values, they should be listed in that order, either descending or ascending.
 
-One of the fundamental issues when conducting MRP is to ensure that
+One of the fundamental issues when conducting MRP is to ensure that the survey and post-stratification datasets are aligned. This object ensures that is the case before a user conducts their analysis.
 
-\subsection*{MRP object}
+For instance, if a user decides to use age-group and gender, then the map would specify the correspondences in age-group and also in gender.
+If they then also decided to use education, then the map would also specify that correspondence.
+They would then use the function `add` to include that third question in the map.
 
-An  MRP object contains survey objects, and a survey map. At the point at which the survey objects are put into the MRP object they become immutable but the survey map object is.
+The function `validate` is used on `SurveyMap` objects to assess whether the proposed map is appropriate.
+This means that an error occurs if there is a mismatch, and a warning occurs if there are variables that are available in either the survey or post-stratification datasets that have not been mapped and hence would be unavailable to model.
 
-The MRP object outputs an analysis. There would be a MRP model fit class that would
+After the `SurveyMap` object has been validated then the function `mapping` is used to prepare it to be modelled.
 
-Many objects on the same class. There are two
+A model is fit to a `SurveyMap` object using the function `fit`.
+After a model has been fit it can be used to get MRP-based estimates using `predictify()`.
+And finally the estimates can be visualized using `visify()`.
 
-\section{Implementation and technical notes} \label{sec:implementation}
+## Putting them together
 
-\section{Vignette} \label{sec:vignette}
+We provide examples of implementation in Section \@ref(vignette) , however, we want to explain putting the three objects together here and to briefly bring the three objects together.
 
-\subsection*{Using CCES data}
+To begin with, one needs two datasets. These would be created using `SurveyObj$new`, specifying the questions, responses, weights, and design.
 
-\subsection*{Using your own data}
+Having established both a survey dataset and a post-stratification dataset, a user need to specify how the questions in the survey dataset correspond to the questions in the post-stratification dataset.
+This is done on a question-by-question basis using `question$new` assigning  a name for that question, the labels that are used in the survey and post-stratification datasets, and the values.
+A user does this in such a way that the values are matched between the two datasets.
 
-\section{Summary and discussion} \label{sec:summary}
+The user now need to put these questions together and does this using `SurveyMap$new` and specifying the survey object, the post-stratification object, and the questions are to be used.
+A user can validate this map with `validate`, and potentially add additional questions using `add`, delete some existing ones using `delete`, or potentially replace a question using `replace`.
+Once a user is happy with the correspondence to be used they can establish a mapping using `mapping`.
 
-\subsection*{Next steps and cautions}
+The relationship can then be modelled using `fit`.
+There are a variety of options that are available as built-into the package, including `rstanarm::stan_glmer`, and `brms::brm`.
+A user could also specify their own modelling approach if they wanted to.
+In any case, once a model has been fit to the survey dataset, it can be used to create estimates based on the post-stratification dataset with `predictify`, which takes a fitted model as an argument.
+Typically a user is interested in some type of high-level estimates, for instance on a province/state basis, and `collapsify` provides the functionality to do this.
+Finally, the estimates on that collapsed basis can be visualized with `visify`
 
-\section*{Acknowledgments}
+# Implementation and technical notes {#implementation}
+
+# Vignette
+
+## Simulated data
+
+## Using CCES data
+
+## Using your own data
+
+# Summary and discussion summary {#summary}
+
+## On the importance of scaffolding for workflows
+
+## Next steps and cautions
+
+\newpage
+
+# References

### 2021-03-16 R/SurveyData.R
Added a way to pass survey design in survey package structure

@@ -9,16 +9,10 @@
 #'  - per-column questions: a list of strings
 #'  - per-column allowed response values: a list of character vectors
 #'  - per-column survey weights: a vector of numeric weights
-#'  - survey design: a string that specifies the survey design using \pkg{lme4}
-#'  style formula syntax.
-#'
-#' Examples of survey designs include:
-#'   - `~.`: a random sample
-#'   - `~ (1|cluster)`: a one stage cluster sample
-#'   - `~ stratum`: a stratified sample
+#'  - survey design: a named list of elements to pass to svydesign in the survey package
 #'
 #' @examples
-#' feline_prefs2 <- SurveyData$new(
+#' feline_prefs <- SurveyData$new(
 #'   feline_survey[,c("age1","gender","pet_own","y")],
 #'   questions = c(
 #'     "Please identify your age group",
@@ -33,7 +27,7 @@
 #'     c("no","yes")
 #'   ),
 #'   weights = feline_survey$wt,
-#'   design = formula("~.")
+#'   design = list(ids =~1, strata=~stype)
 #' )
 #' feline_prefs$print()
 #'
@@ -50,7 +44,7 @@
 #'     levels(approx_popn$pet_pref)
 #'   ),
 #'   weights = approx_popn$wt,
-#'   design = formula("~.")
+#'   design = NULL
 #' )
 #' popn_obj$print()
 #'
@@ -63,7 +57,7 @@ SurveyData <- R6::R6Class(
         questions = character(0),
         responses = list(),
         weights = numeric(),
-        design = as.formula("~."),
+        design = list(),
         initialize = function(survey_data,
                               questions = character(0),
                               responses = list(),
@@ -98,16 +92,6 @@ SurveyData <- R6::R6Class(
         },
         print = function(...) {
             cat("Survey containing", nrow(self$survey_data), "observations", "\n")
-            if (length(self$design == 2)) {
-                cat("Random Sampling Design \n")
-            } else {
-                cat("Simple")
-                if (is.null(findbars(self$design))){
-                    cat(" stratified sample with strata", all.vars(terms(self$design))[[1]], "\n")
-                } else {
-                    cat(" cluster sample with cluster", all.vars(terms(self$design))[[1]], "\n")
-                }
-            }
             for (i in 1:ncol(self$survey_data)) {
                 cat("Column ", i , " label: ", names(self$survey_data)[i], "\n")
                 if (length(self$questions) > 0) {

### 2021-03-16 R/SurveyMap.R

@@ -13,19 +13,19 @@
 #' they should be listed in that order, either descending or ascending.
 #'
 #' @examples
-#' feline_prefs <- SurveyData$new(
+#' feline_prefs = SurveyData$new(
 #'   feline_survey[,c("age1","gender","pet_own","y")],
 #'   questions = c("Please identify your age group","Please select your gender","Which pet do you own?", "Response"),
 #'   responses = list(levels(feline_survey$age1),levels(feline_survey$gender),levels(feline_survey$pet_own),c("no","yes")),
 #'   weights = feline_survey$wt,
-#'   design = formula("~.")
+#'   design = list(ids =~1, strata=~stype)
 #' )
-#' popn_obj <- SurveyData$new(
+#' popn_obj = SurveyData$new(
 #'   approx_popn[,c("age2","gender","pet_pref")],
 #'   questions = c("Which age group are you?","Gender?","Which pet would you like to own?"),
 #'   responses = list(levels(approx_popn$age2),levels(approx_popn$gender),levels(approx_popn$pet_pref)),
 #'   weights = approx_popn$wt,
-#'   design = formula("~.")
+#'   design = NULL
 #' )
 #' q1 <- SurveyQuestion$new(
 #'   name = "age",
@@ -64,8 +64,8 @@
 #' print(tmp_map)
 #' tmp_map$validate()
 #' tmp_map$mapping()
-#' tmp_map$tabulate("age") # Just use age in the poststrat matrix
-#' tmp_map$tabulate() # Use all variables in the map
+#' tmp_map$tabulate("age") #Just use age in the poststrat matrix
+#' tmp_map$tabulate() #Use all variables in the map
 #'
 #'
 #' mod_fit_1 <- tmp_map$fit(
@@ -371,7 +371,7 @@ SurveyMap <- R6::R6Class(
 
       if (is.null(args$fun)) {
         if ("stanreg" %in% class(fitted_model)){
-          require_suggested_package("rstanarm", "2.21.0")
+          require_suggested_package("rstanarm")
           return(
             t(rstanarm::posterior_epred(
               object = fitted_model,
@@ -422,6 +422,12 @@ SurveyMap <- R6::R6Class(
       }
     } else {
       posterior_preds <- data.frame(value = apply(poststrat_fit,2,function(x) sum(poststrat$N_j*x)/sum(poststrat$N_j)))
+    }
+    if(!is.null(self$samp_obj$design) & !is.null(self$samp_obj$weights) ){
[…]

### 2021-03-31 R/sim_data_approx_voters_popn.R
update the vignette and example data.

@@ -8,10 +8,10 @@
 #' @format A data frame with columns:
 #'
 #' \describe{
-#'  \item{age2}{Age group}
+#'  \item{age_group}{Age group}
 #'  \item{gender}{Gender}
-#'  \item{vote_for}{The voted party in the 2016 election}
-#'  \item{y}{Preference to NLP}
+#'  \item{vote_for}{The party that being voted in the 2016 election}
+#'  \item{y}{Preference to NLP in 2020 election}
 #'  \item{wt}{Raked and trimmed survey weight}
 #' }
 #' @examples

### 2021-03-31 R/sim_data_nlp_survey.R

@@ -9,10 +9,10 @@
 #' @format A data frame with columns:
 #'
 #' \describe{
-#'  \item{age1}{Age group}
+#'  \item{age}{Age group}
 #'  \item{gender}{Gender}
 #'  \item{vote_for}{The party that the respondent vote for in the 2016 election}
-#'  \item{y}{Preference to NLP}
+#'  \item{y}{Preference to NLP in 2020 election}
 #'  \item{wt}{Raked and trimmed survey weight}
 #' }
 #' @examples

### 2021-03-31 data-raw/NeverlandVoters.R

@@ -75,19 +75,19 @@ nlp_survey$wt <- weights(wts_trim_nlp_survey)
 nlp_survey <- nlp_survey %>%
   select(-S_inclusion_prob) %>%
   mutate_at(c("age_group","gender","vote_for","y"), as.factor) %>%
-  mutate(age1 =fct_recode(age_group, `18-25` = "1", `26-35` = "2", `36-45` = "3",`46-55` = "4", `56-65` = "5", `66-75` = "6", `76-90` = "7"),
+  mutate(age =fct_recode(age_group, `18-25` = "1", `26-35` = "2", `36-45` = "3",`46-55` = "4", `56-65` = "5", `66-75` = "6", `76-90` = "7"),
          gender = fct_recode(gender, "male" = "1", "female" = "2", "nonbinary" = "3"),
          vote_for = fct_recode(vote_for, "Neverland Labor Party" = "1", "NLP" = "2", "Neverland Democrats" = "3", "The Democrats" = "4"),
          y = fct_recode(y,"no" = "0", "yes" = "1")) %>%
-  select(c(age1, gender, vote_for, y, wt))
+  select(c(age, gender, vote_for, y, wt))
 
 
 approx_voters_popn <- approx_voters_popn %>%
   select(-S_inclusion_prob) %>%
   mutate_at(c("age_group","gender","vote_for"), as.factor) %>%
-  mutate(age2 =fct_recode(age_group, `18-35` = "1", `18-35` = "2", `36-55` = "3",`36-55` = "4", `56-65` = "5", `66+` = "6", `66+` = "7"),
+  mutate(age_group =fct_recode(age_group, `18-35` = "1", `18-35` = "2", `36-55` = "3",`36-55` = "4", `56-65` = "5", `66+` = "6", `66+` = "7"),
          gender = fct_recode(gender, "m" = "1", "f" = "2", "nb" = "3"),
          vote_pref = fct_recode(vote_for, "NLP" = "1", "NLP" = "2", "The Democrats" = "3", "The Democrats" = "4")) %>%
-  select(c(age2, gender, vote_pref, wt))
+  select(c(age_group, gender, vote_pref, wt))
 
 usethis::use_data(nlp_survey, approx_voters_popn, overwrite = TRUE)

### 2021-04-05 R/SurveyData.R
Update SurveyData.R

@@ -151,14 +151,14 @@ SurveyData <- R6::R6Class(
             if (key) {
                 private$survey_data_
             } else {
-                private$survey_data_[, colnames(private$survey_data_) != "key"]
+                private$survey_data_[, colnames(private$survey_data_) != "key", drop = FALSE]
             }
         },
         mapped_data = function(key = TRUE) {
             if (key) {
                 private$mapped_data_
             } else {
-                private$mapped_data_[, colnames(private$mapped_data_) != "key"]
+                private$mapped_data_[, colnames(private$mapped_data_) != "key", drop = FALSE]
             }
         },
         questions = function() private$questions_,

### 2021-04-06 R/SurveyData.R
key -> key_

@@ -95,8 +95,8 @@ SurveyData <- R6::R6Class(
             private$responses_ <- responses
             private$weights_ <- weights
             private$design_ <- design
-            private$survey_data_ <- data.frame(key = 1:nrow(survey_data), survey_data)
-            private$mapped_data_ <- data.frame(key = 1:nrow(survey_data))
+            private$survey_data_ <- data.frame(key_ = 1:nrow(survey_data), survey_data)
+            private$mapped_data_ <- data.frame(key_ = 1:nrow(survey_data))
             invisible(self)
         },
 
@@ -151,14 +151,14 @@ SurveyData <- R6::R6Class(
             if (key) {
                 private$survey_data_
             } else {
-                private$survey_data_[, colnames(private$survey_data_) != "key", drop = FALSE]
+                private$survey_data_[, colnames(private$survey_data_) != "key_", drop = FALSE]
             }
         },
         mapped_data = function(key = TRUE) {
             if (key) {
                 private$mapped_data_
             } else {
-                private$mapped_data_[, colnames(private$mapped_data_) != "key", drop = FALSE]
+                private$mapped_data_[, colnames(private$mapped_data_) != "key_", drop = FALSE]
             }
         },
         questions = function() private$questions_,

### 2021-04-06 R/SurveyData.R
use list instead of vector for questions
closes #15

@@ -20,7 +20,7 @@
 #' @examples
 #' feline_prefs <- SurveyData$new(
 #'   feline_survey[,c("age1","gender","pet_own","y")],
-#'   questions = c(
+#'   questions = list(
 #'     "Please identify your age group",
 #'     "Please select your gender",
 #'     "Which pet do you own?",
@@ -39,7 +39,7 @@
 #'
 #' popn_obj <- SurveyData$new(
 #'   approx_popn[,c("age2","gender","pet_pref")],
-#'   questions = c(
+#'   questions = list(
 #'     "Which age group are you?",
 #'     "Gender?",
 #'     "Which pet would you like to own?"
@@ -59,14 +59,14 @@ SurveyData <- R6::R6Class(
     private = list(
         survey_data_ = data.frame(NULL),
         mapped_data_ = data.frame(NULL),
-        questions_ = character(0),
+        questions_ = list(),
         responses_ = list(),
         weights_ = numeric(),
         design_ = as.formula("~.")
     ),
     public = list(
         initialize = function(survey_data,
-                              questions = character(0),
+                              questions = list(),
                               responses = list(),
                               weights = numeric(),
                               design = as.formula("~.")) {
@@ -77,7 +77,7 @@ SurveyData <- R6::R6Class(
             # allow no question/response, else require info for all columns
             if (length(questions) != 0 || length(responses) != 0) {
                 if (ncol(survey_data) != length(questions) &
-                    length(questions) == sum(stats::complete.cases(questions))) {
+                    length(questions) == sum(stats::complete.cases(survey_data))) {
                     stop("mismatch between number of survey_data columns and questions.",
                          call. = FALSE)
                 }
@@ -126,7 +126,7 @@ SurveyData <- R6::R6Class(
             for (i in 1:ncol(self$survey_data(key = FALSE))) {
                 cat("\nColumn label:", names(self$survey_data(key = FALSE))[i], "\n")
                 if (length(private$questions_) > 0) {
-                    cat("Question:", private$questions_[i], "\n")
+                    cat("Question:", private$questions_[[i]], "\n")
                     cat("Allowed answers:",
                         paste(private$responses_[[i]], collapse = ", "), "\n")
                 }

### 2021-04-06 R/SurveyMap.R

@@ -15,14 +15,14 @@
 #' @examples
 #' feline_prefs <- SurveyData$new(
 #'   feline_survey[,c("age1","gender","pet_own","y")],
-#'   questions = c("Please identify your age group","Please select your gender","Which pet do you own?", "Response"),
+#'   questions = list("Please identify your age group","Please select your gender","Which pet do you own?", "Response"),
 #'   responses = list(levels(feline_survey$age1),levels(feline_survey$gender),levels(feline_survey$pet_own),c("no","yes")),
 #'   weights = feline_survey$wt,
 #'   design = formula("~.")
 #' )
 #' popn_obj <- SurveyData$new(
 #'   approx_popn[,c("age2","gender","pet_pref")],
-#'   questions = c("Which age group are you?","Gender?","Which pet would you like to own?"),
+#'   questions = list("Which age group are you?","Gender?","Which pet would you like to own?"),
 #'   responses = list(levels(approx_popn$age2),levels(approx_popn$gender),levels(approx_popn$pet_pref)),
 #'   weights = approx_popn$wt,
 #'   design = formula("~.")

### 2021-04-06 R/SurveyData.R
allow list(question = response_values)
closes #35

@@ -75,15 +75,20 @@ SurveyData <- R6::R6Class(
                      call. = FALSE)
             }
             # allow no question/response, else require info for all columns
-            if (length(questions) != 0 || length(responses) != 0) {
+            if (length(questions) != 0) {
                 if (ncol(survey_data) != length(questions) &
                     length(questions) == sum(stats::complete.cases(survey_data))) {
                     stop("mismatch between number of survey_data columns and questions.",
                          call. = FALSE)
                 }
-                if (length(responses) != length(questions)) {
-                    stop("mismatch between number of survey questions and answers.",
-                         call. = FALSE)
+                if (length(responses) == 0) {
+                    responses <- unname(questions)
+                    questions <- as.list(names(questions))
+                } else {
+                    if (length(responses) != length(questions)) {
+                        stop("mismatch between number of survey questions and answers.",
+                             call. = FALSE)
+                    }
                 }
             }
             # allow no weights, else require weights for all columns

### 2021-04-18 R/sim_data_approx_voters_popn.R
Add education and state in the simulated data.

@@ -11,7 +11,8 @@
 #'  \item{age_group}{Age group}
 #'  \item{gender}{Gender}
 #'  \item{vote_for}{The party that being voted in the 2016 election}
-#'  \item{y}{Preference to NLP in 2020 election}
+#'  \item{education}{Highest grade of education that is completed}
+#'  \item{state}{State}
 #'  \item{wt}{Raked and trimmed survey weight}
 #' }
 #' @examples

### 2021-04-18 R/sim_data_nlp_survey.R

@@ -12,6 +12,8 @@
 #'  \item{age}{Age group}
 #'  \item{gender}{Gender}
 #'  \item{vote_for}{The party that the respondent vote for in the 2016 election}
+#'  \item{highest_educ}{Highest grade of education that is completed}
+#'  \item{state}{State}
 #'  \item{y}{Preference to NLP in 2020 election}
 #'  \item{wt}{Raked and trimmed survey weight}
 #' }

### 2021-04-18 data-raw/NeverlandVoters.R

@@ -10,10 +10,15 @@ library(forcats)
 design <- declare_population(N = 50000,
                              age_group = sample.int(7, size=N, replace = TRUE, prob = c(1,1,2,2,3,2,1)),
                              gender = sample.int(3, size=N, replace = TRUE, prob = c(.48,.49,.03)),
+                             education = sample.int(6, size=N, replace = TRUE, prob = c(.05,.25,.27,.1,.3,.03)),
+                             state = sample.int(5, size=N, replace = TRUE, prob = c(.28,.4,.12,.15,.05)),
                              vote_for = sample.int(4, size=N, replace = TRUE, prob = abs(rnorm(4,0,1))),
                              y = rbinom(N,1,inv_logit_scaled(c(1,-.5,.5)[gender] +
                                                                c(-.4,.4,.5,1.0)[vote_for] +
-                                                               c(-.2,-.5,.5,.5,1.0,.4,.7)[age_group])))
+                                                               c(-.2,-.5,.5,.5,1.0,.4,.7)[age_group] +
+                                                               c(-.2,.7,-.5,-.5,.4,.7)[education] +
+                                                               c(-.3,-.6,.6,.3,.4)[state])))
+
 ### get population
 popn <- draw_data(design +
                     declare_sampling(n=50000))
@@ -22,7 +27,9 @@ popn <- draw_data(design +
 get_sample <- function(data, n) {
   p = inv_logit_scaled(.5 + c(-1,-2,-1,-.5,1,.5,.1)[data$age_group]+
                          c(2,-.5,0)[data$gender] +
-                         c(1,-.5,-1,.5)[data$vote_for])
+                         c(1,-.5,-1,.5)[data$vote_for] +
+                         c(-1,1,-2,-.5,1,.5)[data$education] +
+                         c(-.2,-.7,.6,.5,.4)[data$state])
   data[sample.int(nrow(data),n,replace=TRUE,prob = p), ]
 }
 
@@ -30,7 +37,9 @@ get_sample <- function(data, n) {
 get_pop_sample <- function(data, n) {
   p = inv_logit_scaled(.5 + c(-.1,-.2,-.1,-.5,.1,.5,.01)[data$age_group]+
                          c(.2,-.5,0)[data$gender] +
-                         c(.1,-.5,-.1,.5)[data$vote_for])
+                         c(.1,-.5,-.1,.5)[data$vote_for] +
[…]

### 2021-04-27 R/SurveyFit.R
document more arguments and methods

@@ -2,11 +2,12 @@
 #'
 #' @name SurveyFit
 #' @export
-#' @description An [R6][R6::R6Class] SurveyFit object stores a fitted model
+#' @description An [R6][R6::R6Class] `SurveyFit` object stores a fitted model
 #'   object and provides methods for generating predicted probabilities for all
 #'   poststrat cells, generating population and group estimates, and visualizing
 #'   results.
 #' @inherit SurveyMap examples
+#'
 SurveyFit <- R6::R6Class(
   classname = "SurveyFit",
   private = list(
@@ -14,6 +15,12 @@ SurveyFit <- R6::R6Class(
     fit_ = NULL
   ),
   public = list(
+
+    #' @description Create a new `SurveyFit` object. This method is called
+    #'   internally by the `$fit()` method of the [`SurveyMap`] object and does
+    #'   not need to be called directly by the user.
+    #' @param fit A fitted model object.
+    #' @param map A [`SurveyMap`] object.
     initialize = function(fit, map) {
       private$fit_ <- fit
       private$map_ <- map
@@ -21,13 +28,11 @@ SurveyFit <- R6::R6Class(
     },
 
     #' @description Access the fitted model object
-    #' @return The fitted model object.
     fit = function() {
       private$fit_
     },
 
     #' @description Access the SurveyMap object
-    #' @return The [SurveyMap] object.
     map = function() {
       private$map_
     },

### 2021-04-27 R/SurveyMap.R

@@ -173,6 +173,7 @@ SurveyMap <- R6::R6Class(
     },
 
     #' @description Print a summary of the mapping.
+    #' @param ... Currently ignored.
     print = function(...) {
       if (length(private$item_map_) > 0) {
         for (i in 1:length(private$item_map_)) {
@@ -193,6 +194,7 @@ SurveyMap <- R6::R6Class(
     },
 
     #' @description Add new [SurveyQuestion]s.
+    #' @param ... The [SurveyQuestion]s to add.
     add = function(...) {
       dots <- list(...)
       for (i in 1:length(dots)) {

### 2021-05-04 R/SurveyMap.R
predictify -> population_predict in tests

@@ -424,7 +424,7 @@ SurveyMap <- R6::R6Class(
       admin_package <- as.character(getNamespaceName(environment(fun)))
       if (!any(admin_package %in% c("lme4","brms","rstanarm"))){
         warning("Only rstanarm, brms and lme4 are supported natively. ",
-                "Other modelling tools will need a custom predictify method.",
+                "Other modeling tools will need a custom population_predict() method.",
                 call. = FALSE)
       }
 

### 2021-05-04 tests/testthat/test-SurveyFit.R

@@ -111,20 +111,20 @@ test_that("Model fits do not cause errors ",{
     ), regexp = NA)
 })
 
-test_that("Predictify runs without errors",{
+test_that("population_predict runs without errors",{
   skip_if_not_installed("rstanarm")
   expect_error(poststrat_preds_rstanarm <-
-                 rstanarm_fit$predictify(),
+                 rstanarm_fit$population_predict(),
                regexp = NA)
 
   skip_if_not_installed("lme4")
   expect_error(poststrat_preds_lme4 <-
-                 lme4_fit$predictify(),
+                 lme4_fit$population_predict(),
                regexp = NA)
 
   skip_if_not_installed("brms")
   expect_error(poststrat_preds_brms <-
-                 brms_fit$predictify(),
+                 brms_fit$population_predict(),
                regexp = NA)
 })
 

### 2021-05-31 R/SurveyMap.R
Change the level of party in the example.

@@ -68,7 +68,7 @@
 #' q_party_pref <- SurveyQuestion$new(
 #'   name = "party_pref",
 #'   col_names = c("vote_for","vote_pref"),
-#'   values_map = list("Box Party" = "The BP", "The BP" = "The BP","The Circle Party" = "The Circles","The Circles" = "The Circles")
+#'   values_map = list("Box Party" = "BP",  "BP" = "BP","Circle Party" = "CP", "CP" = "CP")
 #' )
 #' q_gender <- SurveyQuestion$new(
 #'   name = "gender",

### 2021-06-07 R/SurveyData.R
Typo that prevented examples working, changed

@@ -14,7 +14,7 @@
 #' box_prefs <- SurveyData$new(
 #'   data = shape_survey,
 #'   questions = list(
-#'     age1 = "Please identify your age group",
+#'     age = "Please identify your age group",
 #'     gender = "Please select your gender",
 #'     vote_for = "Which party did you vote for in the 2018 election?",
 #'     y = "If today is the election day, would you vote for the Box Party?"

### 2021-06-07 R/SurveyMap.R

@@ -19,7 +19,7 @@
 #' box_prefs <- SurveyData$new(
 #'   data = shape_survey,
 #'   questions = list(
-#'     age1 = "Please identify your age group",
+#'     age = "Please identify your age group",
 #'     gender = "Please select your gender",
 #'     vote_for = "Which party did you vote for in the 2018 election?",
 #'     y = "If today is the election day, would you vote for the Box Party?"

### 2021-06-07 R/SurveyMap.R
Changed all occurances of SurveyQuestion to QuestionMap in files and filenames, rebuilt doc

@@ -55,7 +55,7 @@
 #' )
 #' popn_obj$print()
 #'
-#' q_age <- SurveyQuestion$new(
+#' q_age <- QuestionMap$new(
 #'   name = "age",
 #'   col_names = c("age","age_group"),
 #'   values_map = list(
@@ -65,12 +65,12 @@
 #' )
 #' print(q1)
 #'
-#' q_party_pref <- SurveyQuestion$new(
+#' q_party_pref <- QuestionMap$new(
 #'   name = "party_pref",
 #'   col_names = c("vote_for","vote_pref"),
 #'   values_map = list("Box Party" = "BP",  "BP" = "BP","Circle Party" = "CP", "CP" = "CP")
 #' )
-#' q_gender <- SurveyQuestion$new(
+#' q_gender <- QuestionMap$new(
 #'   name = "gender",
 #'   col_names = c("gender", "gender"),
 #'   values_map = list("male" = "m","female" = "f", "nonbinary" = "nb")
@@ -167,7 +167,7 @@ SurveyMap <- R6::R6Class(
     #' @description Create a new SurveyMap object
     #' @param sample The [SurveyData] object corresponding to the sample data.
     #' @param population The [SurveyData] object corresponding to the population data.
-    #' @param ... [SurveyQuestion] objects.
+    #' @param ... [QuestionMap] objects.
     initialize = function(sample, population, ...) {
       if (!inherits(sample, "SurveyData")) {
         stop("'sample' must be a SurveyData object.", call. = FALSE)
@@ -209,8 +209,8 @@ SurveyMap <- R6::R6Class(
       invisible(self)
     },
 
-    #' @description Add new [SurveyQuestion]s.
-    #' @param ... The [SurveyQuestion]s to add.
+    #' @description Add new [QuestionMap]s.
+    #' @param ... The [QuestionMap]s to add.
     add = function(...) {
       dots <- list(...)
       for (i in 1:length(dots)) {
@@ -225,12 +225,12 @@ SurveyMap <- R6::R6Class(
       invisible(self)
     },
 
-    #' @description Delete [SurveyQuestion]s.
-    #' @param ... The [SurveyQuestion]s to delete.
+    #' @description Delete [QuestionMap]s.
+    #' @param ... The [QuestionMap]s to delete.
     delete = function(...) {
       tmp_list <- list(...)
       for (i in length(tmp_list)) {
[…]

### 2021-06-07 tests/testthat/test-SurveyFit.R

@@ -31,7 +31,7 @@ popn_obj <- SurveyData$new(
   weights = approx_popn$wt,
   design = list(ids =~1))
 
-q1 <- SurveyQuestion$new(
+q1 <- QuestionMap$new(
   name = "age",
   col_names = c("age1","age2"),
   values_map = list(
@@ -39,12 +39,12 @@ q1 <- SurveyQuestion$new(
     "46-55" = "36-55", "56-65" = "56-65", "66-75" = "66+", "76-90" = "66+"
   )
 )
-q2 <- SurveyQuestion$new(
+q2 <- QuestionMap$new(
   name = "pet",
   col_names = c("pet_own","pet_pref"),
   values_map = list("cat" = "cat", "kitten" = "cat","dog" = "dog","puppy" = "dog")
 )
-q3 <- SurveyQuestion$new(
+q3 <- QuestionMap$new(
   name = "gender",
   col_names = c("gender","gender"),
   values_map = data.frame("male" = "m","female" = "f", "nonbinary" = "nb")

### 2021-06-07 tests/testthat/test-SurveyMap.R

@@ -2,7 +2,7 @@ suppressWarnings({
   samp <- SurveyData$new(data = feline_survey, weights = feline_survey$wt)
   popn <- SurveyData$new(data = approx_popn, weights = approx_popn$wt)
 })
-q1 <- SurveyQuestion$new(
+q1 <- QuestionMap$new(
   name = "age",
   col_names = c("age1","age2"),
   values_map = list(
@@ -10,12 +10,12 @@ q1 <- SurveyQuestion$new(
     "46-55" = "36-55", "56-65" = "56-65", "66-75" = "66+", "76-90" = "66+"
   )
 )
-q2 <- SurveyQuestion$new(
+q2 <- QuestionMap$new(
   name = "pet",
   col_names = c("pet_own","pet_pref"),
   values_map = list("cat" = "cat", "kitten" = "cat","dog" = "dog","puppy" = "dog")
 )
-q3 <- SurveyQuestion$new(
+q3 <- QuestionMap$new(
   name = "gender",
   col_names = c("gender","gender"),
   values_map = data.frame("male" = "m","female" = "f", "nonbinary" = "nb")
@@ -81,7 +81,7 @@ test_that("validate creates correct levels (example1)", {
     design = formula("~.")
   )
 
-  q1 <- SurveyQuestion$new(
+  q1 <- QuestionMap$new(
     name = "age",
     col_names = c("age1", "age2"),
     values_map = list("18-25" = "18-45", "26-45" = "18-45", "46+" = "46+")
@@ -123,7 +123,7 @@ test_that("validate creates correct levels (example2)", {
     design = formula("~.")
   )
 
-  q1 <- SurveyQuestion$new(
+  q1 <- QuestionMap$new(
     name = "age",
     col_names = c("age1", "age2"),
     values_map = list("18-25" = "18-45", "26-45" = "18-45", "46+" = "46+")
@@ -166,7 +166,7 @@ test_that("validate creates correct levels (example3)", {
     design = formula("~.")
   )
 
-  q1 <- SurveyQuestion$new(
+  q1 <- QuestionMap$new(
     name = "age",
     col_names = c("age1", "age2"),
     values_map = list("18-25" = "18-25", "26+" = "26-34","26+" = "35+")
@@ -208,7 +208,7 @@ test_that("validate creates correct levels (example4)", {
     design = formula("~.")
   )
 
-  q1 <- SurveyQuestion$new(
+  q1 <- QuestionMap$new(
     name = "age",
[…]

### 2021-06-07 R/SurveyData.R
Added more examples and description to survey data

@@ -4,9 +4,11 @@
 #' @export
 #' @description
 #' A `SurveyData` object represents a survey and its metadata. The survey itself
-#' is a data frame. The survey metadata consists of the text of the survey
-#' questions, the allowed response values, and optionally survey weights and a
-#' survey design specification.
+#' is a data frame containing all data from the survey. The survey data object also
+#' includes the survey questions and responses (if left empty these will just be the
+#' column and factor level names). To enable weights comparison, a weights column and
+#' survey design can be specified, with the survey design specified using `survey` package
+#' notation.
 #'
 #' @examples
 #'
@@ -60,7 +62,14 @@ SurveyData <- R6::R6Class(
     design_ = list()
   ),
   public = list(
-    #' @description Create a new SurveyData object
+    #' @description Create a new SurveyData object using an existing data frame and other
+    #' survey information.
+    #' The same object is being used for a sample and population.
+    #' If a population is approximated from a large survey (like the ACS or DHS),
+    #' then the package will enable the creation of a weighted poststratification matrix.
+    #' If the population is summarized as a poststratification matrix already, then set the weights
+    #' as the size in each cell $N_j$.
+    #' If the entire individual level population data is given, then specify the weights as 1.
     #' @param data A data frame containing the survey data.
     #' @param questions,responses Named lists containing the text of the survey
     #'   questions and the allowed responses, respectively. The names must
@@ -70,6 +79,63 @@ SurveyData <- R6::R6Class(
     #' @param weights Optionally, a vector of survey weights.
     #' @param design Optionally, a named list of arguments (except `weights` and
     #'   `data`) to pass to `survey::svydesign()` to specify the survey design.
+    #' @examples
+    #' #Population estimated from large survey
+    #' popn_obj1 <- SurveyData$new(
+    #'   data = approx_voters_popn,
+    #'   questions = list(
+    #'     age_group = "Which age group are you?",
+    #'     gender = "Gender?",
+    #'     vote_pref = "Which party do you prefer to vote for?"
+    #'   ),
+    #'   # order doesn't matter (gender before age2 here) because
+    #'   # the list has the names of the variables
+    #'   responses = list(
+    #'     gender = levels(approx_voters_popn$gender),
[…]

### 2021-06-17 paper/paper.Rmd

@@
 Typically MRP requires at least two datasets - usually a survey object and a post-stratification object, which we implement using the `SurveyData` method.
-The questions in each datasets are then aligned by establishing a correspondence between each possible answer to a particular question, and we implement this using the `SurveyQuestion` method.
+The questions in each datasets are then aligned by establishing a correspondence between each possible answer to a particular question, and we implement this using the `QuestionMap` method.
 Having linked each question, the surveys themselves can be  linked and we implement this using the `SurveyMap` method.
@@
 
-The `MRPKit` package has the following key components: `SurveyData` objects, which for most users will be a collection of two surveys where one is larger than the other; a `SurveyQuestion` object, that holds the mapping for one question or demographic between the survey and post-stratification dataset; and a `SurveyMap` object, which holds the mapping between a set of items in a survey and a post-stratification dataset.
+The `MRPKit` package has the following key components: `SurveyData` objects, which for most users will be a collection of two surveys where one is larger than the other; a `QuestionMap` object, that holds the mapping for one question or demographic between the survey and post-stratification dataset; and a `SurveyMap` object, which holds the mapping between a set of items in a survey and a post-stratification dataset.
 These are expanded on in the following sub-sections before being brought together.
@@
 
-## SurveyQuestion objects
+## QuestionMap objects
 
-A `SurveyQuestion` object holds the mapping for one question or demographic between the survey and post-stratification dataset.
+A `QuestionMap` object holds the mapping for one question or demographic between the survey and post-stratification dataset.
 For instance, if both the survey and post-stratification dataset contain a question about age, it may be that answers are expressed as a single integer representing years and so mapping them is relatively straight-forward.
@@
 ```{r, eval = FALSE}
-SurveyQuestion$new(
+QuestionMap$new(
   name = "age",
@@
 A `SurveyMap` object holds the mapping between a set of items in a survey and a population dataset.
-That is, it draws on two `SurveyData` objects, and a `SurveyQuestion` object, and brings the two `SurveyData` objects together following the correspondences provided in the `SurveyQuestion` object.
+That is, it draws on two `SurveyData` objects, and a `QuestionMap` object, and brings the two `SurveyData` objects together following the correspondences provided in the `QuestionMap` object.
 
@@
 A model is fit to a `SurveyMap` object using the function `fit`.
-After a model has been fit it can be used to get MRP-based estimates using `predictify()`.
-And finally the estimates can be visualized using `visify()`.
+After a model has been fit it can be used to get MRP-based estimates using `population_predict()`.
+And finally the estimates can be visualized using `plot()`.
 
@@
 A user could also specify their own modelling approach if they wanted to.
-In any case, once a model has been fit to the survey dataset, it can be used to create estimates based on the post-stratification dataset with `predictify`, which takes a fitted model as an argument.
-Typically a user is interested in some type of high-level estimates, for instance on a province/state basis, and `collapsify` provides the functionality to do this.
-Finally, the estimates on that collapsed basis can be visualized with `visify`
+In any case, once a model has been fit to the survey dataset, it can be used to create estimates based on the post-stratification dataset with `population_predict`, which takes a fitted model as an argument.
+Typically a user is interested in some type of high-level estimates, for instance on a province/state basis, and `aggregate` provides the functionality to do this.
+Finally, the estimates on that collapsed basis can be visualized with `plot`
 

### 2021-06-21 R/SurveyFit.R
change aggregation

@@ -77,17 +77,17 @@ SurveyFit <- R6::R6Class(
         if ("stanreg" %in% class(private$fit_)) {
           require_suggested_package("rstanarm", "2.21.0")
           return(
-            t(suppressMessages(rstanarm::posterior_linpred(
+            suppressMessages(rstanarm::posterior_linpred(
               object = private$fit_,
               newdata = poststrat,
               transform = TRUE,
               ...
-            )))
+            ))
           )
         } else if ("brmsfit" %in% class(private$fit_)) {
           require_suggested_package("brms")
           return(
-            t(brms::posterior_epred(
+            brms::posterior_epred(
               object = private$fit_,
               newdata = poststrat,
               dpar = "mu",
@@ -96,10 +96,11 @@ SurveyFit <- R6::R6Class(
                 if (!is.null(args$sample_new_levels)) args$sample_new_levels
                 else "gaussian",
               ...
-            ))
+            )
           )
         } else if ("glmerMod" %in% class(private$fit_)) {
           require_suggested_package("lme4")
+          # note:  might need to transpose this?
           return(
             sim_posterior_probs(
               object = private$fit_,
@@ -130,45 +131,49 @@ SurveyFit <- R6::R6Class(
     #'   variable each row corresponds to.
     aggregate = function(poststrat_estimates, by = NULL) {
       poststrat_data <- private$map_$poststrat_data()
-      if (!is.null(by)) {
-        if (length(by) != 1) {
-          stop("Currently only one variable can be named in 'by'.", call. = FALSE)
-        }
-        rotate_levels <- levels(private$map_$mapped_sample_data()[, by])
-        out <- expand.grid(by = rotate_levels, draw = 1:ncol(poststrat_estimates), value = NA)
-        colnames(out)[1] <- by
-        for (focus_level in rotate_levels){
-          level_loc <- poststrat_data[by] == focus_level
-          out[out[by] == focus_level, "value"] <-
-            apply(poststrat_estimates[level_loc, ], 2, function(x) sum(poststrat_data$N_j[level_loc]*x)/sum(poststrat_data$N_j[level_loc]))
-        }
-      } else {
[…]

### 2021-06-22 R/SurveyMap.R
Use drop =  FALSE to simplify code

@@ -375,50 +375,37 @@ SurveyMap <- R6::R6Class(
         }
         tmp_mapped_levels <- mapped_levels
         mapped_levels_new <- mapped_levels
-        mapped_col_names <- colnames(mapped_levels)
-        mapped_col_names_new <- colnames(mapped_levels)
-        mapped_col_names_tmp <- colnames(mapped_levels)
         for (unique_samp_levels in 1:length(mapped_levels[,1])) {
           if(sum(mapped_levels[unique_samp_levels,],na.rm = TRUE)>1){
-            mapped_levels_new <- tmp_mapped_levels[,tmp_mapped_levels[unique_samp_levels,]!=1]
-            mapped_col_names_new <- mapped_col_names_tmp[tmp_mapped_levels[unique_samp_levels,]!=1]
+            mapped_levels_new <- tmp_mapped_levels[,tmp_mapped_levels[unique_samp_levels,]!=1, drop = FALSE]
             ll = ncol(mapped_levels_new)
             mapped_levels_new <- cbind(mapped_levels_new,ifelse(rowSums(tmp_mapped_levels[,tmp_mapped_levels[unique_samp_levels,]==1])>=1,1,0))
-            mapped_col_names_new <- c(mapped_col_names_new,paste0(mapped_col_names_tmp[tmp_mapped_levels[unique_samp_levels,]==1], collapse = " + "))
-            tmp_mapped_levels<- mapped_levels_new
-            mapped_col_names_tmp <- mapped_col_names_new
-          }
+            colnames(mapped_levels_new)[ll+1] <- paste0(colnames(tmp_mapped_levels[,tmp_mapped_levels[unique_samp_levels,,drop = FALSE]==1,drop = FALSE]), collapse = " + ")
+            tmp_mapped_levels<- mapped_levels_new        }
         }
 
         tmp_mapped_levels <- mapped_levels_new
         mapped_levels_fin <- mapped_levels_new
-        mapped_row_names <- row.names(mapped_levels)
-        mapped_row_names_new <- row.names(mapped_levels)
-        mapped_row_names_tmp <- row.names(mapped_levels)
         for (unique_popn_levels in 1:length(mapped_levels_fin[1,])) {
           if(sum(mapped_levels_new[,unique_popn_levels],na.rm = TRUE)>1){
-            mapped_levels_new <- tmp_mapped_levels[tmp_mapped_levels[,unique_popn_levels]!=1,]
[…]

### 2021-06-22 R/SurveyMap.R
Added doc

@@ -344,7 +344,13 @@ SurveyMap <- R6::R6Class(
       invisible(self)
     },
 
-    #' @description Prepare the mapped data
+    #' @description The mapping method uses the given maps between questions to create new sample and population data
+    #' that has unified variable names (i.e., if the underlying construct is called `age`, both sample and population will have
+    #' an `age` column, even if in the the raw data both had different variable names). The method also unifies the levels of each
+    #' variable in the sample and population so that the maximum set of consistent levels is created. Names of these new levels are
+    #' made according the the sample level names. If multiple levels are combined, the new name will be existing levels seperated by a
+    #' ` + `.
+    #'
     mapping  = function() {
       for (j in 1:length(private$item_map_)) {
         samp_mapnames <- private$item_map_[[j]]$col_names()[1]
@@ -357,7 +363,10 @@ SurveyMap <- R6::R6Class(
         new_levels_samp <- character(length(levels_map_samp))
         new_levels_popn <- character(length(levels_map_popn))
 
-        # Find the naming clusters of levels  (loosely inspired by complete linkage clustering algorithms)
+        # Find the naming clusters of levels  (loosely inspired by complete linkage clustering algorithms.
+        # Three major steps:
+        # Step 1:  Create a matrix that has the factor level labels for the sample as row names, and factor level labels for the
+        # population as the column names. If the levels are linked, record 1 in the corresponding entry, otherwise record 0.
         mapped_levels <- matrix(0,nrow=length(unique(levels_map_samp)),ncol = length(unique(levels_map_popn)))
         colnames(mapped_levels)<-levels_popn
         row.names(mapped_levels)<-levels_samp
@@ -373,6 +382,9 @@ SurveyMap <- R6::R6Class(
         if(sum(colSums(mapped_levels, na.rm=TRUE)!=0)<ncol(mapped_levels)){
           stop("Levels: ",paste(colnames(sum(colSums(mapped_levels, na.rm=TRUE)!=0)),collapse = " ")," do not have a match in the sample.")
         }
+
+        # Moving down the rows (i.e., for each factor level in the sample), identify if there are multiple corresponding population
+        # values. If there are, merge the columns for the values together and combine the col names together, seperated by a ' + ' symbol.
         tmp_mapped_levels <- mapped_levels
         mapped_levels_new <- mapped_levels
[…]

### 2021-06-22 tests/testthat/test-SurveyFit.R
Changed to expect_equal with a tolerance and increased tolerance

@@ -259,23 +259,20 @@ test_that("Populations are within acceptable tolerance of previous runs (+/- 2% 
 
   skip_if_not_installed("rstanarm")
   x <- fit_stan_glmer$aggregate(fit_stan_glmer$population_predict())
-  expect_lt(mean(x$value), .72 + .02)
-  expect_gt(mean(x$value), .72 - .02)
+  expect_equal(mean(x$value), expected = .72, tolerance = .05)
 
   skip_if_not_installed("lme4")
   #lme4 consistently lower
   x <- fit_glmer$aggregate(fit_glmer$population_predict(nsamples = 50))
-  expect_lt(mean(x$value), .68 + .02)
-  expect_gt(mean(x$value), .68 - .02)
+  expect_equal(mean(x$value), expected = .68, tolerance = .05)
 
   skip_if_not_installed("brms")
   x <- fit_brms$aggregate(fit_brms$population_predict())
-  expect_lt(mean(x$value), .72 + .02)
-  expect_gt(mean(x$value), .72 - .02)
+  expect_equal(mean(x$value), expected = .72, tolerance = .05)
 
+  #This is VERY noisy
   x <- fit_stan_glm$aggregate(fit_stan_glm$population_predict())
-  expect_lt(mean(x$value), .72 + .02)
-  expect_gt(mean(x$value), .72 - .02)
+  expect_equal(mean(x$value), expected = .72, tolerance = .15)
 
 })
 

### 2021-06-29 case-study/Munging_CCES.Rmd
writeup on working with large surveys

---
title: "Processing large surveys with package MRP-KIT"
author: "Mitzi Morris"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, message=FALSE, error=FALSE, warning=FALSE, comment=NA, out.width='95%', collapse=TRUE)
```

<style type="text/css">
.table { width: 40%; }
div.quote-container blockquote {  font-size: small; }
</style>

The goal of this case study is to show how the MRP-KIT package
can facilitate analysis of a large-scale survey sample.
By large-scale, we mean surveys which have both a rich set
of demographics and a large number of responses.
The richer the set of demographics, the larger the number
of possible sub-populations of interest; no matter how large
or representative the survey, the combinatorics of the demographic
categories are such that it is unlikely that the sample
is representative for all subpopulations of interest.
In order to correct this imbalance,
we post-stratify to a truly representative
sample drawn from the general population.
This, in turn, requires obtaining a general population sample;
one which is at least as large and which contains all the
demographics of interest.

In this case study, the survey sample is one year's worth of data
from is Cooperative Congressional Election Study (CCES),
a bi-annual survey of of 60,000 American adults from
all 50 states and the District of Columbia.
The CCES is designed to capture 
the views and experiences of Amercians with respect to Congress
and congressional (and presidential) elections.
While this survey sample is representative at the national level
it is not large enough to provide a representative sample for sub-populations
of interest.
The population sample is the 2018 American Community Survey, the
annual micro-census conducted by the US Census Bureau.  It is designed
to be a representative sample of the US population as a whole; data is
collected from 3 million households.

Big data is precious data, but a researcher's time is even more precious.
A secondary goal of this case study is showing how to work effectively
and efficiently with large samples to produce replicable results.

### Survey sample: the 2018 CCES Common Content Dataset

The 2018 CCES Common Content Dataset is 
available from the Harvard Dataverse project:
https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi%3A10.7910/DVN/ZSBZ7K
The raw survey data is distributed in several from the Harvard Dataverse
as  [cces18_common_vv](https://dataverse.harvard.edu/file.xhtml?persistentId=doi:10.7910/DVN/ZSBZ7K/C1YVSF&version=6.0).
Although the tabular versions of the data file are smaller, we download the
Stata file, as it contains metadata on all the questions; this is critical to
being able to align the CCES sample to the US Census population sample.
We also download the [2018 CCES Guide](https://dataverse.harvard.edu/file.xhtml?persistentId=doi:10.7910/DVN/ZSBZ7K/WZWCZ1&version=6.0)
which describes the survey questions, labels, and responses.

We load the data into R using the [`readstata13`](https://sjewo.github.io/readstata13/) library function
[`read.dta13`](https://sjewo.github.io/readstata13/reference/read.dta13.html) which returns a data.frame with attributes.
```{r}
library(readstata13)
cces_2018 = read.dta13("cces18_common_vv.dta", convert.factors = T, nonint.factors = T, generate.factors=T)
```

Fist we check the type, shape and contents of the sample.
```{r}
str(cces_2018, max.level=2, list.len=12, give.attr=FALSE)
dim(cces_2018)
names(cces_2018)[1:12]
```

The Common Content portion of the survey consists of questions asked of all subjects.
Profile questions provide demographic information,
and opinion questions provide outcomes of interest.
The demographic variables include
gender, education level, employment status, race, and geographic location.
```{r}
table(cces_2018$gender[,drop=TRUE])
table(cces_2018$educ[,drop=TRUE])
table(cces_2018$race[,drop=TRUE])
```
The political opinion questions on the survey are labelled "Abortion", "Immigration", "Taxes", "Health Care".
```{r}
attr(cces_2018, "var.labels")[which(startsWith(attr(cces_2018, "var.labels"), "Health Care"))[1]]
health_care_1 <- which(startsWith(attr(cces_2018, "var.labels"), "Health Care"))[1]
table(cces_2018[[health_care_1]][, drop=TRUE])
```

Of the 525 columns in the survey, we are interested in the set of political opinion questons
and a subset of the demographics.  Although not strictly necessary, we create a new dataframe
which contains just these columns of interest.
The survey metadata includes column names, and question and response labels.
We use the column names to select demographics and column labels to select political opinion questions.
```{r}
col_labels <- attr(cces_2018, "var.labels")
col_names <- attr(cces_2018, "names")
demo_cols <- c(1, which(col_names == "birthyr"),
              which(col_names == "gender" ),
              which(col_names == "educ"),
              which(col_names == "race"),
              which(col_names == "hispanic"),
              which(col_names == "marstat"),
              which(col_names == "region"),
              which(col_names == "employ"),
              which(col_names == "hadjob"),
              which(col_names == "inputstate"))
outcome_cols <- c(which(startsWith(col_labels, "Abortion --")),
                 which(startsWith(col_labels, "Immigration --")),
                 which(startsWith(col_labels, "Taxes --")),
                 which(startsWith(col_labels, "Health Care --")),
                 which(startsWith(col_labels, "President 20")))
col_labels[demo_cols]
col_labels[outcome_cols]
keep_cols <- sort(c(demo_cols, outcome_cols))
cces_2018_subset <- subset(cces_2018[keep_cols])
```

We inspect the subset to make sure that we have captured the question and response names.
```{r}
str(cces_2018_subset, max.level=2, list.len=12, give.attr=FALSE)
```

The CCES data is roughly representative of the state-level political sentiment as a whole; but if we seek to understand certain subsets of the population,
the sample may well be unbalanced.
There are only 60,000 survey responses,
but there are over 20 million possible demographic subsets of interest,
given all possible combinations of gender, education,
race, marital status, employment, birthyear, and state of residence (51).

```{r}
cat(format((length(table(cces_2018_subset$gender[,drop=TRUE])) *
length(table(cces_2018_subset$educ[,drop=TRUE])) *
length(table(cces_2018_subset$race[,drop=TRUE])) *
length(table(cces_2018_subset$marstat[,drop=TRUE])) *
length(table(cces_2018_subset$employ[,drop=TRUE])) *
length(table(cces_2018_subset$birthyr[,drop=TRUE])) *
length(table(cces_2018_subset$inputstate[,drop=TRUE]))), big.mark=",",scientific=FALSE))
```

In order to correct this imbalance, we must poststratify to the general population.

### Population sample:  the 2018 American Community Survey

The US Census data is the largest and most representative population sample we have acces to.
A full census is conducted every 10 years.
The American Community Survey (ACS) is a micro-census conducted every year by the Census Bureau
over 1% of the households in the US, i.e., it has 3 million respondants (as opposed to 300 million in the full census).

The [Integrated Public Use Microdata Series (IPUMS)](https://ipums.org/what-is-ipums) (@ipums2020) is a service run by the University of Minnesota
which allows easy access to census and survey data from around the world.
IPUMS preserves and harmonizes US census microdata, including the American Community Survey.
To use this service, you must register at https://uma.pop.umn.edu/usa/user/new.
To get this data, we work through the IPUMS interface as follows:

* From the ipums.org website, select [IPUMS USA](https://uma.pop.umn.edu/usa/user/new) and then click on “Get Data”.

* Click the "SELECT SAMPLES" button and select the 2018 ACS survey, then click on "SUBMIT SAMPLE".

* Select the variables that will be included in our poststratification table.
The survey variables are categorized by HOUSEHOLD (household-level variables) and PERSON (individual-level variables).
   + On PERSON > DEMOGRAPHIC select SEX and AGE
   + On PERSON > RACE, ETHNICITY, AND NATIVITY select RACE, HISPAN, and CITIZEN
   + On PERSON > EDUCATION select EDUC
   + On HOUSEHOLD > GEOGRAPHIC select STATEFIP

* The selected variables are added to a "data cart", along with a set of preselected variables.
At "checkout", we uncheck the preselected variables, and then click on the "CREATE DATA EXTRACT" button.
The following screenshot shows our selections, outlined
in red.

```{r ipumsExtract, echo=FALSE, out.width = "75%", fig.cap="IPUMS requested ACS variables"}
library(knitr)
include_graphics("ipums_acs_2018_match_cces.png")
```

* As even the US Census microdata files are extremely large, the IPUMS service queues this request for offline processing.
When the data has been assembled, IPUMS sends an email with a link to the download page, shown below.

```{r ipumsDownload, echo=FALSE, out.width = "75%", fig.cap="IPUMS service download page"}
library(knitr)
include_graphics("ipums_download.png")
```

You must download the data file and the DDI file for the codebook, as well as the R command file.
The DDI file is an XML file which documents your data extract.  If you click on this link,
the DDI file displays in the browser; you must download this file as XML.
The IPUMS website provides [detailed instructions](https://usa.ipums.org/usa/extract_instructions.shtml).
Both the data file and xml download file have the same base name;
if you rename them, you must rename both of them accordingly.

The following script is based on the R command script provided by IPUMS,
which requires the [`ipumsr` package](https://cran.r-project.org/web/packages/ipumsr/vignettes/ipums.html).
This package provides methods to read, extract, inspect, and otherwise query the downloaded sample.
It also provides an excellent series of vignettes which describe how to work with the data.
We recommend starting here: http://tech.popdata.org/ipumsr/articles/ipums.html

```{r}
# NOTE: To load data, you must download both the extract's data and the DDI
# and also set the working directory to the folder with these files (or change the path below).

if (!require("ipumsr")) stop("Reading IPUMS data into R requires the ipumsr package. It can be installed using the following command: install.packages('ipumsr')")

# download dataset from IPUMS
# xml and dat file should have same name, be in same directory
acs_2018_subset_ddi <- read_ipums_ddi("usa_00007.xml")
acs_2018_subset <- read_ipums_micro(acs_2018_subset_ddi)
# save as R objects
save(acs_2018_subset, acs_2018_subset_ddi, file= "acs_2018_subset.RData")
```

#### Inspecting the ACS sample

The `ipumsr` package uses several [tidyverse](https://www.tidyverse.org/) packages
including [`dplyr`](https://dplyr.tidyverse.org/) and [`haven`](https://haven.tidyverse.org/).
The [`read_ipums_micro` function](http://tech.popdata.org/ipumsr/reference/read_ipums_micro.html)
returns a single [`tbl_df`](https://tibble.tidyverse.org/reference/tbl_df-class.html) data frame, i.e., a "tibble".
```{r}
library(ipumsr)
load("acs_2018_subset.RData")
str(acs_2018_subset, max.level=2, list.len=6)
dim(acs_2018_subset)
names(acs_2018_subset)
```
We selected variables "RACE" and "HISPAN"; the extract contains two columns for each:
"RACE" and "RACED", and "HISPAN" and "HISPAND".  The "D" columns provide an extremely fine-grained
set of categories for race and hispanic origin.  As these go far beyond the categories from
the CCES which are provided, we ignore them.

The columns of this tibble are  [labelled](https://haven.tidyverse.org/reference/labelled.html) vectors.
Labelled vectors can have text labels associated with specific values.
To see how this works, we compare the values and labels for `AGE`.
The 2018 ACS data spans ages 0 through 96,
of these, only ages 0 and 90 have labels.
```{r}
length(table(acs_2018_subset$AGE))
max(acs_2018_subset$AGE)
```

The function [`as_factor`](https://haven.tidyverse.org/reference/as_factor.html)
allows us to treat labelled vectors as factors.

### Aligning the Samples with MRP-KIT

We need to align our demographics of interest; we will do so column-by-column.

We note that information which must be aggregated from combinations of answers
given to a series of questions in order to match demographics in the ACS
is beyond the scope of what can be done with MRP-KIT, e.g.,
it doesn't provide functions for combining survey variables.

Also problematic: creating aggregate categories, e.g., binning age by decade.

#### Variable CCES "gender" / ACS "sex"

Both the CCES and ACS samples have variables for the sex of the respondant:
in the CCES this is named "gender" and in the ACS this is named "SEX".
In both, the response labels are "Male", "Female".
Creating a mapping between the two is straightforward:
we create a new SurveyQuestion object with name "sex",
then create a mappng from the CCES to the ACS,
both for column names and response labels
where for both mappings, the first element
is the survey sample name/label, and the second element
is the population sample name/label.

```{r}
library(mrpkit)
names(table(cces_2018_subset$gender))
print(ipums_val_labels(acs_2018_subset$SEX))
q_sex <- SurveyQuestion$new(name="sex",
      col_names=c("gender", "SEX"),
      values_map=list("Male" = "Male","Female" = "Female"))
print(q_sex)
```

#### Variable CCES "inputstate" / ACS "STATEFIP"

We examine the state names from both the CCES and ACS; the same labels are used in both surveys.
```{r}
# CCES data is a factor, levels are the labels
inputstate <- levels(cces_2018_subset$inputstate[,drop=TRUE])

# ACS data are structured objects, ipums_val_labels returns a tibble
STATEFIP <- ipums_val_labels(acs_2018_subset$STATEFIP)[[2]][1:51]

cat(inputstate[1:5], sep=", "); cat(STATEFIP[1:5], sep=", ")

table(inputstate == STATEFIP)
```

We need to create a SurveyQuestion object, as above,  but it would be
extremely tedious to write out a pairlist, e.g.:
```
values_map=list("Alabama" = "Alabama", ..., "Wyoming" = "Wyoming")
```
[R Pairlists](https://github.com/hadley/r-internals/blob/master/pairlists.md) are
a kind of linked list data structure where each element in the list has a name.
In order to create a pairlist programmatically, we use the R function `setNames`,
which takes as its first argument an object to be labelled
and as its second argument the corresponding character vector of labels,
i.e., the expression `as.list(c("value1", "value2"),  c("tag1", "tag2")`
is equivalent to the expression `list("tag1" = "value1", "tag2" = "value2").
```{r}
mapping = as.list(setNames(STATEFIP, inputstate))
mapping[1:3]
```

Having established the values map, we create the corresponding SurveyQuestion object.
We use accessor methods `name()`, `col_names()` and `values()` to check our work.
```{r}
q_state <- SurveyQuestion$new(name = "state",
	col_names = c("inputstate", "STATEFIP"),
	values_map= mapping)
q_state$name()
q_state$col_names()
head(q_state$values())
```

### 2021-06-29 case-study/prelim_data_analysis.R

@@ -1,0 +1,33 @@
+setwd("~/github/cces/dataverse")
+
+# data clean - bin race, age, inc, educ - ordered numeric categories
+
+# read 2018 data
+library(readstata13)
+cces_2018 = read.dta13("cces18_common_vv.dta", convert.factors = T, nonint.factors = T, generate.factors=T)
+
+# prelim data analysis
+table(complete.cases(cces_2018))
+
+# get demographics
+profile =c('birthyr','gender', 'educ', 'race', 'hispanic', 'marstat', 'employ', 'faminc_new', 'inputstate')
+profile_2018 = cces_2018[profile]
+
+library(tidyverse)
+categories_bins_plot <- function(df, column_name)
+{
+    df2 = as.data.frame(group_by_(df, column_name) %>% summarise(n = n()))
+    ggplot(df2, aes(x=df2[,1], y=df2[,2])) +
+        geom_bar(stat = "identity") +
+        xlab(column_name) + ylab(element_blank()) +
+        ggtitle(paste("CCES 2018", profile[i], "raw counts")) +
+        theme(axis.text.x = element_text(angle = 90, hjust = 1))
+}
+
+for(i in 1:length(profile)) {
+    name = paste("cces_2018_",profile[i],".png", sep="", collapse="")
+    p = categories_bins_plot(profile_2018, profile[i])
+    png(filename=name)
+    show(p)
+    dev.off()
+}

### 2021-06-29 case-study/process_cces_2018.R

@@ -1,0 +1,57 @@
+setwd("~/github/cces/dataverse")
+
+# read 2018 data
+#cces_2018 <- get(load("cces18_common_vv.RData"))
+#dim(cces_2018)
+
+library(readstata13)
+cces_2018 = read.dta13("cces18_common_vv.dta", convert.factors = T, nonint.factors = T, generate.factors=T)
+
+attr_names <- names(attributes(cces_2018))
+col_labels <- attr(cces_2018, "var.labels")
+col_names <- attr(cces_2018, "names")
+
+demo_cols <- c(1, which(col_names == "birthyr"),
+              which(col_names == "gender" ),
+              which(col_names == "educ"),
+              which(col_names == "race"),
+              which(col_names == "hispanic"),
+              which(col_names == "marstat"),
+              which(col_names == "region"),
+              which(col_names == "employ"),
+              which(col_names == "hadjob"),
+              which(col_names == "inputstate"))
+col_labels[demo_cols]
+
+outcome_cols <- c(which(startsWith(col_labels, "Abortion --")),
+                 which(startsWith(col_labels, "Immigration --")),
+                 which(startsWith(col_labels, "Taxes --")),
+                 which(startsWith(col_labels, "Health Care --")),
+                 which(startsWith(col_labels, "President 20")))
+col_labels[outcome_cols]
+
+keep_cols <- sort(c(demo_cols, outcome_cols))
+
+cces_2018_subset <- subset(cces_2018[keep_cols])
+attr(cces_2018_subset, "var.labels") <- col_labels[keep_cols]
+attr(cces_2018_subset, "names") <- col_names[keep_cols]
+
+table(cces_2018_subset$gender)
+table(cces_2018_subset$educ)
+table(cces_2018_subset$race)
+table(cces_2018_subset$marstat)
+table(cces_2018_subset$employ)
+table(cces_2018_subset$region)
+length(table(cces_2018_subset$inputstate))
+length(table(cces_2018_subset$birthyr))
+
+(length(table(cces_2018_subset$gender))-2) *
+(length(table(cces_2018_subset$educ))-2) *
+(length(table(cces_2018_subset$race))-2) *
+(length(table(cces_2018_subset$marstat))-2) *
+(length(table(cces_2018_subset$employ))-2) *
+(length(table(cces_2018_subset$region))-2) *
+length(table(cces_2018_subset$inputstate)) *
+length(table(cces_2018_subset$birthyr))
+
+save(cces_2018_subset, file="cces_2018_subset.RData")

### 2021-06-29 case-study/process_ipums_acs_2018.R

@@ -1,0 +1,20 @@
+# NOTE: To load data, you must download both the extract's data and the DDI
+# and also set the working directory to the folder with these files (or change the path below).
+
+if (!require("ipumsr")) stop("Reading IPUMS data into R requires the ipumsr package. It can be installed using the following command: install.packages('ipumsr')")
+
+# download dataset from IPUMS
+# xml and dat file should have same name, be in same directory
+acs_2018_ddi <- read_ipums_ddi("usa_00007.xml")
+acs_2018 <- read_ipums_micro(acs_2018_ddi)
+
+# save object for later
+save(acs_2018, acs_2018_ddi, file= "acs_2018_subset.RData")
+
+acs_colnames <- names(acs_2018)
+print(acs_colnames)
+print(class(acs_colnames[1]))
+for (i in 1:length(acs_colnames)) {
+    print(acs_colnames[i])
+    print(ipums_val_labels(acs_2018[[i]]))
+}

### 2021-06-30 case-study/Munging_CCES.Rmd
checkpointing case study

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(cache=TRUE, message=FALSE, error=FALSE, warning=FALSE, comment=NA, out.width='95%', collapse=TRUE)
+knitr::opts_chunk$set(cache=TRUE, message=FALSE, error=FALSE, warning=FALSE, comment=NA, out.width='95%', collapse=TRUE, prompt=TRUE)
 ```
@@
 <style type="text/css">
-.table { width: 40%; }
+.table { width: 85%; }
 div.quote-container blockquote {  font-size: small; }
@@
 
+Multi-level Regresssion and Postrativication (MRP)
+is a two-step process for correcting model estimates
+obtained from a survey sample when the survey population is not
+the same as the target population over which we wish to do inference.
+The first step is the doing a multilevel regression on the survey sample;
+the second step is poststratification to the target population.
+Conceptually, MRP is relatively simple.
+In practice, it requires more than twice as much data processing.
+In addition to managing the survey data, you must also obtain
+and manage the target population data, in addition you must
+develop a mapping between information in these two datasets.
+In order to produce verifiable, replicable results, it is
+critical that this mapping be documented and scripted.
+
 The goal of this case study is to show how the MRP-KIT package
-can facilitate analysis of a large-scale survey sample.
-By large-scale, we mean surveys which have both a rich set
-of demographics and a large number of responses.
-The richer the set of demographics, the larger the number
-of possible sub-populations of interest; no matter how large
-or representative the survey, the combinatorics of the demographic
-categories are such that it is unlikely that the sample
-is representative for all subpopulations of interest.
-In order to correct this imbalance,
-we post-stratify to a truly representative
-sample drawn from the general population.
-This, in turn, requires obtaining a general population sample;
-one which is at least as large and which contains all the
-demographics of interest.
+can facilitate this process for large-scale surveys.
+By large-scale surveys, we mean surveys which have both
+a large number of respondants and detailed demographic profiles for each respondant.
+The larger the number of demographics, the larger the number
+of possible sub-populations of interest and the more difficult it becomes
+to collect a representative sample for all such subpopulations,
+thus even well-designed surveys can require use of MRP.
 
@@
 of interest.
-The population sample is the 2018 American Community Survey, the
+The population sample is the 2018 American Community Survey (ACS), the
 annual micro-census conducted by the US Census Bureau.  It is designed
@@
 https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi%3A10.7910/DVN/ZSBZ7K
-The raw survey data is distributed in several from the Harvard Dataverse
+The raw survey data is distributed in several formats from the Harvard Dataverse
 as  [cces18_common_vv](https://dataverse.harvard.edu/file.xhtml?persistentId=doi:10.7910/DVN/ZSBZ7K/C1YVSF&version=6.0).
-Although the tabular versions of the data file are smaller, we download the
-Stata file, as it contains metadata on all the questions; this is critical to
-being able to align the CCES sample to the US Census population sample.
-We also download the [2018 CCES Guide](https://dataverse.harvard.edu/file.xhtml?persistentId=doi:10.7910/DVN/ZSBZ7K/WZWCZ1&version=6.0)
-which describes the survey questions, labels, and responses.
+While the tabular versions of the data file are smaller, the Stata version `cces18_common_vv.dta`
+contains metadata on all the questions: the question name, label, description, response labels and values.
+The corresponding 
+[2018 CCES Guide](https://dataverse.harvard.edu/file.xhtml?persistentId=doi:10.7910/DVN/ZSBZ7K/WZWCZ1&version=6.0)
+provides a human-readable version of this metadata as well.
 
-We load the data into R using the [`readstata13`](https://sjewo.github.io/readstata13/) library function
-[`read.dta13`](https://sjewo.github.io/readstata13/reference/read.dta13.html) which returns a data.frame with attributes.
+```{r ccesDownload, echo=FALSE, out.width = "75%", fig.cap="Harvard Dataverse download page"}
+library(knitr)
+include_graphics("dataverse_cces_2018_download.png")
+```
+
+We download the Stata file because we will need the metatdata to map the CCES sample to the ACS using MRP-KIT,
+then use the [`readstata13`](https://sjewo.github.io/readstata13/) library function
+[`read.dta13`](https://sjewo.github.io/readstata13/reference/read.dta13.html) to load the Stata file into R.
+The result is an R data.frame with attributes containing the metadata.
 ```{r}
@@
 The demographic variables include
-gender, education level, employment status, race, and geographic location.
+gender, education level, race, and geographic location, among others.
 ```{r}
@@
               which(col_names == "marstat"),
-              which(col_names == "region"),
-              which(col_names == "employ"),
-              which(col_names == "hadjob"),
               which(col_names == "inputstate"))
@@
 
-The CCES data is roughly representative of the state-level political sentiment as a whole; but if we seek to understand certain subsets of the population,
+The CCES data is roughly representative of the state-level political sentiment as a whole,
+but if we seek to understand certain subsets of the population,
 the sample may well be unbalanced.
 There are only 60,000 survey responses,
-but there are over 20 million possible demographic subsets of interest,
-given all possible combinations of gender, education,
-race, marital status, employment, birthyear, and state of residence (51).
-
-```{r}
-cat(format((length(table(cces_2018_subset$gender[,drop=TRUE])) *
-length(table(cces_2018_subset$educ[,drop=TRUE])) *
-length(table(cces_2018_subset$race[,drop=TRUE])) *
-length(table(cces_2018_subset$marstat[,drop=TRUE])) *
-length(table(cces_2018_subset$employ[,drop=TRUE])) *
-length(table(cces_2018_subset$birthyr[,drop=TRUE])) *
-length(table(cces_2018_subset$inputstate[,drop=TRUE]))), big.mark=",",scientific=FALSE))
-```
-
-In order to correct this imbalance, we must poststratify to the general population.
+but there are over 2 million
+possible combinations of gender, education,
+race, marital status, birthyear, and state of residence.
+If we bin ages by decade, there are a mere 200,000 combinations,
+binning states by region further reduces this to 16,000, etc., etc.
+However, if we wish to consider state-level subpopulations of interest,
+the survey data is likely to be sparse and not representative of the general population.
+To address the former we do multi-level modeling; for the latter, poststratification.
 
@@
 
-The US Census data is the largest and most representative population sample we have acces to.
+The US Census data is the largest representative population sample that we have access to.
 A full census is conducted every 10 years.
@@
 
-* As even the US Census microdata files are extremely large, the IPUMS service queues this request for offline processing.
+The IPUMS service queues this request for offline processing.
 When the data has been assembled, IPUMS sends an email with a link to the download page, shown below.
@@
 You must download the data file and the DDI file for the codebook, as well as the R command file.
+The IPUMS website provides [detailed instructions](https://usa.ipums.org/usa/extract_instructions.shtml) on how to do this.
 The DDI file is an XML file which documents your data extract.  If you click on this link,
 the DDI file displays in the browser; you must download this file as XML.
-The IPUMS website provides [detailed instructions](https://usa.ipums.org/usa/extract_instructions.shtml).
 Both the data file and xml download file have the same base name;
@@
 of these, only ages 0 and 90 have labels.
+The function [`ipums_val_labels`](https://tech.popdata.org/ipumsr/reference/ipums_var_info.html)
+allows us to inspect these labels.
 ```{r}
 length(table(acs_2018_subset$AGE))
-max(acs_2018_subset$AGE)
+ipums_val_labels(acs_2018_subset$AGE)
 ```
@@
 allows us to treat labelled vectors as factors.
+See the vignette https://tech.popdata.org/ipumsr/reference/ipums_var_info.html for more details.
 
@@
 
-We need to align our demographics of interest; we will do so column-by-column.
+We use MRP-KIT to map demographics of interest on a column-by-column basis.^[
+If information is spread across multiple columns in either the survey sample
+or the target population sample, it must be combined into a single column's
+worth of information as MRP-KIT doesn't provide functions to do this.
+*Note: all pre-processing should be documented, scripted, and **tested**!*]
 
-We note that information which must be aggregated from combinations of answers
-given to a series of questions in order to match demographics in the ACS
-is beyond the scope of what can be done with MRP-KIT, e.g.,
-it doesn't provide functions for combining survey variables.
+In the following sections we examine each variable in turn.
+This is a tedious process, therefore, we provide the following
+summary of the demographic construct and the kinds of challenge each mapping presents
 
-Also problematic: creating aggregate categories, e.g., binning age by decade.
+Construct | CCES | ACS | Challenge
+--------- | ---- | --- | ---------
+Sex | gender | SEX | (none)
+State | inputstate | STATEFIP | specify 1-1 mapping programatically
+Marital status | marstat | MARST | choice of mapping
+Education | educ | EDUC | choice of mapping
+Race/Hispanic ethnicity | race, hispan | RACE, HISPAN | cannot align CCES race to ACS RACE
+Age | birthyr | AGE | different units, too fine-grained
 
-#### Variable CCES "gender" / ACS "sex"
+#### Trivial mapping: respondant sex, 2 categories
 
-Both the CCES and ACS samples have variables for the sex of the respondant:
+The CCES and ACS samples record the sex of the respondant;
 in the CCES this is named "gender" and in the ACS this is named "SEX".
@@
 names(table(cces_2018_subset$gender))
-print(ipums_val_labels(acs_2018_subset$SEX))
+ipums_val_labels(acs_2018_subset$SEX)
 q_sex <- SurveyQuestion$new(name="sex",
@@
       values_map=list("Male" = "Male","Female" = "Female"))
-print(q_sex)
+q_sex
 ```
 
-#### Variable CCES "inputstate" / ACS "STATEFIP"
+#### 1-1 mapping on a long list: state of residence, 51 categories
 
@@
 ```
+
+#### Choice of mapping: education, marital status
+
+The CCES and ACS samples record the education level of the respondant
+in variables `educ` and `EDUC`, respectively.
+Both the CCES and ACS record the highest level of education completed,
+but the CCES records the degree obtained whereas the ACS records the year completed.
+```{r}
+names(table(cces_2018_subset$educ))
+ipums_val_labels(acs_2018_subset$EDUC)
+```
+We choose to align these categories as follows:
+
+CCES | ACS
+---- | ---
+"No HS" | "Nursery school to grade 4"
+"No HS" | "Grade 5, 6, 7, or 8"
+"No HS" | "Grade 9"
+"No HS" | "Grade 10"
+"No HS" | "Grade 11"
+"High School Graduate" | "Grade 12"
+"Some College" | "1 year of college"
+"2-year" | "2 years of college"
+"2-year" | "3 years of college"
+"4-year" | "4 years of college"
+"Post-grad" | "5+ years of college"
+"skipped" | "N/A or no schooling"
+"not asked" | "N/A or no schooling"
+
+Alternative mappings are equally viable.
+With MRP-KIT, the choice of mappings is recorded and
+easily accessible.
+```{r}
+educ_mapping = list(
+    "No HS" = "Nursery school to grade 4",
+    "No HS" = "Grade 5, 6, 7, or 8",
+    "No HS" = "Grade 9",
+    "No HS" = "Grade 10",
+    "No HS" = "Grade 11",
+    "High School Graduate" = "Grade 12",
+    "Some College" = "1 year of college",
+    "2-year" = "2 years of college",
+    "2-year" = "3 years of college",
+    "4-year" = "4 years of college",
+    "Post-grad" = "5+ years of college",
+    "skipped" = "N/A or no schooling",
+    "not asked" = "N/A or no schooling"
+)
+q_educ <- SurveyQuestion$new(name = "educ",
+    col_names = c("educ", "EDUC"),
+    values_map= educ_mapping)
+q_educ$name()
+q_educ$col_names()
+head(q_educ$values())
+```
+
+As with eduction, the CCES and ACS samples record the marital status
+of the respondant differently and we have to use our judgement as to how best
+to align the two.
+```{r}
+names(table(cces_2018_subset$marstat))
+ipums_val_labels(acs_2018_subset$MARST)
+```
+
+CCES | ACS
+---- | ---
+"Married" | "Married, spouse present"
+"Married" | "Married, spouse absent"
+"Separated" | "Separated"
+"Divorced" | "Divorced"
+"Widowed" | "Widowed"
+"Never married" | "Never married"
+"Domestic / civil partnership" | "Never married"
+"skipped" | "Never married"
+"not asked" | "Never married"
+
+We note that for the 2018 CCES survey, the marital status of all respondants is known; nonetheless,
+we provide a complete mapping.
+
+```{r}
+marstat_mapping = list(
+    "Married" = "Married, spouse present",
+    "Married" = "Married, spouse absent",
+    "Separated" = "Separated",
+    "Divorced" = "Divorced",
+    "Widowed" = "Widowed",
+    "Never married" = "Never married",
+    "Domestic / civil partnership" = "Never married",
+    "skipped" = "Never married",
+    "not asked" = "Never married"
+)
+q_marstat <- SurveyQuestion$new(name = "marstat",
+    col_names = c("marstat", "MARST"),
+    values_map= marstat_mapping)
+q_marstat$name()
+q_marstat$col_names()
+head(q_marstat$values())
+```
+
+#### Information split across survey columns:  race and ethnicity.
+
+We ignore the politics of race and racial categories
+and instead focus on how best to map the CCES information to the ACS information.
+Both surveys have multiple variables which provide information about race and ethnicity,
+and both record race and hispanic origin separately.
+However the CCES treats hispanic origin as race as well as ethnic origin; the ACS does not.
+```{r}
+names(table(cces_2018_subset$race))
+names(table(cces_2018_subset$hispan))
+ipums_val_labels(acs_2018_subset$RACE)
+ipums_val_labels(acs_2018_subset$HISPAN)
+```
+In this case, munging the raw survey data is unavoidable.
+For each survey we create a new variable which blends race and hispanic ethnicity.
+We are scripting everyting, so we add this to our processing scripts.

### 2021-06-30 case-study/prelim_data_analysis.R

@@ -1,33 +1,0 @@
-setwd("~/github/cces/dataverse")
-
-# data clean - bin race, age, inc, educ - ordered numeric categories
-
-# read 2018 data
-library(readstata13)
-cces_2018 = read.dta13("cces18_common_vv.dta", convert.factors = T, nonint.factors = T, generate.factors=T)
-
-# prelim data analysis
-table(complete.cases(cces_2018))
-
-# get demographics
-profile =c('birthyr','gender', 'educ', 'race', 'hispanic', 'marstat', 'employ', 'faminc_new', 'inputstate')
-profile_2018 = cces_2018[profile]
-
-library(tidyverse)
-categories_bins_plot <- function(df, column_name)
-{
-    df2 = as.data.frame(group_by_(df, column_name) %>% summarise(n = n()))
-    ggplot(df2, aes(x=df2[,1], y=df2[,2])) +
-        geom_bar(stat = "identity") +
-        xlab(column_name) + ylab(element_blank()) +
-        ggtitle(paste("CCES 2018", profile[i], "raw counts")) +
-        theme(axis.text.x = element_text(angle = 90, hjust = 1))
-}
-
-for(i in 1:length(profile)) {
-    name = paste("cces_2018_",profile[i],".png", sep="", collapse="")
-    p = categories_bins_plot(profile_2018, profile[i])
-    png(filename=name)
-    show(p)
-    dev.off()
-}

### 2021-06-30 case-study/process_cces_2018.R

@@ -1,9 +1,3 @@
-setwd("~/github/cces/dataverse")
-
-# read 2018 data
-#cces_2018 <- get(load("cces18_common_vv.RData"))
-#dim(cces_2018)
-
 library(readstata13)
 cces_2018 = read.dta13("cces18_common_vv.dta", convert.factors = T, nonint.factors = T, generate.factors=T)
 
@@ -17,9 +11,9 @@ demo_cols <- c(1, which(col_names == "birthyr"),
               which(col_names == "race"),
               which(col_names == "hispanic"),
               which(col_names == "marstat"),
-              which(col_names == "region"),
-              which(col_names == "employ"),
-              which(col_names == "hadjob"),
+#              which(col_names == "region"),
+#              which(col_names == "employ"),
+#              which(col_names == "hadjob"),
               which(col_names == "inputstate"))
 col_labels[demo_cols]
 
@@ -40,8 +34,6 @@ table(cces_2018_subset$gender)
 table(cces_2018_subset$educ)
 table(cces_2018_subset$race)
 table(cces_2018_subset$marstat)
-table(cces_2018_subset$employ)
-table(cces_2018_subset$region)
 length(table(cces_2018_subset$inputstate))
 length(table(cces_2018_subset$birthyr))
 
@@ -49,9 +41,29 @@ length(table(cces_2018_subset$birthyr))
 (length(table(cces_2018_subset$educ))-2) *
 (length(table(cces_2018_subset$race))-2) *
 (length(table(cces_2018_subset$marstat))-2) *
-(length(table(cces_2018_subset$employ))-2) *
-(length(table(cces_2018_subset$region))-2) *
 length(table(cces_2018_subset$inputstate)) *
 length(table(cces_2018_subset$birthyr))
 
 save(cces_2018_subset, file="cces_2018_subset.RData")
+
+## # viz demographics
+## demographics_2018 = cces_2018[demo_cols]
+## demo_names = names(demographics_2018)
+## library(tidyverse)
+## categories_bins_plot <- function(df, column_name)
+## {
+##     df2 = as.data.frame(group_by_(df, column_name) %>% summarise(n = n()))
+##     ggplot(df2, aes(x=df2[,1], y=df2[,2])) +
+##         geom_bar(stat = "identity") +
+##         xlab(column_name) + ylab(element_blank()) +
+##         ggtitle(paste("CCES 2018", column_name, "raw counts")) +
[…]

### 2021-07-02 case-study/Munging_CCES.Rmd
checkpointing

@@
 knitr::opts_chunk$set(cache=TRUE, message=FALSE, error=FALSE, warning=FALSE, comment=NA, out.width='95%', collapse=TRUE, prompt=TRUE)
+options(digits=2)
+if (!require("ipumsr")) stop("This case study requires ipumsr package. It can be installed using the following command: install.packages('ipumsr')")
+if (!require("labelled")) stop("This case study requires labelled package. It can be installed using the following command: install.packages('labelled')")
+library(ipumsr)
+library(labelled)
 ```
@@
 div.quote-container blockquote {  font-size: small; }
+blockquote {  font-size: small; }
 </style>
@@
 In addition to managing the survey data, you must also obtain
-and manage the target population data, in addition you must
-develop a mapping between information in these two datasets.
+and manage the target population data.
+Most importantly, you must develop a mapping between information
+in these two datasets.
 In order to produce verifiable, replicable results, it is
@@
 
-In this case study, the survey sample is one year's worth of data
-from is Cooperative Congressional Election Study (CCES),
+In this case study, the survey data comes from
+the 2018 Cooperative Congressional Election Study (CCES),
 a bi-annual survey of of 60,000 American adults from
@@
 and congressional (and presidential) elections.
-While this survey sample is representative at the national level
-it is not large enough to provide a representative sample for sub-populations
-of interest.
-The population sample is the 2018 American Community Survey (ACS), the
-annual micro-census conducted by the US Census Bureau.  It is designed
+The population data comes from the 2018 American Community Survey (ACS),
+the annual micro-census conducted by the US Census Bureau.  It is designed
 to be a representative sample of the US population as a whole; data is
@@
 A secondary goal of this case study is showing how to work effectively
-and efficiently with large samples to produce replicable results.
+and efficiently with large surveys to produce replicable results.
 
-### Survey sample: the 2018 CCES Common Content Dataset
+### Survey data: the 2018 CCES Common Content Dataset
+
+The zero-th step in any data analysis is obtaining the data and checking its contents.
+When working with your own data, you have already done this.
+When working with someone else's data, you need to understand
+both the contents of the dataset and its structure.
 
@@
 The corresponding 
-[2018 CCES Guide](https://dataverse.harvard.edu/file.xhtml?persistentId=doi:10.7910/DVN/ZSBZ7K/WZWCZ1&version=6.0)
+[2018 CCES Guide](https://sda.berkeley.edu/sdaweb/docs/cces2018/DOC/CCES+Guide+2018.pdf)
 provides a human-readable version of this metadata as well.
@@
 We download the Stata file because we will need the metatdata to map the CCES sample to the ACS using MRP-KIT,
-then use the [`readstata13`](https://sjewo.github.io/readstata13/) library function
-[`read.dta13`](https://sjewo.github.io/readstata13/reference/read.dta13.html) to load the Stata file into R.
+then use the package [`readstata13`](https://www.rdocumentation.org/packages/readstata13/versions/0.10.0)
+function [`read.dta13`](https://www.rdocumentation.org/packages/readstata13/versions/0.10.0/topics/read.dta13)
+to load the Stata file into R.
 The result is an R data.frame with attributes containing the metadata.
@@
 names(cces_2018)[1:12]
+names(attributes(cces_2018))
+```
+
+The `read.dta13` creates attributes corresponding to the Stata file metadata.
+The attribute `names` corresponds to the dataframe column names and
+the attribute `val.labels` provides a short description of that columns.
+For questions which have categorical responses, the attribute `data.label`
+provides the category name and interger code used for that survey question.
+```{r}
+names(attributes(cces_2018))
+attr(cces_2018, "names")[1:10]
+attr(cces_2018, "var.labels")[1:10]
+attr(cces_2018, "label.table")[["gender"]]
 ```
@@
 ```
-The political opinion questions on the survey are labelled "Abortion", "Immigration", "Taxes", "Health Care".
+
+The political opinion questions on the survey are labelled "President 2016", "Abortion", "Immigration", "Taxes", "Health Care".
 ```{r}
-attr(cces_2018, "var.labels")[which(startsWith(attr(cces_2018, "var.labels"), "Health Care"))[1]]
-health_care_1 <- which(startsWith(attr(cces_2018, "var.labels"), "Health Care"))[1]
-table(cces_2018[[health_care_1]][, drop=TRUE])
+attr(cces_2018, "var.labels")[which(startsWith(attr(cces_2018, "var.labels"), "President 20"))]
+pres_vote <- which(startsWith(attr(cces_2018, "var.labels"), "President 20"))
+table(cces_2018[[pres_vote]][,drop=TRUE], useNA="always")
+
+attr(cces_2018, "var.labels")[which(startsWith(attr(cces_2018, "var.labels"), "Health Care --"))[1]]
+health_care_Q1 <- which(startsWith(attr(cces_2018, "var.labels"), "Health Care --"))[1]
+table(cces_2018[[health_care_Q1]][,drop=TRUE], useNA="always")
 ```
@@
 Of the 525 columns in the survey, we are interested in the set of political opinion questons
-and a subset of the demographics.  Although not strictly necessary, we create a new dataframe
+and a subset of the demographics.  We create a new dataframe
 which contains just these columns of interest.
-The survey metadata includes column names, and question and response labels.
-We use the column names to select demographics and column labels to select political opinion questions.
+Using the survey metadata, we select demographics by column names and political opinion questions by the column labels.
+In order to properly postratify, we remove any record where there is missing demographic information.
+In addition, so that we can easily align the CCES variables with the ACS variables
+we removed unused category labels.
+
 ```{r}
@@
 col_names <- attr(cces_2018, "names")
-demo_cols <- c(1, which(col_names == "birthyr"),
+demo_cols <- c(which(col_names == "birthyr"),
               which(col_names == "gender" ),
@@
               which(col_names == "inputstate"))
+# don't allow missing demographics
+keep_rows <- complete.cases(cces_2018[, demo_cols])
+
 outcome_cols <- c(which(startsWith(col_labels, "Abortion --")),
@@
                  which(startsWith(col_labels, "President 20")))
-col_labels[demo_cols]
-col_labels[outcome_cols]
 keep_cols <- sort(c(demo_cols, outcome_cols))
-cces_2018_subset <- subset(cces_2018[keep_cols])
+
+cces_2018_subset <- subset(cces_2018, keep_rows, keep_cols)
+# `subset` drops attributes, keep attribute "var.labels"
+attr(cces_2018_subset, "var.labels") <- attr(cces_2018, "var.labels")
+
+# cleanup factor variables; drop unused category labels
+for (name in names(cces_2018_subset)) {
+    if (is.factor(cces_2018_subset[[name]]))  {
+        cces_2018_subset[[name]] = droplevels(cces_2018_subset[[name]])
+    }
+}
 ```
@@
 
-The CCES data is roughly representative of the state-level political sentiment as a whole,
-but if we seek to understand certain subsets of the population,
-the sample may well be unbalanced.
-There are only 60,000 survey responses,
-but there are over 2 million
+The CCES has 60,000 respondants.
+According to the [Guide to the 2018 Cooperative Congressional Election Survey](https://sda.berkeley.edu/sdaweb/docs/cces2018/DOC/CCES+Guide+2018.pdf)
+
+>This study constructed a very large sample capable of capturing variation across a wide variety of legislative constituencies.
+In fact, the state-level samples are sufficiently large as to measure with a reasonable degree of precision the distribution of voters’ preferences within most states.
+
+While the CCES data may be roughly representative of the state-level political sentiment as a whole,
+if we seek to understand certain subsets of the population,
+it is unlikely to be representative of those groups.
+There are over 2 million
 possible combinations of gender, education,
@@
 
-### Population sample:  the 2018 American Community Survey
+### Population data:  the 2018 American Community Survey
 
-The US Census data is the largest representative population sample that we have access to.
+The US Census data is the largest representative survey of the US population that we have access to.
 A full census is conducted every 10 years.
@@
 ```{r}
+library(labelled)
 library(ipumsr)
@@
 ipums_val_labels(acs_2018_subset$AGE)
+ipums_val_labels(acs_2018_subset$AGE)
 ```
@@
 allows us to treat labelled vectors as factors.
-See the vignette https://tech.popdata.org/ipumsr/reference/ipums_var_info.html for more details.
+See the vignette https://tech.popdata.org/ipumsr/reference/ipums_var_info.html for further information.
+```{r}
+head(table(acs_2018_subset$AGE))
+head(table(as_factor(acs_2018_subset$AGE)))
+```
 
-### Aligning the Samples with MRP-KIT
+### Aligning the survey and target population data with MRP-KIT
 
-We use MRP-KIT to map demographics of interest on a column-by-column basis.^[
-If information is spread across multiple columns in either the survey sample
+We use MRP-KIT to allow us to do inference on subpopulations of interest
+using the following demographics:
+sex, state of residence, education, race, marital status, and age.
+
+MRP-KIT allows mappings that are one-to-one, many-to-one, one-to-many.
+All mappings must be completely deterministic based on category label alone.
+If this is not possible then you must create new, synthetic variables which can be
+cleanly aligned.  Likewise, 
+when information is spread across multiple columns in either the survey sample
 or the target population sample, it must be combined into a single column's
-worth of information as MRP-KIT doesn't provide functions to do this.
-*Note: all pre-processing should be documented, scripted, and **tested**!*]
+worth of information.
+If a demographic is recorded at too fine a level of granularity
+and should be aggregated, e.g., binning date years into decades,
+this much also be done beforehand.
+MRP-KIT doesn't and cannot provide functions to do this; only you know
+the data and what it means.
+All such data pre-processing should be documented, srcipted, and *tested*
+to insure replicability and/or use in production pipelines.
 
-In the following sections we examine each variable in turn.
-This is a tedious process, therefore, we provide the following
-summary of the demographic construct and the kinds of challenge each mapping presents
+The MRP-KIT class `SurveyQuestion` is used to map the survey data demographic
+information to the corresponding population data demographics.
+The mapping process is both tedious and challenging.
+The following table provides a summary of the demographic constructs
+that we wish to use in analyzing the CCES dataset,
+and challenges involved in mapping the CCES to the ACS.
 
@@
 State | inputstate | STATEFIP | specify 1-1 mapping programatically
-Marital status | marstat | MARST | choice of mapping
-Education | educ | EDUC | choice of mapping
-Race/Hispanic ethnicity | race, hispan | RACE, HISPAN | cannot align CCES race to ACS RACE
+Education | educ | EDUC | different possible mappings
+Race/Hispanic ethnicity | race, hispan | RACE, HISPAN | information spread across categories
+Marital status | marstat | MARST | many-to-many mapping
 Age | birthyr | AGE | different units, too fine-grained
 
-#### Trivial mapping: respondant sex, 2 categories
+#### Simple one-to-one mapping:  CCES "gender" / ACS "sex"
 
-The CCES and ACS samples record the sex of the respondant;
-in the CCES this is named "gender" and in the ACS this is named "SEX".
-In both, the response labels are "Male", "Female".
-Creating a mapping between the two is straightforward:
-we create a new SurveyQuestion object with name "sex",
+We create a new SurveyQuestion object with name "sex",
 then create a mappng from the CCES to the ACS,
@@
 # CCES data is a factor, levels are the labels
-inputstate <- levels(cces_2018_subset$inputstate[,drop=TRUE])
+inputstate <- levels(cces_2018_subset$inputstate)
 
@@
 We need to create a SurveyQuestion object, as above,  but it would be
-extremely tedious to write out a pairlist, e.g.:
+extremely tedious to do this as a long list of tag, value pairs:
 ```
@@
 ```
-[R Pairlists](https://github.com/hadley/r-internals/blob/master/pairlists.md) are
-a kind of linked list data structure where each element in the list has a name.
-In order to create a pairlist programmatically, we use the R function `setNames`,
-which takes as its first argument an object to be labelled
-and as its second argument the corresponding character vector of labels,
-i.e., the expression `as.list(c("value1", "value2"),  c("tag1", "tag2")`
-is equivalent to the expression `list("tag1" = "value1", "tag2" = "value2").
+The `values_map` is stored as an [R Pairlists](https://github.com/hadley/r-internals/blob/master/pairlists.md).
+This is actually a labelled list, not a list of pairs, i.e.,
+the expression `list("tag1" = "value1", "tag2" = "value2")`
+creates a list of length 2 where each element is a single named item.
 ```{r}
-mapping = as.list(setNames(STATEFIP, inputstate))
+foo <- list("tag1" = "value1", "tag2" = "value2")
+foo
+```
+To create this list programmatically,
+you must create a list of values and a list of labels,
+then set the names of the values list to the labels.
+```{r}
+# create list, then set names (uppercase)
+mapping <- as.list(STATEFIP)
+names(mapping) <- toupper(inputstate)
+mapping[1:3]
+
+# the one-liner (lowercase)
+
+mapping = as.list(setNames(STATEFIP, tolower(inputstate)))
 mapping[1:3]
@@
 ```{r}
+mapping = as.list(setNames(STATEFIP, inputstate))
 q_state <- SurveyQuestion$new(name = "state",
@@
 
-#### Choice of mapping: education, marital status
+#### Choice of mapping: education
 
@@
 ```
-We choose to align these categories as follows:
+We choose to align these categories as follows.
 
@@
 ---- | ---
+"No HS" | "N/A or no schooling"
 "No HS" | "Nursery school to grade 4"
@@
 "Post-grad" | "5+ years of college"
-"skipped" | "N/A or no schooling"
-"not asked" | "N/A or no schooling"
 
-Alternative mappings are equally viable.
-With MRP-KIT, the choice of mappings is recorded and
-easily accessible.
+Mpping "Some College" to both 1 year and 3 years of college is also possible,
+reasoning that the goal of a 4-year college degree wasn't obtained.
+The choice depends on both an understanding of the survey data and the particular goals
+of your analysis.
+MRP-KIT can only provide the mechanism for recording and carrying out the mapping.
 ```{r}
 educ_mapping = list(
+    "No HS" = "N/A or no schooling",
     "No HS" = "Nursery school to grade 4",
@@
     "4-year" = "4 years of college",
-    "Post-grad" = "5+ years of college",
-    "skipped" = "N/A or no schooling",
-    "not asked" = "N/A or no schooling"
+    "Post-grad" = "5+ years of college"
 )
@@
 q_educ$col_names()
-head(q_educ$values())
+q_educ$values()
 ```
 
-As with eduction, the CCES and ACS samples record the marital status
-of the respondant differently and we have to use our judgement as to how best
-to align the two.
+#### Survey variable cannot be mapped to single census variable:  Race
+
+The [IPUMS description of the ACS variable RACE](https://usa.ipums.org/usa-action/variables/RACE#description_section)
+says:
+
+>Currently, the Census Bureau and others consider race to be a sociopolitical construct,
+not a scientific or anthropological one.  Many detailed RACE categories consist of national origin groups.
+
+This is borne out by the CCES, where the question for variable "race" is
+"What racial or ethnic group best describes you?",
+and includes "Hispanic" as a possible response.
+The ACS variable "RACE" has a different set of responses:
+```{r}
+names(table(cces_2018_subset$race))
+ipums_val_labels(acs_2018_subset$RACE)
+````
+
+In order to use the ACS for poststratification, we need to create a variable
+corresponding to the CCES variable "race" from a combination of ACS variables.
+Both surveys contain variable "Hispanic Origin".
+For the CCES, this is a "Yes/No" question:
+"Are you of Spanish, Latino, or Hispanic origin or descent?".
+In the ACS, the variable HISPAN identifies persons of Hispanic/Spanish/Latino origin and
+classifies them according to their country of origin.
+```{r}
+names(table(cces_2018_subset$hispan))
+ipums_val_labels(acs_2018_subset$HISPAN)
+```
+
+We propose to add a column to the ACS survey which combines information present in the `RACE` and `HISPAN` variables.
+```{r}
+# create column which blends "race" "hispan"
+acs_race2 <- as_factor(acs_2018_subset$RACE)
+
+# population percentages by label
+prop.table(table(acs_race2))
+
+# if "hispan" is "Yes", "race" label is "Hispanic".
+is_hispan <- acs_2018_subset$HISPAN > 0 & acs_2018_subset$HISPAN < 5
+length(is_hispan[is_hispan == TRUE]) / length(is_hispan)
+levels(acs_race2) = c(levels(acs_race2), "Hispanic")
+acs_race2[is_hispan==TRUE] <- "Hispanic"
+
+# adjusted percentages
+prop.table(table(acs_race2))
+
+# add to acs_2018_subset data.frame - respect column types
+acs_2018_subset$race2 = to_labelled(acs_race2)
+```
+
+A further possibility is to similarly combine the CCES variables "race" and "hispan" 
+so that the question of "Hispanic origin" is "Yes", the race-hispan label is "Hispanic".
+Were we to do this, we would create a new variable "race2" and add it to the
+CCES subset so that we could evaluate both classifications.
+```{r}
+# Population percentages in CCES variable "race"
+prop.table(table(cces_2018_subset$race))
+
+# Intersection (and) vs. union (or) of race == "Hispanic" and hispanic == "Yes"
+length(which(cces_2018_subset$race == "Hispanic" & cces_2018_subset$hispan == "Yes"))
+length(which(cces_2018_subset$race == "Hispanic" | cces_2018_subset$hispan == "Yes"))
+
+# create column which adjusts race labels
+cces_race2 <- cces_2018_subset$race
+cces_race2[cces_2018_subset$hispan == "Yes"] <- "Hispanic"
+prop.table(table(cces_race2))
+
+# add to acs_2018_subset data.frame - respect column types
+cces_2018_subset$race2 = cces_race2
+```
+
+Once comparable variables are present in both surveys,
+MRP-KIT can handle the mapping.
+We choose to align CCES "race" with new ACS variable "race2" as follows
+
+CCES "race" | ACS "race2"
+----------- | -----------
+"White" | "White"
+"Black" | "Black/African American/Negro"
+"Native American" | "American Indian or Alaska Native"
+"Asian" | "Chinese"
+"Asian" | "Japanese"
+"Asian" | "Other Asian or Pacific Islander"
+"Other" | "Other race, nec"
+"Middle Eastern" | "Other race, nec"
+"Mixed" | "Two major races"     
+"Mixed" | "Three or more major races"
+"Hispanic" | "Hispanic"
+
+```{r}
+race_mapping = list(
+    "White" = "White",
+    "Black" = "Black/African American/Negro",
+    "Native American" = "American Indian or Alaska Native",
+    "Asian" = "Chinese",
+    "Asian" = "Japanese",
+    "Asian" = "Other Asian or Pacific Islander",
+    "Other" = "Other race, nec",
+    "Middle Eastern" = "Other race, nec",
+    "Mixed" = "Two major races",
+    "Mixed" = "Three or more major races",
+    "Hispanic" = "Hispanic"
+)
+q_race <- SurveyQuestion$new(name = "race",
+    col_names = c("race", "race2"),
+    values_map= race_mapping)
+q_race$name()
+q_race$col_names()
+q_race$values()
+```
+
+This is yet another example of the inherent challenges of MRP.
+The work of creating a mapping is an extension
+of data assembly, validation, and exploratory data analysis,
+all of which requires researcher domain knowledge.
+
+#### Many-to-many mapping: marital status
+
+The CCES variable "marstat" and ACS variable "MARST" record the marital status of the respondant.
+
 ```{r}
@@
 
-CCES | ACS
----- | ---
-"Married" | "Married, spouse present"
-"Married" | "Married, spouse absent"
-"Separated" | "Separated"
-"Divorced" | "Divorced"
-"Widowed" | "Widowed"
-"Never married" | "Never married"
-"Domestic / civil partnership" | "Never married"
-"skipped" | "Never married"
-"not asked" | "Never married"
+For the 2018 CCES survey, the marital status of all respondants is known,
+therefore categories "skipped" and "not asked" are not problematic.
+The ACS variable ["MARST"](https://usa.ipums.org/usa-action/variables/MARST#comparability_section)
+has two categories for corresponding to CCES category "Married":
+"Married, spouse present", "Married, spouse absent".
+Mapping both ACS categories to CCES category "Married" is a one-to-many mapping
+which is allowed by MRP-KIT.
+The CCES has category "Domestic / civil partnership",
+for which there is no corresponding ACS category.
+Mapping both partnerships and "Never married" to the ACS category "Never married/single"
+is a many-to-one mapping from CCES to ACS, also allowed by MRP-KIT.
+However, if we map CES category  "Domestic / civil partnership"
+to either ACS category "Married, spouse present" or "Married, spouse absent",
+we create a many-to-many mapping.
 
-We note that for the 2018 CCES survey, the marital status of all respondants is known; nonetheless,
-we provide a complete mapping.
+OK             |                | NOT OK         | 
+-------------- | -------------- | -------------- | --------------
+CCES | ACS | CCES | ACS
+Married | Married, spouse present |Married | Married, spouse present 
+Married | Married, spouse absent |Married | Married, spouse absent 
+Separated | Separated |Separated | Separated 
+Divorced | Divorced |Divorced | Divorced 
+Widowed | Widowed |Widowed | Widowed 
+Never married | Never married/single |Never married | Never married/single 
+*Domestic / civil partnership* | *Never married/single* | *Domestic / civil partnership* | **Married, spouse present**
 
-```{r}
+```{r, error=TRUE}
+# This is OK
 marstat_mapping = list(
@@
     "Never married" = "Never married",
-    "Domestic / civil partnership" = "Never married",
-    "skipped" = "Never married",
-    "not asked" = "Never married"
+    "Domestic / civil partnership" = "Never married"
 )
@@
 q_marstat$col_names()
-head(q_marstat$values())
+q_marstat$values()
+
+# This is NOT
+marstat_mapping = list(
+    "Married" = "Married, spouse present",
+    "Married" = "Married, spouse absent",
+    "Separated" = "Separated",
+    "Divorced" = "Divorced",
+    "Widowed" = "Widowed",
+    "Never married" = "Never married",
+    "Domestic / civil partnership" = "Married, spouse present"
+)
+q_marstat <- SurveyQuestion$new(name = "marstat",
+    col_names = c("marstat", "MARST"),
+    values_map= marstat_mapping)
+
+```
+Here we propose to add yet another column to the ACS which conflates "Married, spouse present" and "Married, spouse absent".
+
+```{r}
+# create column which conflates spouse present, absent
+acs_marst2 <- as_factor(acs_2018_subset$MARST)
+levels(acs_marst2) = c(levels(acs_marst2), "Married")
+acs_marst2[acs_marst2 =="Married, spouse present"] <- "Married"
+acs_marst2[acs_marst2 =="Married, spouse absent"] <- "Married"
+
+# add to acs_2018_subset data.frame - respect column types
+acs_2018_subset$marst2 = to_labelled(droplevels(acs_marst2))
+ipums_val_labels(acs_2018_subset$marst2)
+```
+Now we can create a mapping which maps both "Married" and ""Domestic / civil partnership"
+to the the (new) ACS label "Married".^[
+Alternatively, we could try to access more ACS variables which account for
+same sex marriages and civil unions and use this information to code new variable "marst2".]
+
+```{r}
+marstat_mapping = list(
+    "Married" = "Married",
+    "Separated" = "Separated",
+    "Divorced" = "Divorced",
+    "Widowed" = "Widowed",
+    "Never married" = "Never married",
+    "Domestic / civil partnership" = "Married"
+)
+q_marstat <- SurveyQuestion$new(name = "marstat",
+    col_names = c("marstat", "marst2"),
+    values_map= marstat_mapping)
+q_marstat$name()
+q_marstat$col_names()
+q_marstat$values()
 ```
 
-#### Information split across survey columns:  race and ethnicity.
+#### Aligned demographics
 
-We ignore the politics of race and racial categories
-and instead focus on how best to map the CCES information to the ACS information.
-Both surveys have multiple variables which provide information about race and ethnicity,
-and both record race and hispanic origin separately.
-However the CCES treats hispanic origin as race as well as ethnic origin; the ACS does not.
-```{r}
-names(table(cces_2018_subset$race))
-names(table(cces_2018_subset$hispan))
-ipums_val_labels(acs_2018_subset$RACE)
-ipums_val_labels(acs_2018_subset$HISPAN)
-```
-In this case, munging the raw survey data is unavoidable.
-For each survey we create a new variable which blends race and hispanic ethnicity.
-We are scripting everyting, so we add this to our processing scripts.
+We have created 5 questions:  `q_sex`, `q_state`, `q_educ`, `q_race`, `q_marstat`.
+This will allow us to slice and dice our surveys to a ridiculous level of granularity.
+Excelsior!

### 2021-07-03 case-study/Munging_CCES.Rmd
completed mapping discussion

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(cache=TRUE, message=FALSE, error=FALSE, warning=FALSE, comment=NA, out.width='95%', collapse=TRUE, prompt=TRUE)
+knitr::opts_chunk$set(cache=TRUE, message=FALSE, error=FALSE, warning=FALSE, comment=NA, out.width='95%', collapse=TRUE, prompt=TRUE, highlight=FALSE)
 options(digits=2)
@@
 The Common Content portion of the survey consists of questions asked of all subjects.
+
 Profile questions provide demographic information,
@@
 col_names <- attr(cces_2018, "names")
+
 demo_cols <- c(which(col_names == "birthyr"),
@@
               which(col_names == "inputstate"))
-# don't allow missing demographics
-keep_rows <- complete.cases(cces_2018[, demo_cols])
+keep_rows <- complete.cases(cces_2018[, demo_cols]) # no missing demographics
 
@@
 ipums_val_labels(acs_2018_subset$AGE)
-ipums_val_labels(acs_2018_subset$AGE)
 ```
@@
 head(table(as_factor(acs_2018_subset$AGE)))
+```
+
+The CCES is a survey of _national adults_, i.e., eligible
+voters over the age of 18 but the ACS is a survey over the entire
+US population, including non-citizens.
+In order to use the ACS for poststratification, we drop all
+records for individuals under the age of 18 and non-citizens.
+
+```{r}
+table(as_factor(acs_2018_subset$CITIZEN))
+# drop underage, non-citizens
+citizen = as_factor(acs_2018_subset$CITIZEN) != "Not a citizen"
+over_18 = acs_2018_subset$AGE > 17
+acs_2018_subset = acs_2018_subset[(over_18 & citizen) , ]  
 ```
@@
 
-#### 1-1 mapping on a long list: state of residence, 51 categories
+#### One-to-one mapping over many labels: state of residence, 51 categories
 
@@
 ```
-We choose to align these categories as follows.
+We choose to align these categories so that the CCES label "2-year"
+maps to the ACS labels for 2 and 3 years of college.
+Mpping "Some College" to both 1 year and 3 years of college is also possible,
+on the rationale that the goal was a 4-year college degree, and that it wasn't obtained.
+The choice depends on both an understanding of the survey data and the particular goals
+of your analysis.
 
@@
 
-Mpping "Some College" to both 1 year and 3 years of college is also possible,
-reasoning that the goal of a 4-year college degree wasn't obtained.
-The choice depends on both an understanding of the survey data and the particular goals
-of your analysis.
-MRP-KIT can only provide the mechanism for recording and carrying out the mapping.
 ```{r}
@@
 acs_race2 <- as_factor(acs_2018_subset$RACE)
-
 # population percentages by label
 prop.table(table(acs_race2))
-
 # if "hispan" is "Yes", "race" label is "Hispanic".
@@
 acs_race2[is_hispan==TRUE] <- "Hispanic"
-
 # adjusted percentages
 prop.table(table(acs_race2))
-
 # add to acs_2018_subset data.frame - respect column types
@@
 prop.table(table(cces_2018_subset$race))
-
 # Intersection (and) vs. union (or) of race == "Hispanic" and hispanic == "Yes"
@@
 length(which(cces_2018_subset$race == "Hispanic" | cces_2018_subset$hispan == "Yes"))
-
 # create column which adjusts race labels
@@
 prop.table(table(cces_race2))
-
 # add to acs_2018_subset data.frame - respect column types
@@
 "Married, spouse present", "Married, spouse absent".
-Mapping both ACS categories to CCES category "Married" is a one-to-many mapping
-which is allowed by MRP-KIT.
+
 The CCES has category "Domestic / civil partnership",
 for which there is no corresponding ACS category.
-Mapping both partnerships and "Never married" to the ACS category "Never married/single"
-is a many-to-one mapping from CCES to ACS, also allowed by MRP-KIT.
-However, if we map CES category  "Domestic / civil partnership"
-to either ACS category "Married, spouse present" or "Married, spouse absent",
-we create a many-to-many mapping.
+We could try to access additional columns of the ACS, as there
+is some information on same-sex partnerships and civil unions.
+Otherwise, we have a choice of mapping partnerships to either
+of the ACS married categories, or to the "Never married/single" category.
+
+If we do the former, we create a many-to-many mapping:
+CCES category "Married" maps to two ACS categories: spouse present/absent;
+ACS category "Married, spouse present" maps to two CCES categories: married and partnerships.
+This is not allowed by MRP-KIT.
+If we do the latter, we create two many-to-one mappings:
+CCES category "Married" maps to two ACS categories: spouse present/absent;
+ACS category "Never married/single" maps to two CCES categories:
+never married and civil partnership.  This perfectly OK.
+
+*question: implication of many-to-one mappings between the survey population and the
+target population - do we get separate postratification cells?*
 
@@
 -------------- | -------------- | -------------- | --------------
-CCES | ACS | CCES | ACS
+**CCES** | **ACS** | **CCES** | **ACS**
+Never married | Never married/single |Never married | Never married/single 
+*Domestic / civil partnership* | *Never married/single* | *Domestic / civil partnership* | **Married, spouse present**
 Married | Married, spouse present |Married | Married, spouse present 
@@
 Widowed | Widowed |Widowed | Widowed
-Never married | Never married/single |Never married | Never married/single 
-*Domestic / civil partnership* | *Never married/single* | *Domestic / civil partnership* | **Married, spouse present**
 
 ```{r, error=TRUE}
-# This is OK
+# OK
 marstat_mapping = list(
+    "Never married" = "Never married/single",
+    "Domestic / civil partnership" = "Never married/single",
     "Married" = "Married, spouse present",
@@
     "Divorced" = "Divorced",
-    "Widowed" = "Widowed",
-    "Never married" = "Never married",
-    "Domestic / civil partnership" = "Never married"
+    "Widowed" = "Widowed"
 )
@@
 
-# This is NOT
+# NOT OK
 marstat_mapping = list(
+    "Never married" = "Never married/single",
+    "Domestic / civil partnership" = "Married, spouse present",
     "Married" = "Married, spouse present",
@@
     "Divorced" = "Divorced",
-    "Widowed" = "Widowed",
-    "Never married" = "Never married",
-    "Domestic / civil partnership" = "Married, spouse present"
+    "Widowed" = "Widowed"
 )
@@
 ```
-Here we propose to add yet another column to the ACS which conflates "Married, spouse present" and "Married, spouse absent".
+Since the distinction in the ACS data between spouse present and spouse absent isn't useful,
+we can solve the many-to-many mapping problem by adding a new column to the ACS
+which conflates "Married, spouse present" and "Married, spouse absent".
 
@@
 Alternatively, we could try to access more ACS variables which account for
-same sex marriages and civil unions and use this information to code new variable "marst2".]
+same sex marriages and civil unions and combine information across multiple columns, as we did
+for ACS variable race2.]
 
@@
 
-#### Aligned demographics
+#### Measurement recorded in different units and not binned:  age
 
-We have created 5 questions:  `q_sex`, `q_state`, `q_educ`, `q_race`, `q_marstat`.
-This will allow us to slice and dice our surveys to a ridiculous level of granularity.
-Excelsior!
+The CCES is records age as birthyear but the ACS records age in years.
+In both surveys, this is an integer, not a categorical measurement but
+MRP-KIT mappings operate over a finite set of categories.
+
+```{r}
+acs_birthyear = 2018 - acs_2018_subset$AGE
+summary(acs_birthyear)
+summary(cces_2018$birthyr)
+```
+We have to decide how many bins and which years they should span.
+We seek to bin the population in ways that respects what we know about
+US generational cohorts (citation?), e.g., "baby-boomers" born between 1945 an 1965.
+
+* 7 10-year+ cohorts: 1920-1935, 1936-1945, 1946-1955, 1956-1965, 1966-1975, 1976-1985, 1986-2000
+* 4 20-year+ cohorts: 1920-1945, 1946-1965, 1966-1985, 1986-2000
+
+Choosing the former will result in more postratification cells (57% more, to be precise).
+
+### Mapping the survey and target population data with MRP-KIT
+
+We now have mappings for our 6 postratification variables:
+`q_sex`, `q_state`, `q_educ`, `q_race`, `q_marstat`, `q_age`;
+allowing for (at least) `2 * 51 * 6 * 8 * 6 * 4`  possible subsets of interest.

### 2021-07-04 case-study/Munging_CCES.Rmd
added styling

@@
 output:
-  html_document:
-    toc: true
-    toc_depth: 2
+  rmdformats::robobook
 ---
@@
 
-### Survey data: the 2018 CCES Common Content Dataset
+# Survey data: the 2018 CCES Common Content Dataset
 
@@
 
-### Population data:  the 2018 American Community Survey
+# Population data:  the 2018 American Community Survey
 
@@
 
-#### Inspecting the ACS sample
+## Inspecting the ACS sample
 
@@
 
-### Aligning the survey and target population data with MRP-KIT
+# Aligning the survey and target population data with MRP-KIT
 
@@
 
-#### Simple one-to-one mapping:  CCES "gender" / ACS "sex"
+## Simple one-to-one mapping:  CCES "gender" / ACS "sex"
 
@@
 
-#### One-to-one mapping over many labels: state of residence, 51 categories
+## One-to-one mapping over many labels: state of residence, 51 categories
 
@@
 
-#### Choice of mapping: education
+## Choice of mapping: education
 
@@
 
-#### Survey variable cannot be mapped to single census variable:  Race
+## Survey variable cannot be mapped to single census variable:  Race
 
@@
 
-#### Many-to-many mapping: marital status
+## Many-to-many mapping: marital status
 
@@
 
-#### Measurement recorded in different units and not binned:  age
+## Measurement recorded in different units and not binned:  age
 
@@
 
-Choosing the former will result in more postratification cells (57% more, to be precise).
+## Possible subpopulations of interest
 
-### Mapping the survey and target population data with MRP-KIT
+The size of the postratification table is the product of the number of categories per demographic.
+The more fine-grained the classification, the larger the postratification table.
+In this example, we have variables:
+`q_sex` (2 values), `q_state` (51 values), `q_educ` (6 values), `q_race` (8 values), `q_marstat` (5 values), `q_age` (many possible values - e.g. 4 or 7).
+The resultng poststratification table will have at least 97,920 cells.
 
-We now have mappings for our 6 postratification variables:
-`q_sex`, `q_state`, `q_educ`, `q_race`, `q_marstat`, `q_age`;
-allowing for (at least) `2 * 51 * 6 * 8 * 6 * 4`  possible subsets of interest.
+# Mapping the survey and target population data with MRP-KIT

### 2021-07-04 case-study/Munging_CCES.Rmd
checkpointing - minor tweaks to layout

@@
 The 2018 CCES Common Content Dataset is 
-available from the Harvard Dataverse project:
-https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi%3A10.7910/DVN/ZSBZ7K
+available from the [Harvard Dataverse](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi%3A10.7910/DVN/ZSBZ7K).
+The Common Content portion of the survey consists of questions asked of all subjects.
+Profile questions provide demographic information
+and opinion questions provide outcomes of interest.
+
 The raw survey data is distributed in several formats from the Harvard Dataverse
@@
 names(cces_2018)[1:12]
-names(attributes(cces_2018))
 ```
@@
 The `read.dta13` creates attributes corresponding to the Stata file metadata.
-The attribute `names` corresponds to the dataframe column names and
-the attribute `val.labels` provides a short description of that columns.
-For questions which have categorical responses, the attribute `data.label`
+```{r}
+names(attributes(cces_2018))
+```
+The attribute `names` corresponds to the dataframe column names.
+```{r}
+attr(cces_2018, "names")[1:10]
+```
+The attribute `val.labels` provides a short description of that columns.
+```{r}
+attr(cces_2018, "var.labels")[1:10]
+```
+For questions which have categorical responses, the attribute `label.table`
 provides the category name and interger code used for that survey question.
 ```{r}
-names(attributes(cces_2018))
-attr(cces_2018, "names")[1:10]
-attr(cces_2018, "var.labels")[1:10]
 attr(cces_2018, "label.table")[["gender"]]
 ```
-
-The Common Content portion of the survey consists of questions asked of all subjects.
-
-Profile questions provide demographic information,
-and opinion questions provide outcomes of interest.
 The demographic variables include
@@
 table(cces_2018[[pres_vote]][,drop=TRUE], useNA="always")
-
 attr(cces_2018, "var.labels")[which(startsWith(attr(cces_2018, "var.labels"), "Health Care --"))[1]]
@@
 STATEFIP <- ipums_val_labels(acs_2018_subset$STATEFIP)[[2]][1:51]
-
 cat(inputstate[1:5], sep=", "); cat(STATEFIP[1:5], sep=", ")
-
 table(inputstate == STATEFIP)
@@
 mapping[1:3]
-
-# the one-liner (lowercase)
-
+# combine the two steps into a one-liner (lowercase)
 mapping = as.list(setNames(STATEFIP, tolower(inputstate)))
@@
 ```{r}
-mapping = as.list(setNames(STATEFIP, inputstate))
+state_mapping = as.list(setNames(STATEFIP, inputstate))
 q_state <- SurveyQuestion$new(name = "state",
 	col_names = c("inputstate", "STATEFIP"),
-	values_map= mapping)
+	values_map= state_mapping)
 q_state$name()

### 2021-07-10 case-study/Mapping_CCES_to_ACS.Rmd
checkpointing; created SurveyData objects

--- 2021-07-04 case-study/Munging_CCES.Rmd
@@
 
-```{r ccesDownload, echo=FALSE, out.width = "75%", fig.cap="Harvard Dataverse download page"}
+```{r ccesDownload, echo=FALSE, out.width="75%", fig.cap="Harvard Dataverse download page"}
 library(knitr)
@@
 cces_2018_subset <- subset(cces_2018, keep_rows, keep_cols)
-# `subset` drops attributes, keep attribute "var.labels"
-attr(cces_2018_subset, "var.labels") <- attr(cces_2018, "var.labels")
+# `subset` drops attributes, add back the ones we need
+attr(cces_2018_subset, "var.labels") <- attr(cces_2018, "var.labels")[keep_cols]
 
-# cleanup factor variables; drop unused category labels
-for (name in names(cces_2018_subset)) {
-    if (is.factor(cces_2018_subset[[name]]))  {
-        cces_2018_subset[[name]] = droplevels(cces_2018_subset[[name]])
-    }
-}
+cces_2018_subset <- droplevels(cces_2018_subset)  # remove unused categories
 ```
@@
 
-```{r ipumsExtract, echo=FALSE, out.width = "75%", fig.cap="IPUMS requested ACS variables"}
+```{r ipumsExtract, echo=FALSE, out.width="75%", fig.cap="IPUMS requested ACS variables"}
 library(knitr)
@@
 
-```{r ipumsDownload, echo=FALSE, out.width = "75%", fig.cap="IPUMS service download page"}
+```{r ipumsDownload, echo=FALSE, out.width="75%", fig.cap="IPUMS service download page"}
 library(knitr)
@@
 US population, including non-citizens.
-In order to use the ACS for poststratification, we drop all
-records for individuals under the age of 18 and non-citizens.
+We would like to poststratify to a representative survey population,
+i.e., the ACS micro-census, but that subset of the general population
+who meet the CCES survey criteria, i.e., elibible voters living in
+one of the 50 states or Washington DC.
+Therefore,  we drop all records for individuals under the age of 18,
+non-citizens, or residing outside of the 50 states or DC.
 
@@
 over_18 = acs_2018_subset$AGE > 17
-acs_2018_subset = acs_2018_subset[(over_18 & citizen) , ]  
+state_or_DC = acs_2018_subset$STATEFIP < 52
+acs_2018_subset = acs_2018_subset[(over_18 & citizen & state_or_DC) , ]  
 ```
@@
 
-# ACS data are structured objects, ipums_val_labels returns a tibble
-STATEFIP <- ipums_val_labels(acs_2018_subset$STATEFIP)[[2]][1:51]
-cat(inputstate[1:5], sep=", "); cat(STATEFIP[1:5], sep=", ")
-table(inputstate == STATEFIP)
+# ACS dataframe is haven_labelled, treat as factor, only keep states and DC
+STATEFIP_levels <- levels(as_factor(acs_2018_subset$STATEFIP))[1:51]
+cat("CCES:", inputstate[1:5], sep=", ");
+cat("ACS:", STATEFIP_levels[1:5], sep=", ")
+table(inputstate == STATEFIP_levels)
 ```
@@
 # create list, then set names (uppercase)
-mapping <- as.list(STATEFIP)
+mapping <- as.list(STATEFIP_levels)
 names(mapping) <- toupper(inputstate)
@@
 # combine the two steps into a one-liner (lowercase)
-mapping = as.list(setNames(STATEFIP, tolower(inputstate)))
+mapping = as.list(setNames(STATEFIP_levels, tolower(inputstate)))
 mapping[1:3]
@@
 ```{r}
-state_mapping = as.list(setNames(STATEFIP, inputstate))
+state_mapping = as.list(setNames(STATEFIP_levels, inputstate))
 q_state <- SurveyQuestion$new(name = "state",
@@
 ipums_val_labels(acs_2018_subset$RACE)
-````
+```
 
@@
 length(which(cces_2018_subset$race == "Hispanic" | cces_2018_subset$hispan == "Yes"))
-# create column which adjusts race labels
+# create column which adjusts race labels, add to dataframe
 cces_race2 <- cces_2018_subset$race
@@
 prop.table(table(cces_race2))
-# add to acs_2018_subset data.frame - respect column types
 cces_2018_subset$race2 = cces_race2
+attr(cces_2018_subset, "var.labels")[['race2']] <- "Race or Hispanic"
 ```
@@
 
-```{r, error=TRUE}
+```{r error=TRUE}
 # OK
@@
 
-## Measurement recorded in different units and not binned:  age
+## Measurement recorded in different units and not binned:  brithyear and age
 
@@
 
+```{r}
+# use R `cut` functon, choose breakpoints, labels
+cohort_breaks = c(1900, 1935, 1945, 1955, 1965, 1975, 1985, 2000)
+cohort_labels=c("before-1936",
+         "1936-1945",
+         "1946-1955",
+         "1956-1965",
+         "1966-1975",
+         "1976-1985",
+         "after 1986")
+
+# add to CCES
+cces_2018_subset$birthyr_cohort = cut(cces_2018_subset$birthyr, cohort_breaks, cohort_labels)
+attr(cces_2018_subset, "var.labels")[['birthyr_cohort']] <- "Birthyear cohort"
+
+# transform ACS age to birthyr, then add to ACS
+acs_2018_subset$birthyr = 2018 - acs_2018_subset$AGE
+acs_2018_subset$birthyr_cohort = to_labelled(cut(acs_2018_subset$birthyr, cohort_breaks, cohort_labels))
+```
+
+This is a straightforward 1:1 mapping
+```{r}
+birthyr_cohort_mapping = as.list(setNames(cohort_labels, cohort_labels))
+q_birthyr <- SurveyQuestion$new(name = "birthyear",
+	col_names = c("birthyr_cohort", "birthyr_cohort"),
+	values_map= birthyr_cohort_mapping)
+q_birthyr$name()
+q_birthyr$col_names()
+head(q_birthyr$values())
+```
+
 ## Possible subpopulations of interest
@@
 In this example, we have variables:
-`q_sex` (2 values), `q_state` (51 values), `q_educ` (6 values), `q_race` (8 values), `q_marstat` (5 values), `q_age` (many possible values - e.g. 4 or 7).
+`q_sex` (2 values), `q_state` (51 values), `q_educ` (6 values), `q_race` (8 values), `q_marstat` (5 values), `q_birthyr` (many possible values - e.g. 4 or 7).
 The resultng poststratification table will have at least 97,920 cells.
@@
 # Mapping the survey and target population data with MRP-KIT
+
+A SurveyData object requires a dataframe, a question list, and a response list.
+A question list is a list of tag-value pairs where the tag (name) is the dataframe column name and the value is a description of corresponding survey question.
+For the CCES, the dataframe column descriptions are lists in the attrbute "var.labels".
+As before, we create this list programatically using the `as.list(setNames(tags,values))` trick.
+```{r}
+cces_2018_questions <- as.list(setNames(attr(cces_2018_subset,"var.labels"), names(cces_2018_subset)))
+head(cces_2018_questions)
+```
+For the ACS, we use the function `ipumsr::ipums_var_label` to get a short description of the variable.
+```{r}
+acs_2018_var_labels <- c()
+for (name in names(acs_2018_subset)) {
+    acs_2018_var_labels <- c(acs_2018_var_labels, ipums_var_label(acs_2018_subset[[name]]))
+}
+acs_2018_questions <- as.list(setNames(acs_2018_var_labels, names(acs_2018_subset)))
+head(acs_2018_questions)
+```
+
+A response list is a list of name-value pairs which map the column name to the set of
+possible response values.
+We can create this list using a for loop, or by using R's `apply` function.
+```{r}
+# create response list using a for loop
+cces_2018_responses <- list()
+for (name in names(cces_2018_subset)) {
+    cces_2018_responses[[name]] <- levels(as.factor(cces_2018_subset[[name]]))
+}
+head(cces_2018_responses)
+
+acs_2018_responses <- list()
+for (name in names(acs_2018_subset)) {
+    acs_2018_responses[[name]] <- levels(as_factor(acs_2018_subset[[name]]))
+}
+acs_2018_responses[["MARST"]]
+acs_2018_responses[["race2"]]
+acs_2018_responses[["birthyr_cohort"]]
+
+# cces_2018_subset is a dataframe; R's `apply` function would also work
+# cces_2018_responses <- apply(cces_2018_subset, 2, function(x) levels(as.factor(x)))
+# if we coerce acs_2018_subset to dataframe, can do the same
+# acs_2018_responses <- apply(as.data.frame(cces_2018_subset), 2, function(x) levels(as.factor(x)))
+```
+
+We use these lists to create SurveyData objects for both the CCES and ACS.
+```{r}
+cces_2018_data <- SurveyData$new(cces_2018_subset, cces_2018_questions, cces_2018_responses)
+cces_2018_data
+acs_2018_data <- SurveyData$new(acs_2018_subset, acs_2018_questions, acs_2018_responses)
+acs_2018_data
+```

### 2021-07-10 case-study/process_cces_2018.R

@@ -1,69 +1,48 @@
 library(readstata13)
-cces_2018 = read.dta13("cces18_common_vv.dta", convert.factors = T, nonint.factors = T, generate.factors=T)
 
-attr_names <- names(attributes(cces_2018))
+cces_2018 = read.dta13("cces18_common_vv.dta", convert.factors = T, nonint.factors = T, generate.factors=T)
 col_labels <- attr(cces_2018, "var.labels")
 col_names <- attr(cces_2018, "names")
 
-demo_cols <- c(1, which(col_names == "birthyr"),
+demo_cols <- c(which(col_names == "birthyr"),
               which(col_names == "gender" ),
               which(col_names == "educ"),
               which(col_names == "race"),
               which(col_names == "hispanic"),
               which(col_names == "marstat"),
-#              which(col_names == "region"),
-#              which(col_names == "employ"),
-#              which(col_names == "hadjob"),
               which(col_names == "inputstate"))
-col_labels[demo_cols]
+keep_rows <- complete.cases(cces_2018[, demo_cols]) # no missing demographics
 
 outcome_cols <- c(which(startsWith(col_labels, "Abortion --")),
                  which(startsWith(col_labels, "Immigration --")),
                  which(startsWith(col_labels, "Taxes --")),
                  which(startsWith(col_labels, "Health Care --")),
                  which(startsWith(col_labels, "President 20")))
-col_labels[outcome_cols]
-
 keep_cols <- sort(c(demo_cols, outcome_cols))
 
-cces_2018_subset <- subset(cces_2018[keep_cols])
-attr(cces_2018_subset, "var.labels") <- col_labels[keep_cols]
-attr(cces_2018_subset, "names") <- col_names[keep_cols]
+cces_2018_subset <- subset(cces_2018, keep_rows, keep_cols)
+# `subset` drops attributes, add back the ones we need
+attr(cces_2018_subset, "var.labels") <- attr(cces_2018, "var.labels")[keep_cols]
+droplevels(cces_2018_subset)
 
-table(cces_2018_subset$gender)
-table(cces_2018_subset$educ)
-table(cces_2018_subset$race)
-table(cces_2018_subset$marstat)
-length(table(cces_2018_subset$inputstate))
-length(table(cces_2018_subset$birthyr))
+cces_race2 <- cces_2018_subset$race
[…]

### 2021-07-10 case-study/process_ipums_acs_2018.R

@@ -1,20 +1,50 @@
-# NOTE: To load data, you must download both the extract's data and the DDI
-# and also set the working directory to the folder with these files (or change the path below).
+if (!require("ipumsr")) stop("This case study requires ipumsr package. It can be installed using the following command: install.packages('ipumsr')")
+if (!require("labelled")) stop("This case study requires labelled package. It can be installed using the following command: install.packages('labelled')")
+library(ipumsr)
+library(labelled)
+library(readstata13)
 
-if (!require("ipumsr")) stop("Reading IPUMS data into R requires the ipumsr package. It can be installed using the following command: install.packages('ipumsr')")
 
 # download dataset from IPUMS
-# xml and dat file should have same name, be in same directory
+# xml ddi file and Stata dat file should have same name, be in same directory
 acs_2018_ddi <- read_ipums_ddi("usa_00007.xml")
 acs_2018 <- read_ipums_micro(acs_2018_ddi)
 
 # save object for later
-save(acs_2018, acs_2018_ddi, file= "acs_2018_subset.RData")
+save(acs_2018, acs_2018_ddi, file= "acs_2018_subset_raw.RData")
 
-acs_colnames <- names(acs_2018)
-print(acs_colnames)
-print(class(acs_colnames[1]))
-for (i in 1:length(acs_colnames)) {
-    print(acs_colnames[i])
-    print(ipums_val_labels(acs_2018[[i]]))
-}
+# drop underage, non-citizens
+citizen = as_factor(acs_2018$CITIZEN) != "Not a citizen"
+over_18 = acs_2018$AGE > 17
+acs_2018_subset = acs_2018[(over_18 & citizen) , ]  
+
+# create column which blends "race" "hispan"
+acs_race2 <- factor(as_factor(acs_2018_subset$RACE), levels=c(levels(as_factor(acs_2018_subset$RACE)), "Hispanic"))
+
+# if "hispan" is "Yes", "race" label is "Hispanic".
+is_hispan <- acs_2018_subset$HISPAN > 0 & acs_2018_subset$HISPAN < 5
+acs_race2[is_hispan==TRUE] <- "Hispanic"
+acs_2018_subset$race2 = to_labelled(acs_race2)
+
+# create column which conflates spouse present, absent
+acs_marst2 <- as_factor(acs_2018_subset$MARST)
[…]

### 2021-07-10 case-study/Mapping_CCES_to_ACS.Rmd
checkpointing; convert haven to datafram

@@
 acs_2018_subset_ddi <- read_ipums_ddi("usa_00007.xml")
-acs_2018_subset <- read_ipums_micro(acs_2018_subset_ddi)
+acs_2018_subset_tbl <- read_ipums_micro(acs_2018_subset_ddi)
 # save as R objects
-save(acs_2018_subset, acs_2018_subset_ddi, file= "acs_2018_subset.RData")
+save(acs_2018_subset_tbl, acs_2018_subset_ddi, file= "acs_2018_subset.RData")
 ```
@@
 load("acs_2018_subset.RData")
-str(acs_2018_subset, max.level=2, list.len=6)
-dim(acs_2018_subset)
-names(acs_2018_subset)
+str(acs_2018_subset_tbl, max.level=2, list.len=6)
+dim(acs_2018_subset_tbl)
+names(acs_2018_subset_tbl)
 ```
@@
 ```{r}
-length(table(acs_2018_subset$AGE))
-ipums_val_labels(acs_2018_subset$AGE)
+length(table(acs_2018_subset_tbl$AGE))
+ipums_val_labels(acs_2018_subset_tbl$AGE)
 ```
@@
 allows us to treat labelled vectors as factors.
-See the vignette https://tech.popdata.org/ipumsr/reference/ipums_var_info.html for further information.
+The function [`zap_labels`](https://haven.tidyverse.org/reference/zap_labels.html)
+allows us to treat labelled vectors as vectors.
+See https://haven.tidyverse.org/articles/semantics.html for further information.
 ```{r}
-head(table(acs_2018_subset$AGE))
-head(table(as_factor(acs_2018_subset$AGE)))
+head(table(acs_2018_subset_tbl$AGE))
+head(table(as_factor(acs_2018_subset_tbl$AGE)))
+head(table(zap_labels(acs_2018_subset_tbl$AGE)))
 ```
@@
 who meet the CCES survey criteria, i.e., elibible voters living in
-one of the 50 states or Washington DC.
-Therefore,  we drop all records for individuals under the age of 18,
-non-citizens, or residing outside of the 50 states or DC.
+one of the 50 states or Washington DC.  (Note: STATEFIP codes have gaps, Wyoming has STATEFIP code 56).
+Therefore,  we drop all records for individuals under the age of 18 and non-citizens, or citizens not
+living in the 50 states or DC.
+```{r}
+# using the level name here
+table(as_factor(acs_2018_subset_tbl$CITIZEN))
+citizen = as_factor(acs_2018_subset_tbl$CITIZEN) != "Not a citizen"
+# using the column value
+over_18 = acs_2018_subset_tbl$AGE > 17
+state_or_DC = acs_2018_subset_tbl$STATEFIP < 57  
+acs_2018_subset_tbl = acs_2018_subset_tbl[(over_18 & citizen & state_or_DC) , ]  
+```
 
+The metadata is in the object `acs_2018_subset_ddi`.
+Working with the `tbl_df` object is awkward, as the MRP-KIT API expects to work with simple dataframes.
+Therefore we create a data.frame object which contains only the columns of interest as factors,
+excepting "AGE" which is count, not categorical, data.
 ```{r}
-table(as_factor(acs_2018_subset$CITIZEN))
-# drop underage, non-citizens
-citizen = as_factor(acs_2018_subset$CITIZEN) != "Not a citizen"
-over_18 = acs_2018_subset$AGE > 17
-state_or_DC = acs_2018_subset$STATEFIP < 52
-acs_2018_subset = acs_2018_subset[(over_18 & citizen & state_or_DC) , ]  
+acs_2018_subset <- data.frame(
+    "STATEFIP" = droplevels(as_factor(acs_2018_subset_tbl$STATEFIP)),
+    "SEX" = droplevels(as_factor(acs_2018_subset_tbl$SEX)),
+    "AGE" = zap_labels(acs_2018_subset_tbl$AGE),
+    "MARST" = droplevels(as_factor(acs_2018_subset_tbl$MARST)),
+    "RACE" = droplevels(as_factor(acs_2018_subset_tbl$RACE)),
+    "HISPAN" = droplevels(as_factor(acs_2018_subset_tbl$HISPAN)),
+    "CITIZEN" = droplevels(as_factor(acs_2018_subset_tbl$CITIZEN)),
+    "EDUC" = droplevels(as_factor(acs_2018_subset_tbl$EDUC))
+)
+str(acs_2018_subset, max.level=2, list.len=12, give.attr=FALSE)
 ```
@@
 library(mrpkit)
-names(table(cces_2018_subset$gender))
-ipums_val_labels(acs_2018_subset$SEX)
+levels(cces_2018_subset$gender)
+levels(acs_2018_subset$SEX)
 q_sex <- SurveyQuestion$new(name="sex",
@@
 ```{r}
-# CCES data is a factor, levels are the labels
-inputstate <- levels(cces_2018_subset$inputstate)
-
-# ACS dataframe is haven_labelled, treat as factor, only keep states and DC
-STATEFIP_levels <- levels(as_factor(acs_2018_subset$STATEFIP))[1:51]
-cat("CCES:", inputstate[1:5], sep=", ");
-cat("ACS:", STATEFIP_levels[1:5], sep=", ")
-table(inputstate == STATEFIP_levels)
+cat("CCES:", levels(cces_2018_subset$inputstate)[1:5], "\n", sep=", ");
+cat("ACS:", levels(acs_2018_subset$STATEFIP)[1:5], "\n", sep=", ");
+table(levels(cces_2018_subset$inputstate) == levels(acs_2018_subset$STATEFIP))
 ```
@@
 ```{r}
-# create list, then set names (uppercase)
-mapping <- as.list(STATEFIP_levels)
-names(mapping) <- toupper(inputstate)
+state_names <- levels(cces_2018_subset$inputstate)
+state_values <- levels(acs_2018_subset$STATEFIP)
+# create named list in two steps
+mapping <- as.list(state_values)
+names(mapping) <- state_names 
 mapping[1:3]
-# combine the two steps into a one-liner (lowercase)
-mapping = as.list(setNames(STATEFIP_levels, tolower(inputstate)))
+# one-liner using `setNames`
+mapping = as.list(setNames(state_values, state_names))
 mapping[1:3]
@@
 ```{r}
-state_mapping = as.list(setNames(STATEFIP_levels, inputstate))
+state_mapping = as.list(setNames(state_values, state_names))
 q_state <- SurveyQuestion$new(name = "state",
@@
 ```{r}
-names(table(cces_2018_subset$educ))
-ipums_val_labels(acs_2018_subset$EDUC)
+levels(cces_2018_subset$educ)
+levels(acs_2018_subset$EDUC)
 ```
@@
 ```{r}
-names(table(cces_2018_subset$race))
-ipums_val_labels(acs_2018_subset$RACE)
+levels(cces_2018_subset$race)
+levels(acs_2018_subset$RACE)
 ```
@@
 ```{r}
-names(table(cces_2018_subset$hispan))
-ipums_val_labels(acs_2018_subset$HISPAN)
+levels(cces_2018_subset$hispan)
+levels(acs_2018_subset$HISPAN)
 ```
@@
 # create column which blends "race" "hispan"
-acs_race2 <- as_factor(acs_2018_subset$RACE)
+acs_race2 <- acs_2018_subset$RACE
 # population percentages by label
@@
 # if "hispan" is "Yes", "race" label is "Hispanic".
-is_hispan <- acs_2018_subset$HISPAN > 0 & acs_2018_subset$HISPAN < 5
+is_hispan <- acs_2018_subset$HISPAN != "Not Hispanic"
 length(is_hispan[is_hispan == TRUE]) / length(is_hispan)
@@
 # add to acs_2018_subset data.frame - respect column types
-acs_2018_subset$race2 = to_labelled(acs_race2)
+acs_2018_subset$race2 = acs_race2
 ```
@@
 ```{r}
-names(table(cces_2018_subset$marstat))
-ipums_val_labels(acs_2018_subset$MARST)
+levels(cces_2018_subset$marstat)
+levels(acs_2018_subset$MARST)
 ```
@@
 # create column which conflates spouse present, absent
-acs_marst2 <- as_factor(acs_2018_subset$MARST)
+acs_marst2 <- acs_2018_subset$MARST
 levels(acs_marst2) = c(levels(acs_marst2), "Married")
@@
 
-# add to acs_2018_subset data.frame - respect column types
-acs_2018_subset$marst2 = to_labelled(droplevels(acs_marst2))
-ipums_val_labels(acs_2018_subset$marst2)
+# add to acs_2018_subset data.frame
+acs_2018_subset$marst2 = acs_marst2
 ```
@@
 acs_2018_subset$birthyr = 2018 - acs_2018_subset$AGE
-acs_2018_subset$birthyr_cohort = to_labelled(cut(acs_2018_subset$birthyr, cohort_breaks, cohort_labels))
+acs_2018_subset$birthyr_cohort = cut(acs_2018_subset$birthyr, cohort_breaks, cohort_labels)
 ```
@@
 As before, we create this list programatically using the `as.list(setNames(tags,values))` trick.
-```{r}
+```{r, EVAL=FALSE}
 cces_2018_questions <- as.list(setNames(attr(cces_2018_subset,"var.labels"), names(cces_2018_subset)))
@@
 ```
-For the ACS, we use the function `ipumsr::ipums_var_label` to get a short description of the variable.
-```{r}
+For the ACS, we use function `ipumsr::ipums_var_label` to get a short description of the variable
+from the metadata in the DDI object.
+```{r, EVAL=FALSE}
 acs_2018_var_labels <- c()
@@
 We can create this list using a for loop, or by using R's `apply` function.
-```{r}
+```{r, EVAL=FALSE}
 # create response list using a for loop
@@
 We use these lists to create SurveyData objects for both the CCES and ACS.
-```{r}
+```{r, EVAL=FALSE}
 cces_2018_data <- SurveyData$new(cces_2018_subset, cces_2018_questions, cces_2018_responses)
@@
 ```
+Now we can create the mapping object from the CCES to the ACS.
+```{r, EVAL=FALSE}
+cces_acs_mapping <- SurveyMap$new(cces_2018_data, acs_2018_data, q_sex, q_state, q_race, q_educ)
+# poststrat table size: 2 * 51 * 8 * 6
+```

### 2021-07-11 case-study/Mapping_CCES_to_ACS.Rmd
checkpointing

@@
 ```
-We choose to align these categories so that the CCES label "2-year"
-maps to the ACS labels for 2 and 3 years of college.
-Mpping "Some College" to both 1 year and 3 years of college is also possible,
-on the rationale that the goal was a 4-year college degree, and that it wasn't obtained.
+
+For the ACS data, a possible label is "3 years of college" however
+in the ACS subset used here, this label isn't used.
+If it were used, we would have a choice between mapping it to CCES
+labels "Some college" or "2-year", on the rational that if the
+ goal was a 4-year college degree, and that it wasn't obtained.
 The choice depends on both an understanding of the survey data and the particular goals
@@
 "No HS" | "Grade 11"
-"High School Graduate" | "Grade 12"
-"Some College" | "1 year of college"
+"High school graduate" | "Grade 12"
+"Some college" | "1 year of college"
 "2-year" | "2 years of college"
-"2-year" | "3 years of college"
 "4-year" | "4 years of college"
@@
     "No HS" = "Grade 11",
-    "High School Graduate" = "Grade 12",
-    "Some College" = "1 year of college",
+    "High school graduate" = "Grade 12",
+    "Some college" = "1 year of college",
     "2-year" = "2 years of college",
-    "2-year" = "3 years of college",
     "4-year" = "4 years of college",
@@
 acs_marst2[acs_marst2 =="Married, spouse absent"] <- "Married"
-
+acs_marst2 <- droplevels(acs_marst2)  # unused levels not allowed
 # add to acs_2018_subset data.frame
@@
     "Widowed" = "Widowed",
-    "Never married" = "Never married",
+    "Never married" = "Never married/single",
     "Domestic / civil partnership" = "Married"
@@
 
-A SurveyData object requires a dataframe, a question list, and a response list.
-A question list is a list of tag-value pairs where the tag (name) is the dataframe column name and the value is a description of corresponding survey question.
+The components of an MRP-KIT mapping are the survey data, the target population data,
+and one or more mappings between demographics of interest.  In the previous sections
+we assembled the sample and target population datasets and defined the SurveyQuestion
+objects for to each demographic, adjusting the datasets accordingly.
+Next we create SurveyData objects for each dataset.
+
+A SurveyData object records the contents of a dataset in terms of questions and
+responses.  Every column is (potentially) a question, and the responses are the set
+of category labels for that column.  (*what about continuous variables?*)
+The question and responses are name-value pair lists.
+For the question list, the name is the dataframe column name and the value is the description of the variable or the survey question asked.
+For the response list, the name is also the dataframe column name and the value is the set of possible responses.
+
 For the CCES, the dataframe column descriptions are lists in the attrbute "var.labels".
-As before, we create this list programatically using the `as.list(setNames(tags,values))` trick.
-```{r, EVAL=FALSE}
+As before, we create this list programatically using the `as.list(setNames(tags,values))` one-liner.
+```{r}
 cces_2018_questions <- as.list(setNames(attr(cces_2018_subset,"var.labels"), names(cces_2018_subset)))
@@
 For the ACS, we use function `ipumsr::ipums_var_label` to get a short description of the variable
-from the metadata in the DDI object.
-```{r, EVAL=FALSE}
+from the metadata in the `acs_2018_subset_tbl` object.
+```{r}
 acs_2018_var_labels <- c()
 for (name in names(acs_2018_subset)) {
-    acs_2018_var_labels <- c(acs_2018_var_labels, ipums_var_label(acs_2018_subset[[name]]))
+    acs_2018_var_labels <- c(acs_2018_var_labels, ipums_var_label(acs_2018_subset_tbl[[name]]))
 }
@@
 possible response values.
-We can create this list using a for loop, or by using R's `apply` function.
-```{r, EVAL=FALSE}
-# create response list using a for loop
-cces_2018_responses <- list()
-for (name in names(cces_2018_subset)) {
-    cces_2018_responses[[name]] <- levels(as.factor(cces_2018_subset[[name]]))
-}
+We create this list by using R's `apply` function.
+```{r}
+cces_2018_responses <- apply(cces_2018_subset, 2, function(x) levels(as.factor(x)))
 head(cces_2018_responses)
-
-acs_2018_responses <- list()
-for (name in names(acs_2018_subset)) {
-    acs_2018_responses[[name]] <- levels(as_factor(acs_2018_subset[[name]]))
-}
-acs_2018_responses[["MARST"]]
-acs_2018_responses[["race2"]]
-acs_2018_responses[["birthyr_cohort"]]
-
-# cces_2018_subset is a dataframe; R's `apply` function would also work
-# cces_2018_responses <- apply(cces_2018_subset, 2, function(x) levels(as.factor(x)))
-# if we coerce acs_2018_subset to dataframe, can do the same
-# acs_2018_responses <- apply(as.data.frame(cces_2018_subset), 2, function(x) levels(as.factor(x)))
+acs_2018_responses <- apply(acs_2018_subset, 2, function(x) levels(as.factor(x)))
+head(acs_2018_responses)
 ```
@@
 We use these lists to create SurveyData objects for both the CCES and ACS.
-```{r, EVAL=FALSE}
+```{r}
 cces_2018_data <- SurveyData$new(cces_2018_subset, cces_2018_questions, cces_2018_responses)
@@
 Now we can create the mapping object from the CCES to the ACS.
-```{r, EVAL=FALSE}
-cces_acs_mapping <- SurveyMap$new(cces_2018_data, acs_2018_data, q_sex, q_state, q_race, q_educ)
-# poststrat table size: 2 * 51 * 8 * 6
+The `print` function displays the mappings between the survey data and target population data.
+```{r}
+cces_acs_mapping <- SurveyMap$new(cces_2018_data, acs_2018_data, q_sex, q_state, q_educ, q_race, q_marstat, q_birthyr)
+print(cces_acs_mapping)
 ```
+The `validate` function identifies mapped/unmapped columns and validates the mappings defined by the SurveyQuestion objects
+against the levels defined on the datasets.
+```{r}
+cces_acs_mapping$validate()
+```
+
+The `tabulate` function creates the post-stratification table.
+```{r}
+cces_acs_mapping$tabulate()
+```
+
+# MRP Part One:  Multi-level regression
+
+Given the survey data, we fit a model to the survey data, using our favorite R package of choice.
+
+# MRP Part Two:  Post-stratification

### 2021-07-23 mrpkit_haven_labelled.Rmd
add the code for labelled handling to be reviewed.

---
title: "handle the labelled data in mrpkit"
author: "Dewi Amaliah"
date: "12/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(haven)

```

The data:

1. Smoke data (NYTS2017) could be downloaded [here](https://www.cdc.gov/tobacco/data_statistics/surveys/nyts/data/index.html)
2. Wages data could be downloaded [here](https://onedrive.live.com/?authkey=%21AP6PMk0EBH%2Dx59M&cid=07C4D937B78DE3CA&id=7C4D937B78DE3CA%2127454&parId=7C4D937B78DE3CA%2127445&o=OneUp)
3. titanic data could be downloaded [here](https://drive.google.com/file/d/1FTE_0cFH8mroRrgTvtWVsPMTDvUD71T4/view?usp=sharing)

```{r}
smoke <- read_sas("nyts2020.sas7bdat")
wages <- read_sav("Wages.sav")
shape <- mrpkit::shape_survey
titan <- read_sav("titanic.sav")
```

```{r}

# here, the example is using titanic data
data <- wages

# this function will keep the factor, character, binary response, and labelled data (which is essentially a factor data, and if it is, it would be factorize in the next code)
keep <- function(x) is.factor(x) || is.character(x) || length(unique(x)) == 2 || !is.null(attr(x, "labels"))

# subset the variable that is in the keep criteria
data_use <- data[, sapply(data, keep), drop = FALSE]

# factorize the data that has a value labels
data_factorize <- as.data.frame(lapply(data_use, function(x) 
  if(!is.null(attr(x, "labels"))) {
    if (0 %in% x){as.factor(names(attr(x, "labels")[as.numeric(paste(x))+1]))}
    else {as.factor(names(attr(x, "labels")[as.numeric(paste(x))]))}}
  else 
    {x}))

# the downside of data_factorize is that it omit the variable's label (i.e., the question)
# hence, to extract the question, we will use data_use, instead of data factorize
# data factorize will only be used to extract the response, and its change the value of variable (according to its level, so might be this gonna be used in the next stage?)

#### extract the question

# select the variable that has a label and the class of its label is character. It is because, the variables that has labels, but doesn't have question label, would be counted as variable that has a question label, but its class is numeric, for example, sex and south in wages data, and sex in the titanic data. 
q_labelled <- dplyr::select_if(data_use, function(x) !is.null(attr(x, "label")) & class(attr(x, "label")) == "character")

# extract the question
questions <- lapply(q_labelled, function(x) setNames(attr(x, "label"), names(x)))

# select the variable that does not have label or having a label, but its label is numeric (for example, sex and south in wages data, and sex in the titanic data)
q_not_labelled <- dplyr::select_if(data_use, function(x) is.null(attr(x, "label")) | (!is.null(attr(x, "label")) & class(attr(x, "label")) == "numeric"))

# append the list of questions
append(questions, setNames(as.list(colnames(q_not_labelled)), colnames(q_not_labelled)))
```

```{r}
# get the response from the factorize data
responses <- lapply(data_factorize, function(x) if (is.factor(x)) levels(x) else unique(x))
responses
```

### 2021-08-04 R/data.R
fix r cmd check warnings

@@ -70,7 +70,7 @@
 #' \describe{
 #'  \item{age_group}{Age group}
 #'  \item{gender}{Gender}
-#'  \item{vote_for}{Party voted for in the 2016 election}
+#'  \item{vote_pref}{Party voted for in the 2016 election}
 #'  \item{education}{Highest grade of education completed}
 #'  \item{state}{State}
 #'  \item{wt}{Raked and trimmed survey weight}

### 2021-08-09 R/SurveyData.R
Pull the latest change from master branch.

@@ -200,6 +200,7 @@ SurveyData <- R6::R6Class(
 
       if (length(weights) == 0) {
         wts <- rep(1, nrow(data))
+        warning("'Weights have not been provided, assume all data weighted with weight 1.", call. = FALSE)
       } else {
         if (!is.character(weights) || !weights %in% colnames(data)) {
           stop("'weights' must be a string naming a column in 'data'.", call. = FALSE)

### 2021-08-09 R/SurveyFit.R

@@ -142,12 +142,17 @@ SurveyFit <- R6::R6Class(
           out[out[by] == focus_level, "value"] <-
             apply(poststrat_estimates[level_loc, ], 2, function(x) sum(poststrat_data$N_j[level_loc]*x)/sum(poststrat_data$N_j[level_loc]))
         }
+        out <- out %>%
+          dplyr::select(-"draw")
       } else {
         out <- data.frame(value = apply(poststrat_estimates, 2, function(x) sum(poststrat_data$N_j*x)/sum(poststrat_data$N_j)))
       }
       out
     },
-
+    #' @description Plot takes the aggregated estimates and produces a quick visualization total and sub-population estimates.
+    #' @param aggregated_estimates The object returned by `aggregate`
+    #' @param weights TRUE (default) if weighted estimates are included for comparison. Weighted interval is a 95% interval.
+    #'  If no weights are specified, weights are assumed to be 1, which is roughly equivalent to the observed data average.
     plot = function(aggregated_estimates, weights = TRUE) {
       if (ncol(aggregated_estimates) > 2) {
         focus_var <- colnames(aggregated_estimates)[1]

### 2021-08-09 tests/testthat/test-SurveyData.R

@@ -37,11 +37,11 @@ test_that("error if weights have NAs", {
 })
 
 test_that("weights equal to 1 if not specified", {
-  x <- SurveyData$new(
+  suppressWarnings(x <- SurveyData$new(
     data = feline_survey,
     questions = list(age1 = "Please identify your age group"),
     responses = list(age1 = levels(feline_survey$age1))
-  )
+  ))
   expect_true(all(x$weights() == 1))
 })
 

### 2021-08-30 R/SurveyData.R
another case of na.omit in SurveyData

@@ -146,7 +146,7 @@ SurveyData <- R6::R6Class(
         keep <- function(x) is.factor(x) || is.character(x) || length(unique(na.omit(x))) == 2
         data_use <- data[, sapply(data, keep), drop = FALSE]
         questions <- setNames(as.list(colnames(data_use)), colnames(data_use))
-        responses <- lapply(data_use, function(x) if (is.factor(x)) levels(x) else unique(x))
+        responses <- lapply(data_use, function(x) if (is.factor(x)) levels(x) else unique(na.omit(x)))
         warning(
           "No 'questions' and 'responses' provided. ",
           "Using all factor, character, and binary variables in 'data' by default.",

### 2021-08-30 R/SurveyData.R
Fix typo in error message
Fix typo in line 176

@@ -173,7 +173,7 @@ SurveyData <- R6::R6Class(
         stop("Names in 'questions' and 'responses' lists must be the same.")
       }
       if (!all(nms_q %in% colnames(data))) {
-        stop("Names of 'questions' much match column names in 'data'.", call. = FALSE)
+        stop("Names of 'questions' must match column names in 'data'.", call. = FALSE)
       }
       questions <- questions[nms_q]
       responses <- responses[nms_q]

### 2021-08-31 tests/testthat/test-SurveyData.R
Update test-SurveyData.R

@@ -63,7 +63,7 @@ test_that("error if question names don't match data names", {
       questions = list(banana = "Please identify your age group"),
       responses = list(banana = 1:4)
     ),
-    "Names of 'questions' much match column names in 'data'."
+    "Names of 'questions' must match column names in 'data'."
   )
 })
 

### 2021-09-14 R/SurveyFit.R
Added formula to survey fit obj

@@ -12,7 +12,8 @@ SurveyFit <- R6::R6Class(
   classname = "SurveyFit",
   private = list(
     map_ = NULL,
-    fit_ = NULL
+    fit_ = NULL,
+    private$formula_ <- NULL
   ),
   public = list(
 
@@ -21,9 +22,10 @@ SurveyFit <- R6::R6Class(
     #'   not need to be called directly by the user.
     #' @param fit A fitted model object.
     #' @param map A [`SurveyMap`] object.
-    initialize = function(fit, map) {
+    initialize = function(fit, map,formula) {
       private$fit_ <- fit
       private$map_ <- map
+      private$map_ <- formula
       invisible(self)
     },
 
@@ -35,6 +37,11 @@ SurveyFit <- R6::R6Class(
     #' @description Access the SurveyMap object
     map = function() {
       private$map_
+    },
+
+    #' @description Access the SurveyMap object
+    formula = function() {
+      private$formula_
     },
 
     #' @description Call the fitted model object's print method

### 2021-09-14 R/SurveyMap.R

@@ -545,7 +545,7 @@ SurveyMap <- R6::R6Class(
       args$formula <- formula
       args$data <- cbind(mapped_data, y_and_x)
       fit <- do.call(fun, args)
-      SurveyFit$new(fit = fit, map = self)
+      SurveyFit$new(fit = fit, map = self, formula = formula)
     },
 
     #' @description Access the `item_map`

### 2021-09-14 R/SurveyFit.R
typo

@@ -13,7 +13,7 @@ SurveyFit <- R6::R6Class(
   private = list(
     map_ = NULL,
     fit_ = NULL,
-    private$formula_ <- NULL
+    formula_ <- NULL
   ),
   public = list(
 
@@ -22,7 +22,7 @@ SurveyFit <- R6::R6Class(
     #'   not need to be called directly by the user.
     #' @param fit A fitted model object.
     #' @param map A [`SurveyMap`] object.
-    initialize = function(fit, map,formula) {
+    initialize = function(fit, map, formula) {
       private$fit_ <- fit
       private$map_ <- map
       private$map_ <- formula

### 2021-09-14 R/SurveyFit.R
typo

@@ -25,7 +25,7 @@ SurveyFit <- R6::R6Class(
     initialize = function(fit, map, formula) {
       private$fit_ <- fit
       private$map_ <- map
-      private$map_ <- formula
+      private$formula_ <- formula
       invisible(self)
     },
 

### 2021-09-14 R/SurveyFit.R
typo

@@ -13,7 +13,7 @@ SurveyFit <- R6::R6Class(
   private = list(
     map_ = NULL,
     fit_ = NULL,
-    formula_ <- NULL
+    formula_ = NULL
   ),
   public = list(
 

### 2021-09-15 R/SurveyFit.R
Change formula calls to refer to formula in fit object

@@ -167,8 +167,7 @@ SurveyFit <- R6::R6Class(
       if(length(dim(aggregated_estimates))!=2 | is.null(dim(aggregated_estimates))){
         stop("dimensions of aggregated estimates must be two")
       }
-      model_fit <- private$fit_
-      lhs_var <- as.character(formula(model_fit))[[2]]
+      lhs_var <- as.character(self$formula()[[2]])
       if (ncol(aggregated_estimates) > 1) {
         by_var <- colnames(aggregated_estimates)[1]
         wtd_ests <- data.frame(create_wtd_ests(self, lhs_var, by=by_var), method = "wtd")
@@ -176,15 +175,15 @@ SurveyFit <- R6::R6Class(
         mrp_ests_df <- sapply(mrp_ests_tmp, function(x) x)
         mrp_ests <- data.frame(t(mrp_ests_df),by_var = colnames(mrp_ests_df), method = "mrp")
         colnames(mrp_ests)[3] = by_var
-        lhs_binary <- force_factor(private$fit_$data[,lhs_var])
-        raw_ests_tmp <- by(lhs_binary, private$fit_$data[,by_var],function(x)c(mean = mean(x),sd = sqrt(mean(x)*(1-mean(x))/length(x))))
+        lhs_binary <- force_factor(self$map()$sample()$survey_data()[,lhs_var])
+        raw_ests_tmp <- by(lhs_binary, self$map()$sample()$survey_data()[,by_var],function(x)c(mean = mean(x),sd = sqrt(mean(x)*(1-mean(x))/length(x))))
         raw_ests_df <- sapply(raw_ests_tmp, function(x) x)
         raw_ests <- data.frame(t(raw_ests_df),by_var = colnames(raw_ests_df), method = "raw")
         colnames(raw_ests)[3] = by_var
       } else{
         wtd_ests <- data.frame(create_wtd_ests(self, lhs_var), method = "wtd")
         mrp_ests <- data.frame(mean=mean(aggregated_estimates$value),sd = sd(aggregated_estimates$value), method = "mrp")
-        lhs_binary <- force_factor(private$fit_$data[,lhs_var])
+        lhs_binary <- force_factor(self$map()$sample()$survey_data()[,lhs_var])
         raw_ests <- data.frame(mean = mean(lhs_binary),sd = sqrt(mean(lhs_binary)*(1-mean(lhs_binary))/length(lhs_binary)), method = "raw")
       }
       out <- rbind(mrp_ests, raw_ests,wtd_ests)
[…]

### 2021-09-15 R/SurveyMap.R

@@ -115,19 +115,19 @@
 #' )
 #'
 #' # Example lme4 usage
-#' # fit_2 <- ex_map$fit(
-#' #   fun = lme4::glmer,
-#' #   formula = y ~ (1|age) + (1|gender),
-#' #   family = "binomial"
-#' # )
-#' #
+#' fit_2 <- ex_map$fit(
+#'   fun = lme4::glmer,
+#'   formula = y ~ (1|age) + (1|gender),
+#'   family = "binomial"
+#' )
+#'
 #' # Example brms usage
-#' # fit_3 <- ex_map$fit(
-#' #   fun = brms::brm,
-#' #   formula = y ~ (1|age) + (1|gender),
-#' #   family = "bernoulli",
-#' #   seed = 1111
-#' # )
+#' fit_3 <- ex_map$fit(
+#'   fun = brms::brm,
+#'   formula = y ~ (1|age) + (1|gender),
+#'   family = "bernoulli",
+#'   seed = 1111
+#' )
 #'
 #'
 #' # predicted probabilities
@@ -137,8 +137,9 @@
 #' # estimates by age level
 #' estimates_by_age <- fit_1$aggregate(poststrat_estimates, by = "age")
 #'
-#'fit_1$summary()
-#'fit_1$summary(estimates_by_age)
+#' fit_1$summary(estimates_by_age)
+#' fit_2$summary(estimates_by_age)
+#' fit_3$summary(estimates_by_age)
 #'
 #' # plot estimates by age
 #' fit_1$plot(estimates_by_age)

### 2021-09-15 tests/testthat/test-SurveyFit.R

@@ -229,12 +229,12 @@ test_that("aggregate throws correct errors", {
 
 test_that("plot returns ggplot object", {
   popn <- fit_stan_glmer$aggregate(fit_stan_glmer$population_predict())
-  expect_s3_class(fit_stan_glmer$plot(popn, weights = TRUE), "ggplot")
-  expect_s3_class(fit_stan_glmer$plot(popn, weights = FALSE), "ggplot")
+  expect_s3_class(fit_stan_glmer$plot(popn, additional_stats = "mrp"), "ggplot")
+  expect_s3_class(fit_stan_glmer$plot(popn, additional_stats = "none"), "ggplot")
 
   by_age <- fit_stan_glmer$aggregate(fit_stan_glmer$population_predict(), by = "age")
-  expect_s3_class(fit_stan_glmer$plot(by_age, weights = TRUE), "ggplot")
-  expect_s3_class(fit_stan_glmer$plot(by_age, weights = FALSE), "ggplot")
+  expect_s3_class(fit_stan_glmer$plot(by_age, additional_stats = c("wtd","raw")), "ggplot")
+  expect_s3_class(fit_stan_glmer$plot(by_age, additional_stats = "none"), "ggplot")
 })
 
 test_that("Warnings provided if weights are all equal to 1",{

### 2021-09-15 R/SurveyFit.R
Refer to the mapped data not original data

@@ -175,15 +175,15 @@ SurveyFit <- R6::R6Class(
         mrp_ests_df <- sapply(mrp_ests_tmp, function(x) x)
         mrp_ests <- data.frame(t(mrp_ests_df),by_var = colnames(mrp_ests_df), method = "mrp")
         colnames(mrp_ests)[3] = by_var
-        lhs_binary <- force_factor(self$map()$sample()$survey_data()[,lhs_var])
-        raw_ests_tmp <- by(lhs_binary, self$map()$sample()$survey_data()[,by_var],function(x)c(mean = mean(x),sd = sqrt(mean(x)*(1-mean(x))/length(x))))
+        lhs_binary <- force_factor(self$map()$mapped_sample_data()[,lhs_var])
+        raw_ests_tmp <- by(lhs_binary, self$map()$mapped_sample_data()[,by_var],function(x)c(mean = mean(x),sd = sqrt(mean(x)*(1-mean(x))/length(x))))
         raw_ests_df <- sapply(raw_ests_tmp, function(x) x)
         raw_ests <- data.frame(t(raw_ests_df),by_var = colnames(raw_ests_df), method = "raw")
         colnames(raw_ests)[3] = by_var
       } else{
         wtd_ests <- data.frame(create_wtd_ests(self, lhs_var), method = "wtd")
         mrp_ests <- data.frame(mean=mean(aggregated_estimates$value),sd = sd(aggregated_estimates$value), method = "mrp")
-        lhs_binary <- force_factor(self$map()$sample()$survey_data()[,lhs_var])
+        lhs_binary <- force_factor(self$map()$mapped_sample_data()[,lhs_var])
         raw_ests <- data.frame(mean = mean(lhs_binary),sd = sqrt(mean(lhs_binary)*(1-mean(lhs_binary))/length(lhs_binary)), method = "raw")
       }
       out <- rbind(mrp_ests, raw_ests,wtd_ests)

### 2021-09-15 tests/testthat/test-SurveyFit.R

@@ -296,6 +296,23 @@ test_that("plot method returns warnings for incorrect additional stats input",{
                "when choosing no additional statistics, only supply `none`")
 })
 
+test_that("summary works with different fit objects",{
+
+  x <- fit_brms$aggregate(fit_brms$population_predict(), by = "age")
+  expect_s3_class(fit_brms$summary(x), "data.frame")
+
+  x <- fit_glmer$aggregate(fit_glmer$population_predict(), by = "age")
+  expect_s3_class(fit_brms$summary(x), "data.frame")
+
+  x <- fit_stan_glm$aggregate(fit_stan_glm$population_predict(), by = "age")
+  expect_s3_class(fit_stan_glm$summary(x), "data.frame")
+
+  x <- fit_stan_glmer$aggregate(fit_stan_glmer$population_predict(), by = "age")
+  expect_s3_class(fit_stan_glmer$summary(x), "data.frame")
+
+
+})
+
 test_that("plot appearance hasn't changed", {
   skip_on_cran()
   skip_if_not_installed("vdiffr")

### 2021-09-15 R/SurveyFit.R
binary variable is only in the OG data

@@ -175,7 +175,7 @@ SurveyFit <- R6::R6Class(
         mrp_ests_df <- sapply(mrp_ests_tmp, function(x) x)
         mrp_ests <- data.frame(t(mrp_ests_df),by_var = colnames(mrp_ests_df), method = "mrp")
         colnames(mrp_ests)[3] = by_var
-        lhs_binary <- force_factor(self$map()$mapped_sample_data()[,lhs_var])
+        lhs_binary <- force_factor(self$map()$sample()$survey_data()[,lhs_var])
         raw_ests_tmp <- by(lhs_binary, self$map()$mapped_sample_data()[,by_var],function(x)c(mean = mean(x),sd = sqrt(mean(x)*(1-mean(x))/length(x))))
         raw_ests_df <- sapply(raw_ests_tmp, function(x) x)
         raw_ests <- data.frame(t(raw_ests_df),by_var = colnames(raw_ests_df), method = "raw")
@@ -183,7 +183,7 @@ SurveyFit <- R6::R6Class(
       } else{
         wtd_ests <- data.frame(create_wtd_ests(self, lhs_var), method = "wtd")
         mrp_ests <- data.frame(mean=mean(aggregated_estimates$value),sd = sd(aggregated_estimates$value), method = "mrp")
-        lhs_binary <- force_factor(self$map()$mapped_sample_data()[,lhs_var])
+        lhs_binary <- force_factor(self$map()$sample()$survey_data()[,lhs_var])
         raw_ests <- data.frame(mean = mean(lhs_binary),sd = sqrt(mean(lhs_binary)*(1-mean(lhs_binary))/length(lhs_binary)), method = "raw")
       }
       out <- rbind(mrp_ests, raw_ests,wtd_ests)

### 2021-09-15 tests/testthat/test-SurveyFit.R
Tidied up tests that would fail if issue #124 happened again

@@ -298,19 +298,20 @@ test_that("plot method returns warnings for incorrect additional stats input",{
 
 test_that("summary works with different fit objects",{
 
+  skip_if_not_installed("brms")
   x <- fit_brms$aggregate(fit_brms$population_predict(), by = "age")
   expect_s3_class(fit_brms$summary(x), "data.frame")
 
+  skip_if_not_installed("lme4")
   x <- fit_glmer$aggregate(fit_glmer$population_predict(), by = "age")
   expect_s3_class(fit_brms$summary(x), "data.frame")
 
+  skip_if_not_installed("rstanarm")
   x <- fit_stan_glm$aggregate(fit_stan_glm$population_predict(), by = "age")
   expect_s3_class(fit_stan_glm$summary(x), "data.frame")
 
   x <- fit_stan_glmer$aggregate(fit_stan_glmer$population_predict(), by = "age")
   expect_s3_class(fit_stan_glmer$summary(x), "data.frame")
-
-
 })
 
 test_that("plot appearance hasn't changed", {

### 2021-09-15 tests/testthat/test-SurveyFit.R
Changed image tests

@@ -328,32 +328,41 @@ test_that("plot appearance hasn't changed", {
                                  0.804,0.847, 0.850, 0.823)
                        )
 
-  # vdiffr::write_svg(fit_stan_glmer$plot(popn, additional_stats = "none"),
-  #                   "tests/figs/SurveyFit/plot-population.svg", title = "plot-population")
-  #
-  # vdiffr::expect_doppelganger(
-  #   "plot-population",
-  #   fit_stan_glmer$plot(popn, additional_stats = "none"),
-  #   path = "tests/figs/SurveyFit/plot-population.svg"
-  # )
-  #
-  # vdiffr::expect_doppelganger(
-  #   "plot-population-weights",
-  #   fit_stan_glmer$plot(popn, additional_stats = c("mrp","raw","wtd")),
-  #   path = "SurveyFit"
-  # )
-  #
-  # vdiffr::expect_doppelganger(
-  #   "plot-group",
-  #   fit_stan_glmer$plot(by_age, additional_stats = "none"),
-  #   path = "SurveyFit"
-  # )
-  #
-  # vdiffr::expect_doppelganger(
-  #   "plot-group-weights",
-  #   fit_stan_glmer$plot(by_age, additional_stats = c("mrp","raw","wtd")),
-  #   path = "SurveyFit"
-  # )
+  #vdiffr::write_svg(fit_stan_glmer$plot(popn, additional_stats = "none"),
+  #                  "tests/figs/SurveyFit/plot-population.svg", title = "plot-population")
+
+  vdiffr::expect_doppelganger(
+    "plot-population",
+    fit_stan_glmer$plot(popn, additional_stats = "none"),
+    path = "tests/figs/SurveyFit/plot-population.svg"
+  )
+
+  #vdiffr::write_svg(fit_stan_glmer$plot(popn, additional_stats = c("mrp","raw","wtd")),
+  #                  "tests/figs/SurveyFit/plot-population-stats.svg", title = "plot-population-stats")
+
+  vdiffr::expect_doppelganger(
+    "plot-population-stats",
+    fit_stan_glmer$plot(popn, additional_stats = c("mrp","raw","wtd")),
+    path = "tests/figs/SurveyFit/plot-population-stats.svg"
+  )
+
+  #vdiffr::write_svg(fit_stan_glmer$plot(by_age, additional_stats  = "none"),
[…]

### 2021-09-15 R/SurveyFit.R
Added warning, changed snapshots of images

@@ -256,6 +256,9 @@ SurveyFit <- R6::R6Class(
         stop("additional statistics can only be weighted (wtd) or raw (raw)")
       }
       additional_ests_filtered <- additional_ests[additional_ests$method %in% additional_stats,]
+      if(sum(self$map()$sample()$weights() ==1)==length(self$map()$sample()$weights())){
+        warning("weights are all equal to 1 or no weights provided. Raw estimate and wtd estimate will be equivalent.")
+      }
       gg <- gg +
         ggplot2::geom_vline(data = additional_ests_filtered, ggplot2::aes(xintercept = .data[["mean"]], colour = method)) +
         ggplot2::theme(legend.position = "bottom",

### 2021-09-15 tests/testthat/test-workflow.R

@@ -93,15 +93,19 @@ test_that("Complete run through w/o weights produces only expected warnings", {
   estimates_by_age <- fit_2$aggregate(poststrat_estimates, by = "age")
 
   # plot estimates by age
-  fit_2$plot(estimates_by_age, weights = FALSE)
-  fit_2$plot(estimates_by_age, weights = TRUE)
+  expect_warning(fit_2$plot(estimates_by_age),
+  "weights are all equal to 1 or no weights provided. Raw estimate and wtd estimate will be equivalent.")
+  expect_warning(fit_2$plot(estimates_by_age, additional_stats = "none"),
+                 'weights are all equal to 1 or no weights provided. Raw estimate and wtd estimate will be equivalent.')
 
   # population estimate
   estimates_popn <- fit_2$aggregate(poststrat_estimates)
 
   # plot population estimate
-  fit_2$plot(estimates_popn, weights = FALSE)
-  fit_2$plot(estimates_popn, weights = TRUE)
+  expect_warning(fit_2$plot(estimates_popn),
+                 'weights are all equal to 1 or no weights provided. Raw estimate and wtd estimate will be equivalent.')
+  expect_warning(fit_2$plot(estimates_popn, additional_stats = "none"),
+                 "weights are all equal to 1 or no weights provided. Raw estimate and wtd estimate will be equivalent.")
 
 })
 

### 2021-09-15 R/SurveyFit.R
Updated doc

@@ -22,6 +22,7 @@ SurveyFit <- R6::R6Class(
     #'   not need to be called directly by the user.
     #' @param fit A fitted model object.
     #' @param map A [`SurveyMap`] object.
+    #' @param formula A formula object representing the model ran.
     initialize = function(fit, map, formula) {
       private$fit_ <- fit
       private$map_ <- map

### 2021-09-15 R/SurveyMap.R

@@ -143,7 +143,7 @@
 #'
 #' # plot estimates by age
 #' fit_1$plot(estimates_by_age)
-#' fit_1$plot(estimates_by_age, additional_stats = "none)
+#' fit_1$plot(estimates_by_age, additional_stats = "none")
 #' fit_1$plot(estimates_by_age, additional_stats = "wtd")
 #' fit_1$plot(estimates_by_age, additional_stats = "raw")
 #' fit_1$plot(estimates_by_age, additional_stats = c("wtd","raw","mrp"))

### 2021-10-04 R/SurveyMap.R
Add the files and deal with the URLs and propose new way of describing the package

@@ -105,6 +105,7 @@
 #'
 #' # Example rstanarm usage
 #' # Returns a SurveyFit object
+#' \dontrun{
 #' fit_1 <- ex_map$fit(
 #'   fun = rstanarm::stan_glmer,
 #'   formula = y ~ (1|age) + (1|gender),
@@ -152,7 +153,7 @@
 #' # plot population estimate
 #' fit_1$plot(estimates_popn, weights = FALSE)
 #' fit_1$plot(estimates_popn, weights = TRUE)
-#'
+#' }
 SurveyMap <- R6::R6Class(
   classname = "SurveyMap",
   private = list(

### 2021-10-08 R/SurveyFit.R
Minor changes to doc and examples as per PR

@@ -156,12 +156,10 @@ SurveyFit <- R6::R6Class(
     #' @description Creates a set of summary statistics of the mrp estimate,
     #' and corresponding weighted and raw data estimates
     #' @param aggregated_estimates The data frame returned by the `aggregate() method`.
-    #' @return A data frame. If `by` is not specified then the data frame will
-    #'   have number of rows equal to the number of posterior draws. If `by` is
-    #'   specified the data frame will have number of rows equal to the number
-    #'   of posterior draws times the number of levels of the `by` variable,
-    #'   and there will be an extra column indicating which level of the `by`
-    #'   variable each row corresponds to.
+    #' @return A data frame that consists of a minimum three rows with the raw, MRP
+    #' and weighted estimates, plus an estimate of standard error. If the aggregated estimates
+    #' were specified with a `by` argument (indicating sub population or small area estimates),
+    #' then produces a dataframe with number of rows equal to three times the number of small areas.
     summary = function(aggregated_estimates) {
       if (!is.data.frame(aggregated_estimates)) {
         stop("'aggregated_estimates' must be a data frame ",
@@ -217,7 +215,8 @@ SurveyFit <- R6::R6Class(
       rownames(out) <- NULL
       out
     },
-    #' @description Plot takes the aggregated MRP estimates and produces a quick visualization of total and sub-population estimates.
+    #' @description Plot takes the aggregated MRP estimates and produces a quick
+    #' visualization of total and sub-population estimates.
     #' @param aggregated_estimates The data frame returned by `aggregate`.
     #' @param additional_stats A vector that specifies which of three additional
     #'   stats ("wtd", "raw", "mrp", "none") should be included on the plot. The
@@ -225,9 +224,11 @@ SurveyFit <- R6::R6Class(
     #'   mean (raw), but an analogous bar for MRP (mrp) can be added using the
     #'   posterior mean and sd. The sd for the weighted estimate uses the survey
     #'   design and the \pkg{survey} package, whilst the raw estimate is a
-    #'   direct mean and binomial sd of the binary responses. Intervals are by
-    #'   default 95% CI.
-    #' @return A ggplot object.
+    #'   direct mean and binomial sd of the binary responses. Uncertainty is included on
+    #'   violin plots but not on density plots. Intervals are 95% CI.
+    #' @return A ggplot object that is either a violin plot (if aggregated estimates is
+    #' at a small area level) or density plot (if aggregated estimates is at the population
[…]

### 2021-10-08 R/SurveyMap.R

@@ -122,12 +122,12 @@
 #' )
 #'
 #' # Example brms usage
-#' fit_3 <- ex_map$fit(
-#'   fun = brms::brm,
-#'   formula = y ~ (1|age) + (1|gender),
-#'   family = "bernoulli",
-#'   seed = 1111
-#' )
+#' # fit_3 <- ex_map$fit(
+#' #   fun = brms::brm,
+#' #   formula = y ~ (1|age) + (1|gender),
+#' #   family = "bernoulli",
+#' #   seed = 1111
+#' # )
 #'
 #'
 #' # predicted probabilities
@@ -139,7 +139,7 @@
 #'
 #' fit_1$summary(estimates_by_age)
 #' fit_2$summary(estimates_by_age)
-#' fit_3$summary(estimates_by_age)
+#'# fit_3$summary(estimates_by_age)
 #'
 #' # plot estimates by age
 #' fit_1$plot(estimates_by_age)

### 2021-12-06 paper/paper.Rmd
Finally start to get on top of this thing. Tomorrow at 2020 US election examples

--- 2020-09-30 paper/paper.Rmd
@@
 ---
-documentclass: jss
+title: "`mrpkit`: A reproducible, opinionated, workflow for multilevel regression with post-stratification in R"
 author:
-  - name: Lauren Kennedy
-    affiliation: Monash University
-    email: \email{name@company.com}
-    url: http://rstudio.com
-  - name: Mitzi Morris
-    affiliation: Columbia University
-  - name: Jonah Gabry
-    affiliation: Columbia University
-  - name: Rohan Alexander
-    affiliation: University of Toronto
-title:
-  formatted: "\\pkg{MRP-Kit} A 'grammar' of multilevel regression with post-stratification and implementation in \\proglang{R}"
-  # If you use tex in the formatted title, also supply version without
-  plain:     "A 'grammar' of multilevel regression with post-stratification and implementation in R"
-  # For running headers, if needed
-  short:     "\\pkg{foo}: A grammar of MRP"
-abstract: >
-  In this paper we define a grammar for multilevel regression with post-stratification (MRP) and implement this grammar in the R Package \pkg{MRPKit}.
-Our grammar is centered around the following verbs: add, remove, delete, and replace.
-These act on survey objects.
-Our grammar, and its implementation, provides a detailed workflow when conducting MRP that will be useful for researchers and in teaching.
-keywords:
-  # at least one keyword must be supplied
-  formatted: [multilevel regression with post-stratification, "\\proglang{R}", reproducibility, statistics, political science]
-  plain:     [multilevel regression with post-stratification, R, reproducibility, statistics, political science]
-preamble: >
-  \usepackage{amsmath}
-output: rticles::jss_article
+  - Lauren Kennedy^[Monash University, lauren.kennedy729@gmail.com]
+  - Jonah Gabry^[Columbia University]
+  - Dewi Amaliah^[Monash University]
+  - Mitzi Morris^[Columbia University]
+  - Rohan Alexander^[University of Toronto]
+
+thanks: "**Feedback:** Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper are welcome at: lauren.kennedy729@gmail.com.
+**Funding statement:** We have no direct financial disclosures.
+**Conflicts statement:** We have no conflicts to disclose.
+**Code and data accessibility statement:** Code and data are available at: https://github.com/lauken13/mrpkit.
+We are unable to share direct access to the datasets used in the 2020 US Presidential example in this paper, however they are freely available and we provide detailed instructions for how to obtain them.
+**Author contributions:** Kennedy had the original idea.
+Gabry, Kennedy, and Morris made architectural and other high-level decisions.
+Gabry and Kennedy wrote most of the code and tests in the package.
+Amaliah wrote the vignette and contributed tests.
+Alexander wrote the first draft of this paper.
+All authors contributed to writing the paper, and approved the final version."
+date: "`r format(Sys.time(), '%d %B %Y')`"
+abstract: |
+    The R package `mrpkit` implements a structured workflow for Multilevel Regression and Poststratification (MRP).
+The package helps appropriately set-up the survey and post-stratification datasets and then establish a relationship between different variables in these datasets.
+A substantial amount of the work to do this is specified by the methods in the package, saving time, enhancing reproducibility, and reducing the risk of coding errors.
+The package natively supports multilevel binomial and Bernoulli models fit with `lme4` and `Stan` (via `brms` and `rstanarm`) and allows for the use of custom modeling functions.
+Summaries and visualizations of the resulting post-stratified estimates can be conducted using the methods in the package.
+This implementation provides a detailed workflow when conducting MRP that will be useful for researchers and in teaching.
+Unlike writing all the code yourself, using `mrpkit` proactively addresses many common issues, and makes it possible for folks who are new to MRP to quickly conduct their first analysis.
+
+    **Keywords:** multilevel regression with post-stratification, R, reproducibility, statistics.
+output:
+  bookdown::pdf_document2:
+toc: FALSE
+bibliography: references.bib
 ---
 
-```{r, setup, include=FALSE}
-options(prompt = 'R> ', continue = '+ ')
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
 ```
 
-\section[Introduction]{Introduction} \label{sec:intro}
+# Introduction
 
 Multilevel regression with post-stratification (MRP) is a statistical approach in which surveys are related to each other using a statistical model.
-This is important because known biases in one survey, can be adjusted for by another in a statistically reasonable way.
-This enables better use of non-representative surveys, additional information, and propagation of uncertainty.
-However it can be difficult to use MRP due to this need to related two different datasets.
-This package defines a grammar, or list of underlying rules, of MRP and then describes an R package, \proglang{MRPKit}, that implements this grammar.
+This is useful because known biases in one survey can be adjusted by another in a statistically reasonable way.
+This enables better use of non-representative surveys, consideration of additional information, and for the propagation of uncertainty.
+However, it can be difficult to use MRP due to this need to relate two different datasets.
+This seemingly straight-forward requirement imposes considerable hassle in practice and tends to substantially complicate reproducibility.
+The R [@citer] package `mrpkit` allows users to conduct multilevel regression with post-stratification in an opinionated and specific way.
+The package enables users to specify two datasets (a sample dataset and a post-stratification dataset), establish a correspondence between them, and then conduct analysis.
+
+The process of conducting MRP involves considering some survey outcome as being due to a combination of many different cells that are typically demographic or geographic in nature, estimating this relationship using a multilevel regression model and then aggregating cell-level estimates to match the population [@wang2015forecasting].
+The survey of interest would typically be a regular survey, such as a political poll of, say, 1,000 respondents, but it could also be a larger survey, such as the American National Election Studies, or similar.
+The post-stratification dataset would typically be a larger survey, such as, in the case of the US, the American Community Surveys, or a census.
+The correspondences need to be established for each possible answer for each question, for instance, the age-groups used in both surveys need to be homogenized.
+The model would typically be a Bayesian hierarchical model, but in principle any variety could be used.
 
 At its core, MRP is a mapping between a survey object and a population object.
-It is from this mapping that the power of MRP exists, but it also establishing this mapping that is the difficult part of implementing MRP models.
-Making this implementation easier and more reproducible is important as interest in, and the use of, MRP increases.
-It can be difficult even for those experienced with MRP to ensure there are no mistakes in this mapping and easing this is an important contribution to enhancing the reproducibility of MRP analysis.
+It is from this mapping that the power of MRP exists, but it can be difficult to establish this mapping in a reproducible way that fits into a typical applied statistics workflow.
+Making this implementation easier and more reproducible is important because interest in, and the use of, MRP has considerably increased.
+It can be difficult even for those experienced with MRP to ensure there are no mistakes in this mapping.
+The `mrpkit` package creates a well-defined workflow for MRP that represents an important contribution for enhancing the reproducibility of MRP analysis.
 
-We first define a grammar of MRP, which we define as the underlying rules and principles that are common to every analysis based on MRP.
-This grammar is based around the following verbs: 'add', 'remove', 'delete', and 'replace'.
-These verbs are applied to a survey object, a post-stratification object, and survey\_map object.
-These three objects come together to create a mapping object, which is what processes such as regression act on.
-Finally, common diagnostics and graphs are implemented.
-In this way, our grammar and package implement an entire statistical workflow for conducting MRP.
+Our package complements existing packages such as `survey` [@lumleyone; @lumleytwo; @lumleythree], and `DeclareDesign` [@citeDeclareDesign].
+The `survey` package is well-established and allows a user to specify a survey design, such as clustered, stratified, etc, generate summary statistics, and conduct sub-population analysis.
+The `srvyr` package [@citesrvyr] is a variant of `survey` that adds a syntax similar to `dplyr` [@citedplyr].
+The `DeclareDesign` package [@citeDeclareDesign] is a newer package that is based around a grammar of research design aspects including: data generation, creating treatment and control groups, and sampling.
+In general, these packages are focused on the designing, implementing, and simulating from, surveys.
+Ours is focused on what is needed for MRP, and we draw on their packages where possible.
 
-The survey object would typically be a regular survey, such as a political poll of 1,000 respondents, but it could also be a larger survey, such as the Canadian Election Survey, or similar.
-The post-stratification object would typically be a larger survey, such as, in the case of the US, the (INSERT THE USUAL ONE), or a census.
+In terms of packages to specifically help with MRP, the most common alternative at the moment to this package is for users to do all aspects of the MRP workflow themselves.
+While there is nothing inherently difficult about MRP, the implementation can be difficult.
+In particular, preparing and matching different levels between surveys can be time consuming and potentially introduce undocumented errors.
+More recently `tidymrp` [@jospackage] is being developed.
+This package focuses on conducting analysis in a tidy way [@citetidyverse].
+And `ccesMRPprep` [@shirospackage] is focused on using the Cooperative Congressional Election Study (CCES).
 
-Our grammar and package complements existing packages such as \proglang{survey} (ADD CITATION), X and Y. These packages... Instead, ours is focused on what is needed for MRP.
+The main alternative approach is for MRP to be conducted on a case-specific basis.
+This means that a researcher who is interested in MRP estimates writes the code needed to obtain, clean, prepare, analyze and ultimately interpret the estimates.
+Again, while there is nothing inherently wrong with this approach, successfully conducting MRP requires dealing with many small issues.
+Each of these is small but getting them wrong can have large effects that are potentially unnoticed on the estimates.
+It is also easy to overlook steps, or to not appropriately document them.
+`mrpkit` is prescriptive about the steps that must be taken, and ensures that all aspects are documented.
 
-The remainder of this paper is structured as follows: Section \ref{sec:review} reviews similar packages and contributions and places ours within that context.
-Section \ref{sec:components} discusses the grammar and the core aspects of MRP as implemented in \proglang{MRPKit}.
-Section \ref{sec:implementation} discusses some of the implementation issues and technical notes related to the decisions that were made.
-Section \ref{sec:vignette} provides two examples of the package in use, one using SOMETHING, and the other using SOMETHING ELSE.
-Finally, Section \ref{sec:summary} provides a summary discussion, some cautions and weaknesses, as well as notes about next steps.
+The remainder of this paper is structured as follows: Section \@ref(usecase) discusses the core use case and workflow for MRP as implemented in `mrpkit`.
+Section \@ref(implementation) discusses some of the implementation issues and technical notes related to the decisions that were made.
+Section \@ref(example) provides an example of the package in use.
+Finally, Section \@ref(summary) provides a summary discussion, some cautions, and weaknesses, as well as notes about next steps.
 
-\section{Review of the other packages} \label{sec:review}
+# Use cases and workflow {#usecase}
 
-The most common alternative at the moment to this package is for users to do all aspects themselves.
-While there is nothing conceptually difficult about MRP, the implementation can be difficult.
-In particular, preparing and matching different levels between surveys can be time consuming and potentially introduce errors.
+The MRP workflow implemented by `mrpkit` is summarised in Figure \@ref(fig:illustrate).
+In this section, these steps are expanded on.
+At each of them there are a variety of decisions that must be made.
+The `mrpkit` package makes each of these explicit.
 
-HERE
+```{r illustrate, warning = FALSE, echo=FALSE, fig.cap="Illustration of a `mrpkit` workflow"}
+library(DiagrammeR)
 
-Survey package
+grViz("
+digraph a_nice_graph {
 
-Declare design
+graph [compound = true]
 
-DIY
+node [shape = box, fontsize = 12]
 
-MRP package - https://github.com/gelman/mrp
+subgraph cluster0 {
+    node [fixedsize = true, width = 3]
+    'Read in datasets' -> 'Prepare each dataset' -> 'Bring the datasets together' -> 'Fit the model' -> 'Generate outputs and summarize results'
+  }
+}
+")
+```
 
-And how we're different
+We see a variety of use cases including:
 
-\section{Components and grammar} \label{sec:components}
+- A professor wants to introduce MRP to a class that might not be especially experienced with R, nor have time to develop the necessary skills, instead wanting to focus on the estimates and applications.
+- A researcher with domain or survey expertise, but who does not have the resources to ensure that the code has been appropriately tested.
+- A researcher who is experienced with MRP needs to run many different models on the same dataset.
 
-The \proglang{MRPKit} package has the following key components: survey objects, which for most users will be a collection of two surveys where one is larger than the other; a survey map, that relates the survey objects and then an MRP object, which is created once the survey map is applied to the survey objects.
-Regression acts on the MRP object.
-These objects are subject to the following verbs: 'add', 'delete', 'new', and 'replace'.
-For instance: \proglang{SurveyMap\$add}, \proglang{SurveyOb\$delete}, \proglang{SurveyMap\$new}, and \proglang{SurveyMap\$replace}.
+## Set-up, load package, and read in the datasets
 
-\subsection*{Survey objects}
+At the moment the `mrpkit` package has not been submitted to CRAN as it is still be finalised. For now it can be installed from GitHub.
 
-Survey objects are the surveys that MRP will bring together.
-A user can \proglang{add} a survey object by providing a CSV file location.
-The user then needs to identify the data types of each column.
-Typically, there will need to be two survey objects, where one will be the survey that is of interest for its response variables, such as political opinion, and another will be a survey that is to be used for post-stratification.
+```{r}
+# devtools::install_github("lauken13/mrpkit")
+library(mrpkit)
+```
 
-New is when you have a new survey object and you are bringing it in. e.g.
+Having installed and loaded the package, a user first needs to get their survey and post-stratification dataset into R.
+This means ensuring that it is in a rectangular format such as a CSV and then reading it in as a dataframe.
+The `mrpkit` package does not explicitly contain methods for this, instead a user could use `readr::read_csv()` [@citereadr] or `haven::read_dta()` [@citehaven].
 
-% sampl <- surveyStr$new(...)
-% pop <- surveyStr$new(...)
-% map <- surveyMap$new(sampl, pop, ...)
-% mrp1 <- MRP$new(sampl, pop, map)
+A user will need two datasets.
+At this point, one aspect to be aware of is the class of the variables.
+The `mrpkit` package will convert all variables to factor variables.
+Here for the survey and post-stratification datasets we will use simulated datasets that are available as part of the package `mrpkit::shape_survey` and `mrpkit::approx_voters_popn`.
 
-% option 1
-% mod1 <- mrp1$fit(y ~ x + (1|group))
-% mod1 <- mrp1$fit(y ~ x + (1|group) + (1|group2))
+```{r}
+survey <- shape_survey
+poststrat <- approx_voters_popn
 
-% mod1$poststratify()
-% mod2$poststratify("model1")
+head(survey)
+head(poststrat)
+```
 
-% or
+## Preparing each dataset
 
-% option 2
-% mrp1$fit(y ~ x + (1|group), name = "model1")
-% mrp1$fit(y ~ x + (1|group) + (1|group2), name = "model2")
+Typically MRP requires at least two datasets---usually a survey dataset and a post-stratification dataset.
+A user specifies these using the `SurveyData` method.
+`SurveyData` objects are used to represent both the survey and the post-stratification datasets, along with their metadata.
+A user needs to create `SurveyData` objects for both the survey and post-stratification datasets separately using the `SurveyData$new` method and assigning a name to each.
+The `SurveyData$new` method is used to transform a regular data frame into a `SurveyData` object.
 
-% mrp1$poststartify("model1")
-% mrp1$poststartify("model2")
+As part of creating these objects a user will typically specify the dataset (from the earlier step), the survey questions, the possible responses, and any weights and design formula that are applicable.
+A user does this separately for the survey and the post-stratification dataset, so there will be two objects created.
+Having created this object, a user can use the `SurveyData$print` method to display the created `SurveyData` object in the console.
 
-What is the difference between add and new? What is replace doing? URGH, I just need to look at the repo.
+Building on the datasets earlier, we will show the steps to create `SurveyData` objects using the `SurveyData$new` method. This method takes the following arguments:
 
-\subsection*{Survey map}
+1.
+`data`: The name of the dataframe, i.e., the survey or post-stratification data, that is to be transformed to a `SurveyData` object.
+By default, this method uses each of the factor, character, and binary variables to create questions and responses, unless a user specifies the list of questions and responses.
+2.
+`questions`: The actual questions asked in the survey form or questionnaire for each of the columns in the dataframe.
+For instance, a column called 'age' may contain the results of a question 'Please identify your age-group'.
+3. `responses`: All possible responses for each of the questions. For instance, possible responses in the 'age' column could be: '18-25', '26-35', '36-45', '46-55', '56-65', '66-75', and '76-90'.
+4. (Optional) `weights`: If the weight is specified as a column in the datasets, then this argument should be specified as a string.
+5. (Optional) `design`: If there is a design formula for the survey, then this can be specified using the notation of the `survey` package [@lumleyone; @lumleytwo; @lumleythree].
 
-A `SurveyMap` object holds the mapping between a set of items in a survey and a population dataset.
-The label is the item label in each dataset and the values is a list of all possible values.
-The values for the survey and population must be aligned, i.e., the lists must have the same number of elements and the values at index i in each list are equivalent.
-If there is a meaningful ordering over the values, they should be listed in that order, either descending or ascending.
+Here we create a `SurveyData` object for the survey dataset.
 
-One of the fundamental issues when conducting MRP is to ensure that
+```{r survey-object}
+# Create a SurveyData object for the survey dataset 
+box_pref <- SurveyData$new(
+  data = survey,
+  questions = list(
+    age = "Please identify your age group",
+    gender = "Please select your gender",
+    vote_for = "Which party did you vote for in the 2018 election?",
+    highest_educ = "Please identify your completed highest education",
+    state = "Which State do you live in?",
+    y = "If today is election day, will you vote for the Box Party?"
+  ),
+  responses = list(
+    age = levels(survey$age),
+    gender = levels(survey$gender),
+    vote_for = levels(survey$vote_for),
+    highest_educ = levels(survey$highest_educ),
+    state = levels(survey$state),
+    y = c("no", "yes")
+  ),
+  weights = "wt",
+  design = list(ids =  ~ 1)
+)
 
-\subsection*{MRP object}
+# Display it
+box_pref$print()
+```
 
-An  MRP object contains survey objects, and a survey map. At the point at which the survey objects are put into the MRP object they become immutable but the survey map object is.
+Now we create a `SurveyData` object for the post-stratification dataset.
+This code is used to create a `SurveyData` object for post-stratification data when the weight is already specified in the dataset.
+Usually, the post-stratification data is from a large survey, for example, the American Community Survey (ACS) or Demographic and Health Survey (DHS).
+It is possible to have an already summarized post-stratification dataset.
+If this is the case, then the weight would be the size of each cell.
+Alternatively (as in this example), if the entire individual-level population is given, then this argument should be omitted, and it would be automatically specified as 1.
 
-The MRP object outputs an analysis. There would be a MRP model fit class that would
+```{r poststrat-object}
+poststrat_obj <- SurveyData$new(
+  data = poststrat,
+  questions = c(
+    age_group = "Which age group are you?",
+    gender = "Which gender are you identified?",
+    vote_pref = "Which party do you prefer to vote?",
+    education = "What is the highest grade or level of school you have completed",
+    state = "Please identify the state where you live in"
+  ),
+  responses = list(
+    age_group = levels(poststrat$age_group),
+    gender = levels(poststrat$gender),
+    vote_pref = levels(poststrat$vote_pref),
+    education = levels(poststrat$education),
+    state = levels(poststrat$state)
+  ),
+  weights = "wt",
+  design = list(ids =  ~ 1)
+)
 
-Many objects on the same class. There are two
+# print it
+poststrat_obj$print() 
+```
 
-\section{Implementation and technical notes} \label{sec:implementation}
+A user can access the properties of the `SurveyData` object using methods for this.
+For instance, the weights, `SurveyData$weights()`, the design, `SurveyData$design()`, the questions, `SurveyData$questions()`, and the number of questions, `SurveyData$n_questions()`, or the responses, `SurveyData$responses()`.
 
-\section{Vignette} \label{sec:vignette}
+## Aligning the datasets
 
-\subsection*{Using CCES data}
+Having established these two datasets separately, a user needs to begin to create a relationship between them.
+The first step is to relate each question, and their answers using `QuestionMap$new`.
+This needs to be done for every question that will be used and typically requires specifying a name that will be used for the underlying construct, the column names in each of the two `SurveyData` objects, and a mapping between the values in each.
+This is a deliberately specific task requiring the user to think about and manually link the answers.
 
-\subsection*{Using your own data}
+In the example, the survey (`box_pref`) and the post-stratification (`poststrat_obj`) objects have different column labels and response levels.
+Specifically, the column label for age in `box_pref` is 'age', whereas, in `poststrat_obj`, the column label is 'age_group'.
+This variable also has different levels; `box_pref` has seven levels of age, whereas `poststrat_obj` has only four levels of age.
+We do this alignment using the `QuestionMap$new` method.
 
-\section{Summary and discussion} \label{sec:summary}
+This method takes three arguments:
 
-\subsection*{Next steps and cautions}
+1. `name`: The name of the underlying construct. This will be used in the modeling stage.
+2.
+`col_names`: A character vector of the column label in the survey and post-stratification objects.
+The column name of the survey should always be the first element, followed by the column name of the post-stratification object.
+This order is not interchangeable.
+3.
+`values_map`: The list of the mapped values between the survey and post-stratification object.
+If there is a meaningful ordering over the values, they should be sorted over that order, either descending or ascending.
 
-\section*{Acknowledgments}
+Here we define how age-groups, political preferences, gender, education, and states should come together.
+While this is tedious it is worth noting that in general much of this code is likely to be re-useable across different situations.
+
+```{r questions}
+# Create QuestionMap$object for the question related to age
+q_age <- QuestionMap$new(
+  name = "age",
+  col_names = c("age", "age_group"),
+  values_map = list(
+    "18-25" = "18-35",
+    "26-35" = "18-35",
+    "36-45" = "36-55",
+    "46-55" = "36-55",
+    "56-65" = "56-65",
+    "66-75" = "66+",
+    "76-90" = "66+"
+  )
+)
+
+# Create QuestionMap$object for the question related to party preference
+q_party_pref <- QuestionMap$new(
+  name = "party_pref",
+  col_names = c("vote_for", "vote_pref"),
+  values_map = list(
+    "Box Party" = "BP",
+    "BP" = "BP",
+    "Circle Party" = "CP",
+    "CP" = "CP"
+  )
+)
+
+# Create QuestionMap$object for the question related to gender
+q_gender <- QuestionMap$new(
+  name = "gender",
+  col_names = c("gender", "gender"),
+  values_map = data.frame(
+    "male" = "m",
+    "female" = "f",
+    "nonbinary" = "nb"
+  )
+)
+
+# Create QuestionMap$object for the question related to education
+q_educ <- QuestionMap$new(
+  name = "highest_education",
+  col_names = c("highest_educ", "education"),
+  values_map = list(
+    "no high school" = "no high school",
+    "high school" = "high school",
+    "some college" = "some college",
+    "associates" = "some college",
+    "4-year college" = "4-years college",
+    "post-graduate" = "post-grad"
+  )
+)
+
+# Create QuestionMap$object for the question related to state
+q_state <- QuestionMap$new(
+  name = "state",
+  col_names = c("state", "state"),
+  values_map = list(
+    "State A" = "A",
+    "State B" = "B",
+    "State C" = "C",
+    "State D" = "D",
+    "State E" = "E"
+  )
+)
+```
+
+The maps between each question print nicely to allow checking.
+
+```{r}
+q_age
+q_party_pref
+q_gender
+q_educ
+q_state
+```
+
+## Bringing the datasets together
+
+The relationship between the two datasets is finalised by bringing all of this together to create a map between them using `SurveyMap$new` and assigning this to an object.
+This typically requires specifying:
+
+1) the name of the `SurveyData` object related to the sample, in this case 'box_pref'; 
+2) the name of the `SurveyData` object related to the post-stratification dataset, in this case `poststrat_obj`; and 
+3) the question mappings that should be included, which specifies the matched labels and values provided by `QuestionMap` objects.
+
+<!-- Probs get rid of: -->
+<!-- The `SurveyMap` object has `new`, `add`, `delete`, `replace`, `mapping`, `tabulate`, and `fit` methods. -->
+
+The method initializes a new `SurveyMap` object is `SurveyMap$new`. This takes `SurveyData` and `QuestionMap` objects as argument.
+
+```{r map-object}
+# Create a new SurveyMap object
+ex_map <- SurveyMap$new(sample = box_pref, population = poststrat_obj,
+                        q_age, q_educ, q_gender, q_party_pref
+                        )
+
+print(ex_map)
+```
+
+Questions can also be added using `SurveyMap$add`. For instance, a user could add 'state' to the `QuestionMap` object.
+
+```{r add}
+# Add questions incrementally
+ex_map$add(q_state)
+
+print(ex_map)
+```
+
+The user has mapped all the questions in the post-stratification object.
+Variables could be excluded using the `SurveyMap$delete` method.
+Similarly, `SurveyMap$replace` enables a user to use another level of matching.
+
+## Finalising the mapping and tabulation
+
+At this point, the datasets have been brought together.
+This process is finalised with `SurveyMap$mapping`.
+This creates new sample and post-stratification datasets with unified variable names, and levels.
+It prepares the mapped data for model fitting.
+
+```{r mapping}
+ex_map$mapping()
+```
+
+If a user wants to include all the variables in the post-stratification step, then a user uses the method `SurveyMap$tabulate` without any arguments.
+This prepares the dataset for post-stratification.
+It is also possible to specify only certain variables in the post-stratification matrix, by including those variables as arguments.
+
+```{r tabulate}
+ex_map$tabulate()
+```
+
+## Modelling the survey data
+
+The user can then fit a model to the survey dataset using `SurveyMap$fit` and naming it to create an object.
+Built-in options include: `rstanarm::stan_glmer()` and `rstanarm::stan_glm()` [@citestanarm], `lme4::glmer()` [@citelme4] , `brms::brm()` [@citebrms].
+A user could also specify their own function.
+For that purpose there is a `data` argument that accepts a dataframe.
+In this example, we fit three models using `rstanarm::stan_glmer()`, `lme4::glmer()`, and `brms:brm()` with `age`, `gender`, and `education` as the predictors.
+
+```{r  fitting}
+# Example of using rstanarm::stan_glmer
+fit1 <- ex_map$fit(
+  fun = rstanarm::stan_glmer,
+  formula = y ~ (1 | age) + (1 | gender) + (1 | state),
+  family = "binomial",
+  refresh = 100,
+  cores = 2)
+
+# Example of using lme4::glmer
+fit2 <- ex_map$fit(
+  fun = lme4::glmer,
+  formula = y ~ (1 | age) + (1 | gender) + (1 | state),
+  family = "binomial")
+```
+
+```{r fit-brm, eval = FALSE}
+# Example using brms::brm
+fit3 <- ex_map$fit(
+  fun = brms::brm,
+  formula = y ~ (1 | age) + (1 | gender) + (1 | state),
+  family = "bernoulli",
+  refresh = 100,
+  cores = 2)
+```
+
+The resulting object is not the fitted model object created by the modeling function, but rather a `SurveyFit` object that contains the fitted model and provides useful methods for working with it.
+
+## Using that model
+
+At this stage, a user has fitted model objects, i.e., `fit1` and `fit2`.
+These objects have been automatically stored as `SurveyFit` objects.
+`mrpkit` has various built-in methods that enable a user to perform common tasks using the `SurveyFit` object.
+These include:
+
+1) Generating the predicted probabilities for the post-stratification cells using `SurveyFit$population_predict`.
+2) Considering these post-stratified estimates at a more-aggregated level, for instance state or province level, using `SurveyFit$aggregate`.
+3) Make a table that summarises the posterior mean and standard deviation, along with comparisons with raw estimates and weighted estimates, using the `summary` method.
+4) Visualise the estimates using the `plot` method.
+
+### Get the probability of the outcome for each poststratification cell
+
+After creating the `SurveyFit` object, the user can generate the predicted probability of the outcome of each post-stratification cell using the `SurveyFit$population_predict` method.
+This returns a matrix with rows that correspond to the columns of post-stratification data and columns that correspond to the posterior samples.
+
+Furthermore, if the model fitting is done with one of the built-in functions (`rstanarm::stan_glmer`, `lme4::glmer`, `brms::brm`), then this method does not take any arguments.
+However, if you used a customized function in the model fitting, then this method takes multiple arguments, including a custom prediction function
+(see `?SurveyFit` for details). In this example, we will show how to use this method for the `fit1` object created using `rstanarm::stan_glmer`.
+
+```{r predict}
+# predict the probability of voting for the Box Party using the fit1 model
+poststrat_est_fit1 <- fit1$population_predict()
+```
+
+### Aggregate the probability of the outcome to a higher level of estimate
+
+The next step is to generate the population estimate or group estimate using the `SurveyFit$aggregate` method.
+This method takes two arguments, namely the post-stratification estimate data and the variable whose level to which the estimated value would be aggregated to.
+In this example, we want to aggregate the estimated value by the level of `age` and `state`.
+If the variable is not specified, then this method would generate a population estimate.
+
+```{r aggregate}
+
+# aggregate the predicted value by age
+age_estimation <- fit1$aggregate(poststrat_est_fit1, by = "age")
+
+# aggregate the predicted value by state
+state_estimation <- fit1$aggregate(poststrat_est_fit1, by = "state")
+
+# generate the population estimate
+popn_estimation <- fit1$aggregate(poststrat_est_fit1)
+
+```
+
+This method then will return a data frame with the variable levels and probability of the outcome, `y`.
+In this example, `age_estimation` represents the age group and the probability of voting for the Box Party for each age group.
+
+```{r show-aggr}
+head(age_estimation)
+
+# get the mean and sd for each age group
+library(dplyr)
+age_estimation %>%
+  group_by(age) %>%
+  summarize(mean = mean(value), sd = sd(value))
+```
+
+### Summarise and compare
+
+The output of aggregate is a posterior predictive distribution for the population (or sub-populations) of interest.
+This can be used with the plot function to visualize.
+However, we also might wish to make a table summary of the posterior mean and standard deviation, along with comparisons with raw estimates and weighted estimates.
+We can use the `summary` method do to this.
+This method takes the aggregated data frame and produces a dataframe of mean and standard deviation estimates for each group (if in the aggregated frame) or overall in the population.
+The standard deviation for the weights is created by wrapping the survey package using the original design specified, and the raw standard deviation is $\sqrt\frac{pq}{n}$.
+This is called automatically in the plot method.
+
+### Visualize
+
+Once we have the aggregated estimates, we can visualize them using the `plot` method.
+This method takes the aggregated data frame and additional_stats option (the default is wtd and raw estimates) as its arguments.
+It generates a violin plot of the aggregated estimate by the level of a certain variable.
+When additional statistics are specified (options are raw, wtd and mrp; more than one can be specified as a vector), these are added as points with error bars representing the 95% confidence interval.
+In this example, we display the estimated plot for `age` and `state` and a density plot for the population estimates.
+
+```{r age-vis, fig.width= 6, fig.height=4}
+plot_age <- fit1$plot(age_estimation) 
+plot_age
+```
+
+The plot above shows the distribution of people who vote for the Box Party for each age group. The plot implies that people who are aged 36-55 are most likely to vote for the BP.
+
+The next plot displays the probability of voting for the BP by the state without using additional statistics.
+We learn that the probability of the BP winning the election is high since most people in almost all the states (around 75%) vote for that party.
+State B is the state with the lowest probability of vote for the BP.
+The distribution still ranged more than 50%.
+Note that this violin plot is generated using `ggplot2` (Wickham, 2016).
+Hence, it could be modified the same way `ggplot2`'s plot is modified.
+For example, here, we can add the title and change the theme of the plot.
+
+```{r state-vis, fig.width= 6, fig.height=4}
+library(ggplot2)
+plot_state <- fit1$plot(state_estimation, additional_stats = "none") +
+  ggtitle("Probability of voting the BP by state") +
+  theme_bw()
+plot_state
+```
+
+Lastly, we show the plot of the population estimate, with lines representing the mean weighted and raw estimates.
+About 70 per cent of people in the Shape World will be likely to vote for the Box Party.
+
+```{r,popn_estimate, fig.width= 6, fig.height=4}
+fit1$plot(popn_estimation, additional_stats = c("raw","wtd"))
+```
+
+# Implementation and technical notes {#implementation}
+
+# Worked example {#example}
+
+# Summary and discussion summary {#summary}
+
+## On the importance of scaffolding for workflows
+
+It is increasingly recognized that workflows are amongst the most important aspect of statistical practice.
+In the context of introducing a Bayesian workflow @gelman2020bayesian says it is important because of factors including: computational challenges, model evolution, and model comparison.
+In the case of MRP we have similar concerns and the scaffolding provided by `mrpkit` is important.
+
+Computational challenges mean that it is time-consuming to run models. This means that it is important to be  and
+
+The analysis often shifts focus which requires
+
+The variety of skills
+
+## The use of R6 objects
+
+## Next steps and cautions
+
+We expect that more papers will be
+
+\newpage
+
+# References

### 2021-12-07 paper/paper.Rmd
Add the US election example to the paper

@@
 
-<!-- Probs get rid of: -->
-<!-- The `SurveyMap` object has `new`, `add`, `delete`, `replace`, `mapping`, `tabulate`, and `fit` methods. -->
-
 The method initializes a new `SurveyMap` object is `SurveyMap$new`. This takes `SurveyData` and `QuestionMap` objects as argument.
@@
 It generates a violin plot of the aggregated estimate by the level of a certain variable.
-When additional statistics are specified (options are raw, wtd and mrp; more than one can be specified as a vector), these are added as points with error bars representing the 95% confidence interval.
-In this example, we display the estimated plot for `age` and `state` and a density plot for the population estimates.
+When additional statistics are specified (options are raw, wtd and mrp; more than one can be specified as a vector), these are added as points with error bars representing the 95 per cent confidence interval.
+In this example, we display the estimated plot for 'age' and 'state' and a density plot for the population estimates.
 
@@
 State B is the state with the lowest probability of vote for the BP.
-The distribution still ranged more than 50%.
-Note that this violin plot is generated using `ggplot2` (Wickham, 2016).
+The distribution still ranged more than 50 per cent.
+Note that this violin plot is generated using `ggplot2` [@citeggplot].
 Hence, it could be modified the same way `ggplot2`'s plot is modified.
@@
 
+There are a variety of technical notes related to this package and the architectural decisions that were made.
+
+Section \@ref(implementation) discusses some of the implementation issues and technical notes related to the decisions that were made
+
 # Worked example {#example}
+
+In this example we will forecast the 2020 US presidential election using survey data from the Democracy Fund Voter Study Group and use the Integrated Public Use Microdata Series (IPUMS), to access the 2018 American Community Survey (ACS) to use as post-stratification data.
+We are interested in considering the following variables: state, age-group, and gender.
+
+The first step is to gather the datasets.
+For the survey data we are interested in the Nationscape dateset, which can be requested from the Voter Study Group.
+Access is usually granted within hours and is completely free.
+We are going to use the survey from 25 June 2020.
+
+```{r}
+# This code based on that from Alen Mitrovski, Xiaoyan Yang, and Matthew Wankiewicz.
+library(haven)
+library(tidyverse)
+
+raw_nationscape_data <- 
+  read_dta(here::here("paper/ns20200625.dta"))
+
+# The Stata format separates labels so reunite those
+raw_nationscape_data <- 
+  labelled::to_factor(raw_nationscape_data)
+
+# Just keep relevant variables
+nationscape_data <- 
+  raw_nationscape_data %>% 
+  select(vote_2020,
+         gender,
+         state,
+         age)
+
+# For simplicity, remove anyone undecided or planning to vote for someone other than Biden/Trump and make vote a binary variable: 1 for Biden, 0 for Trump.
+nationscape_data <- 
+  nationscape_data %>% 
+  filter(vote_2020 == "Joe Biden" | vote_2020 == "Donald Trump") %>% 
+  mutate(vote_biden = if_else(vote_2020 == "Joe Biden", 1, 0)) %>% 
+  select(-vote_2020)
+
+# Create the dependent variables by grouping the existing variables
+nationscape_data <-
+  nationscape_data %>%
+  mutate(
+    age_group = case_when( # case_when works in order and exits when there's a match
+      age <= 29 ~ 'age_18-29',
+      age <= 44 ~ 'age_30-44',
+      age <= 59 ~ 'age_45-59',
+      age >= 60 ~ 'age_60_or_more',
+      TRUE ~ 'Trouble'
+      )
+#     gender = case_when(
+#       gender == "Female" ~ 'female',
+#       gender == "Male" ~ 'male',
+#       TRUE ~ 'Trouble'
+#       ),
+#     education_level = case_when(
+#       education == "3rd Grade or less" ~ "High school or less",
+#       education == "Middle School - Grades 4 - 8" ~ "High school or less",
+#       education == "Completed some high school" ~ "High school or less",
+#       education == "High school graduate" ~ "High school or less",
+#       education == "Other post high school vocational training" ~ "Some post secondary",
+#       education == "Completed some college, but no degree" ~ "Some post secondary",
+#       education == "Associate Degree" ~ "Post secondary or higher",
+#       education == "College Degree (such as B.A., B.S.)" ~ "Post secondary or higher",
+#       education == "Completed some graduate, but no degree" ~ "Post secondary or higher",
+#       education == "Masters degree" ~ "Graduate degree",
+#       education == "Doctorate degree" ~ "Graduate degree",
+#       TRUE ~ 'Trouble'
+#       )
+    ) 
+#   select(-education, -age)
+
+# Tidy the class
+nationscape_data <-
+  nationscape_data %>%
+  mutate(across(c(gender, state, age_group, vote_biden), as_factor))
+
+nationscape_data %>% 
+  head()
+```
+
+```{r}
+# This code is very directly from Alen Mitrovski, Xiaoyan Yang, and Matthew Wankiewicz.
+# Format state names so the whole state name is written out, to match IPUMS data
+# states_names_and_abbrevs <- 
+#   tibble(stateicp = state.name, state = state.abb)
+# 
+# nationscape_data <-
+#   nationscape_data %>%
+#   left_join(states_names_and_abbrevs)
+# 
+# rm(states_names_and_abbrevs)
+# 
+# # Make lowercase to match IPUMS data
+# nationscape_data <- 
+#   nationscape_data %>% 
+#   mutate(stateicp = tolower(stateicp))
+# 
+# nationscape_data$stateicp %>% unique()
+# 
+# # Replace NAs with DC
+# nationscape_data$stateicp <- 
+#   replace_na(nationscape_data$stateicp, "district of columbia")
+# 
+# # Tidy the class
+# nationscape_data <- 
+#   nationscape_data %>% 
+#   mutate(across(c(gender, stateicp, education_level, age_group), as_factor))
+# 
+# # Save data
+# write_csv(nationscape_data, "outputs/data/polling_data.csv")
+
+```
+
+We also need a post-stratification dataset. In this case we can use the American Community Survey (ACS) accessed through IPUMS.
+
+```{r}
+# Again, closely following code from Alen Mitrovski, Xiaoyan Yang, and Matthew Wankiewicz.
+library(haven)
+library(tidyverse)
+
+raw_poststrat_data <- 
+  read_dta(here::here("paper/usa_00009.dta"))
+
+# The Stata format separates labels so reunite those
+raw_poststrat_data <- 
+  labelled::to_factor(raw_poststrat_data)
+head(raw_poststrat_data)
+
+raw_poststrat_data$age <- as.numeric(raw_poststrat_data$age)
+raw_poststrat_data$stateicp <- as.character(raw_poststrat_data$stateicp)
+
+poststrat_data <- 
+  raw_poststrat_data %>% 
+  # filter(inctot < 9999999) %>% 
+  filter(age >= 18) %>% 
+#   mutate(gender = sex) %>% 
+  mutate(
+    age_group = case_when( # case_when works in order and exits when there's a match
+      age <= 29 ~ 'age_18-29',
+      age <= 44 ~ 'age_30-44',
+      age <= 59 ~ 'age_45-59',
+      age >= 60 ~ 'age_60_or_more',
+      TRUE ~ 'Trouble'
+      )
+#     education_level = case_when(
+#       educd == "nursery school, preschool" ~ "High school or less",
+#       educd == "kindergarten" ~ "High school or less",
+#       educd == "grade 1" ~ "High school or less",
+#       educd == "grade 2" ~ "High school or less",
+#       educd == "grade 3" ~ "High school or less",
+#       educd == "grade 4" ~ "High school or less",
+#       educd == "grade 5" ~ "High school or less",
+#       educd == "grade 6" ~ "High school or less",
+#       educd == "grade 7" ~ "High school or less",
+#       educd == "grade 8" ~ "High school or less",
+#       educd == "grade 9" ~ "High school or less",
+#       educd == "grade 10" ~ "High school or less",
+#       educd == "grade 11" ~ "High school or less",
+#       educd == "12th grade, no diploma" ~ "High school or less",
+#       educd == "regular high school diploma" ~ "High school or less",
+#       educd == "ged or alternative credential" ~ "High school or less",
+#       educd == "some college, but less than 1 year" ~ "Some post secondary",
+#       educd == "1 or more years of college credit, no degree" ~ "Some post secondary",
+#       educd == "associate's degree, type not specified" ~ "Post secondary or higher",
+#       educd == "bachelor's degree" ~ "Post secondary or higher",
+#       educd == "master's degree" ~ "Graduate degree",
+#       educd == "professional degree beyond a bachelor's degree" ~ "Graduate degree",
+#       educd == "doctoral degree" ~ "Graduate degree",
+#       educd == "no schooling completed" ~ "High school or less",
+#       TRUE ~ 'Trouble'
+#       )
+    ) %>% 
+  filter(!stateicp %in% c("puerto rico", "state groupings (1980 urban/rural sample)", "military/mil. reservations", "state not identified"))
+
+# Just keep relevant variables
+poststrat_data <- 
+  poststrat_data %>% 
+  select(sex,
+         age_group,
+         stateicp)
+
+# Tidy the class
+poststrat_data <-
+  poststrat_data %>%
+  mutate(across(c(sex, stateicp, age_group), as_factor))
+
+# 
+# # Save data
+# write_csv(poststrat_data, "outputs/data/us_poststrat.csv")
+# 
+# poststrat_data %>% 
+#   head()
+
+```
+
+Now that we have the data, we need to create a `SurveyData` object for the survey dataset.
+
+```{r survey-object}
+library(mrpkit)
+# Create a SurveyData object for the survey dataset 
+us_presidential_election <- 
+  SurveyData$new(
+    data = nationscape_data,
+    questions = list(
+      age_group = "Please identify your age",
+      gender = "Please select your gender",
+      vote_biden = "If today is election day, will you vote for Biden?",
+      state = "Which state do you live in?"
+      ),
+    responses = list(
+      age_group = levels(nationscape_data$age_group),
+      gender = levels(nationscape_data$gender),
+      vote_biden = levels(nationscape_data$vote_biden),
+      state = levels(nationscape_data$state)
+      )
+    )
+
+# Display it
+us_presidential_election$print()
+```
+
+And we now need the same for the post-stratification dataset.
+
+```{r survey-object}
+# Create a SurveyData object for the survey dataset 
+us_post_strat <- 
+  SurveyData$new(
+    data = poststrat_data,
+    questions = list(
+      age_group = "Please identify your age",
+      sex = "Please select your sex",
+      stateicp = "Which state do you live in?"
+      ),
+    responses = list(
+      age_group = levels(poststrat_data$age_group),
+      sex = levels(poststrat_data$sex),
+      stateicp = levels(poststrat_data$stateicp)
+      )
+    )
+
+# Display it
+us_post_strat$print()
+```
+
+Now we need to start to bring the datasets together.
+
+```{r questions}
+# Create QuestionMap$object for the question related to age
+q_age <- QuestionMap$new(
+  name = "age",
+  col_names = c("age_group", "age_group"),
+  values_map = list(
+    "age_18-29" = "age_18-29",
+    "age_30-44" = "age_30-44",
+    "age_45-59" = "age_45-59",
+    "age_60_or_more" = "age_60_or_more"
+  )
+)
+# 
+# # Create QuestionMap$object for the question related to party preference
+# q_party_pref <- QuestionMap$new(
+#   name = "party_pref",
+#   col_names = c("vote_for", "vote_pref"),
+#   values_map = list(
+#     "Box Party" = "BP",
+#     "BP" = "BP",
+#     "Circle Party" = "CP",
+#     "CP" = "CP"
+#   )
+# )
+
+# Create QuestionMap$object for the question related to gender
+q_gender <- QuestionMap$new(
+  name = "gender",
+  col_names = c("gender", "sex"),
+  values_map = data.frame(
+    "Female" = "female",
+    "Male" = "male"
+  )
+)
+
+# Create QuestionMap$object for the question related to state
+q_state <- QuestionMap$new(
+  name = "state",
+  col_names = c("state", "stateicp"),
+  values_map = list(
+  "AL" = "alabama",
+  "AK" = "alaska",
+  "AZ" = "arizona",
+  "AR" = "arkansas",
+  "CA" = "california",
+  "CO" = "colorado",
+  "CT" = "connecticut",
+  "DE" = "delaware",
+  "DC" = "district of columbia",
+  "FL" = "florida",
+  "GA" = "georgia",
+  "HI" = "hawaii",
+  "ID" = "idaho",
+  "IL" = "illinois",
+  "IN" = "indiana",
+  "IA" = "iowa",
+  "KS" = "kansas",
+  "KY" = "kentucky",
+  "LA" = "louisiana",
+  "ME" = "maine",
+  "MD" = "maryland",
+  "MA" = "massachusetts",
+  "MI" = "michigan",
+  "MN" = "minnesota",
+  "MS" = "mississippi",
+  "MO" = "missouri",
+  "MT" = "montana",
+  "NE" = "nebraska",
+  "NV" = "nevada",
+  "NH" = "new hampshire",
+  "NJ" = "new jersey",
+  "NM" = "new mexico",
+  "NY" = "new york",
+  "NC" = "north carolina",
+  "ND" = "north dakota",
+  "OH" = "ohio",
+  "OK" = "oklahoma",
+  "OR" = "oregon",
+  "PA" = "pennsylvania",
+  "RI" = "rhode island",
+  "SC" = "south carolina",
+  "SD" = "south dakota",
+  "TN" = "tennessee",
+  "TX" = "texas",
+  "UT" = "utah",
+  "VT" = "vermont",
+  "VA" = "virginia",
+  "WA" = "washington",
+  "WV" = "west virginia",
+  "WI" = "wisconsin",
+  "WY" = "wyoming"
+  )
+)
+```
+
+```{r}
+q_age
+q_gender
+q_state
+```
+
+We can now bring them together.
+
+```{r map-object}
+# Create a new SurveyMap object
+us_map <- SurveyMap$new(sample = us_presidential_election, 
+                        population = us_post_strat,
+                        q_state, q_age, q_gender
+                        )
+
+print(us_map)
+```
+
+We now finish the process of bringing the datasets together.
+
+```{r mapping}
+us_map$mapping()
+```
+
+If a user wants to include all the variables in the post-stratification step, then a user uses the method `SurveyMap$tabulate` without any arguments.
+This prepares the dataset for post-stratification.
+It is also possible to specify only certain variables in the post-stratification matrix, by including those variables as arguments.
+
+```{r tabulate}
+us_map$tabulate()
+```
+
+We can now model the dataset.
+
+```{r  fitting}
+# Example of using rstanarm::stan_glmer
+fit1 <- us_map$fit(
+  fun = rstanarm::stan_glmer,
+  formula = vote_biden ~ (1 | age) + (1 | gender) + (1 | state),
+  family = "binomial",
+  refresh = 100,
+  cores = 2)
+```
+
+```{r predict}
+# predict the probability of voting for Biden using the fit1 model
+poststrat_est_fit1 <- fit1$population_predict()
+```
+
+```{r}
+# aggregate the predicted value by state
+state_estimation <- fit1$aggregate(poststrat_est_fit1, by = "state")
+
+```
+
+```{r}
+
+# generate the population estimate
+popn_estimation <- fit1$aggregate(poststrat_est_fit1)
+```
+
+```{r}
+library(ggplot2)
+plot_state <- fit1$plot(state_estimation, additional_stats = "none") +
+  ggtitle("Probability of voting the BP by state") +
+  theme_bw()
+plot_state
+```
 

### 2021-12-13 paper/paper.Rmd
Continue updating

--- 2021-12-06 paper/paper.Rmd
@@
 This implementation provides a detailed workflow when conducting MRP that will be useful for researchers and in teaching.
-Unlike writing all the code yourself, using `mrpkit` proactively addresses many common issues, and makes it possible for folks who are new to MRP to quickly conduct their first analysis.
+Using `mrpkit` proactively addresses many common issues and makes it possible to conduct MRP in an efficient way.
 
-    **Keywords:** multilevel regression with post-stratification, R, reproducibility, statistics.
+    **Keywords:** multilevel regression with post-stratification; R; reproducibility; workflow; statistics.
 output:
@@
 The R [@citer] package `mrpkit` allows users to conduct multilevel regression with post-stratification in an opinionated and specific way.
-The package enables users to specify two datasets (a sample dataset and a post-stratification dataset), establish a correspondence between them, and then conduct analysis.
+The package provides a workflow whereby users specify two datasets (a sample dataset and a post-stratification dataset), establish a correspondence between them, and then conduct analysis.
 
@@
 The `DeclareDesign` package [@citeDeclareDesign] is a newer package that is based around a grammar of research design aspects including: data generation, creating treatment and control groups, and sampling.
-In general, these packages are focused on the designing, implementing, and simulating from, surveys.
+In general, these packages are focused on designing, implementing, and simulating from, surveys.
 Ours is focused on what is needed for MRP, and we draw on their packages where possible.
 
-In terms of packages to specifically help with MRP, the most common alternative at the moment to this package is for users to do all aspects of the MRP workflow themselves.
-While there is nothing inherently difficult about MRP, the implementation can be difficult.
+In terms of packages to specifically help with MRP, the most common alternative to this package at the moment is for users to do all aspects of the MRP workflow themselves.
+This can be difficult.
 In particular, preparing and matching different levels between surveys can be time consuming and potentially introduce undocumented errors.
-More recently `tidymrp` [@jospackage] is being developed.
-This package focuses on conducting analysis in a tidy way [@citetidyverse].
-And `ccesMRPprep` [@shirospackage] is focused on using the Cooperative Congressional Election Study (CCES).
+The package `ccesMRPprep` [@shirospackage] is focused on MRP using the Cooperative Congressional Election Study (CCES).
+More recently the package `tidymrp` [@jospackage] is being developed, which focuses on conducting analysis in a tidy way [@citetidyverse].
 
 The main alternative approach is for MRP to be conducted on a case-specific basis.
-This means that a researcher who is interested in MRP estimates writes the code needed to obtain, clean, prepare, analyze and ultimately interpret the estimates.
+This means that a researcher who is interested in MRP estimates writes the code needed to obtain, clean, prepare, analyze, and ultimately interpret the estimates.
 Again, while there is nothing inherently wrong with this approach, successfully conducting MRP requires dealing with many small issues.
@@
 
-The MRP workflow implemented by `mrpkit` is summarised in Figure \@ref(fig:illustrate).
+The MRP workflow implemented by `mrpkit` is summarized in Figure \@ref(fig:illustrate).
 In this section, these steps are expanded on.
@@
 
-We see a variety of use cases including:
+We see a variety of use cases for `mrpkit` including:
 
@@
 
+The main contribution of `mrpkit` across these situations is making steps explicit that are often implicit.
+
 ## Set-up, load package, and read in the datasets
 
-At the moment the `mrpkit` package has not been submitted to CRAN as it is still be finalised. For now it can be installed from GitHub.
+At the moment the `mrpkit` package has not been submitted to CRAN as it is being finalized. For now it can be installed from GitHub.
+
+```{r, eval=FALSE}
+devtools::install_github("lauken13/mrpkit")
+```
 
 ```{r}
-# devtools::install_github("lauken13/mrpkit")
 library(mrpkit)
@@
 A user will need two datasets.
-At this point, one aspect to be aware of is the class of the variables.
-The `mrpkit` package will convert all variables to factor variables.
-Here for the survey and post-stratification datasets we will use simulated datasets that are available as part of the package `mrpkit::shape_survey` and `mrpkit::approx_voters_popn`.
+At this point, a user should be keenly aware of the class of the variables as this becomes important later in the workflow.
+This is because the `mrpkit` package will convert all variables to factor variables.
+
+In this example we will use simulated datasets that are available as part of the package for the survey dataset, `mrpkit::shape_survey`, and the post-stratification dataset, `mrpkit::approx_voters_popn`.
 
@@
 
-Typically MRP requires at least two datasets---usually a survey dataset and a post-stratification dataset.
-A user specifies these using the `SurveyData` method.
+Typically, MRP requires at least two datasets---usually a survey dataset and a post-stratification dataset.
+A user specifies these using the `SurveyData` class.
 `SurveyData` objects are used to represent both the survey and the post-stratification datasets, along with their metadata.
+The `SurveyData$new` method transforms a regular dataframe into a `SurveyData` object.
 A user needs to create `SurveyData` objects for both the survey and post-stratification datasets separately using the `SurveyData$new` method and assigning a name to each.
-The `SurveyData$new` method is used to transform a regular data frame into a `SurveyData` object.
 
-As part of creating these objects a user will typically specify the dataset (from the earlier step), the survey questions, the possible responses, and any weights and design formula that are applicable.
+As part of creating these objects a user will typically specify the dataset (brought into R in the earlier step), the survey questions, the possible responses, and any weights and design formula that are applicable.
 A user does this separately for the survey and the post-stratification dataset, so there will be two objects created.
@@
 
-Building on the datasets earlier, we will show the steps to create `SurveyData` objects using the `SurveyData$new` method. This method takes the following arguments:
+Building on the datasets that were loaded earlier, we will show the steps to create `SurveyData` objects using the `SurveyData$new` method. This method takes the following arguments:
 
 1.
-`data`: The name of the dataframe, i.e., the survey or post-stratification data, that is to be transformed to a `SurveyData` object.
-By default, this method uses each of the factor, character, and binary variables to create questions and responses, unless a user specifies the list of questions and responses.
+`data`: The name of the dataset, i.e., the survey or post-stratification data that will be transformed to a `SurveyData` object.
+By default, this method uses all of the factor, character, and binary variables in the dataset to create questions and responses, unless a user specifies the list of questions and responses that are of interest.
 2.
-`questions`: The actual questions asked in the survey form or questionnaire for each of the columns in the dataframe.
-For instance, a column called 'age' may contain the results of a question 'Please identify your age-group'.
+`questions`: The actual questions asked in the survey form or questionnaire for each of the columns in the dataset For instance, a column called 'age' may contain the results of a question 'Please identify your age-group'.
 3. `responses`: All possible responses for each of the questions. For instance, possible responses in the 'age' column could be: '18-25', '26-35', '36-45', '46-55', '56-65', '66-75', and '76-90'.
-4. (Optional) `weights`: If the weight is specified as a column in the datasets, then this argument should be specified as a string.
+4. (Optional) `weights`: If the weight is specified as a column in the dataset, then this argument should be specified as a string.
 5. (Optional) `design`: If there is a design formula for the survey, then this can be specified using the notation of the `survey` package [@lumleyone; @lumleytwo; @lumleythree].
@@
 Now we create a `SurveyData` object for the post-stratification dataset.
-This code is used to create a `SurveyData` object for post-stratification data when the weight is already specified in the dataset.
-Usually, the post-stratification data is from a large survey, for example, the American Community Survey (ACS) or Demographic and Health Survey (DHS).
+In this case, the weight is already specified in the dataset.
+Often in MRP, the post-stratification dataset is from a large survey, for example, the American Community Survey (ACS) or Demographic and Health Survey (DHS).
 It is possible to have an already summarized post-stratification dataset.
@@
 
-A user can access the properties of the `SurveyData` object using methods for this.
-For instance, the weights, `SurveyData$weights()`, the design, `SurveyData$design()`, the questions, `SurveyData$questions()`, and the number of questions, `SurveyData$n_questions()`, or the responses, `SurveyData$responses()`.
+A user can access the properties of the `SurveyData` object using various methods.
+For instance, the weights, `SurveyData$weights()`, the design, `SurveyData$design()`, the questions, `SurveyData$questions()`, the number of questions, `SurveyData$n_questions()`, and the responses, `SurveyData$responses()`.
 
@@
 This needs to be done for every question that will be used and typically requires specifying a name that will be used for the underlying construct, the column names in each of the two `SurveyData` objects, and a mapping between the values in each.
-This is a deliberately specific task requiring the user to think about and manually link the answers.
+This is a deliberately specific task requiring the user to think about and manually link the values.
 
@@
 Here we define how age-groups, political preferences, gender, education, and states should come together.
-While this is tedious it is worth noting that in general much of this code is likely to be re-useable across different situations.
+While this is tedious it is worth noting that in general much of this code is likely to be re-useable across different projects.
 
@@
 q_age
-q_party_pref
-q_gender
-q_educ
-q_state
 ```
@@
 
-The relationship between the two datasets is finalised by bringing all of this together to create a map between them using `SurveyMap$new` and assigning this to an object.
+The relationship between the two datasets is finalized by bringing all of this together to create a map between them using `SurveyMap$new` and assigning this to an object.
 This typically requires specifying:
@@
 1) the name of the `SurveyData` object related to the sample, in this case 'box_pref'; 
-2) the name of the `SurveyData` object related to the post-stratification dataset, in this case `poststrat_obj`; and 
+2) the name of the `SurveyData` object related to the post-stratification dataset, in this case 'poststrat_obj'; and 
 3) the question mappings that should be included, which specifies the matched labels and values provided by `QuestionMap` objects.
 
-<!-- Probs get rid of: -->
-<!-- The `SurveyMap` object has `new`, `add`, `delete`, `replace`, `mapping`, `tabulate`, and `fit` methods. -->
-
-The method initializes a new `SurveyMap` object is `SurveyMap$new`. This takes `SurveyData` and `QuestionMap` objects as argument.
+To initialize a new `SurveyMap` object a user can use the `SurveyMap$new` method. This takes `SurveyData` and `QuestionMap` objects as argument.
 
@@
 
-The user has mapped all the questions in the post-stratification object.
-Variables could be excluded using the `SurveyMap$delete` method.
-Similarly, `SurveyMap$replace` enables a user to use another level of matching.
+At this point the user has mapped the questions of interest between the survey and the post-stratification datasets.
+Variables could be excluded using the `SurveyMap$delete` method, or replaced using the `SurveyMap$replace` method.
 
-## Finalising the mapping and tabulation
+## Finalizing the mapping and tabulation
 
-At this point, the datasets have been brought together.
-This process is finalised with `SurveyMap$mapping`.
+The process of bringing the datasets together is finalized with the method `SurveyMap$mapping`.
 This creates new sample and post-stratification datasets with unified variable names, and levels.
@@
 
-If a user wants to include all the variables in the post-stratification step, then a user uses the method `SurveyMap$tabulate` without any arguments.
-This prepares the dataset for post-stratification.
-It is also possible to specify only certain variables in the post-stratification matrix, by including those variables as arguments.
+If a user wants to include all the variables in the post-stratification step, then a user can use the `SurveyMap$tabulate` method without any arguments.
+It is also possible to specify only certain variables in the post-stratification step, by including those variables as arguments.
 
@@
 The user can then fit a model to the survey dataset using `SurveyMap$fit` and naming it to create an object.
-Built-in options include: `rstanarm::stan_glmer()` and `rstanarm::stan_glm()` [@citestanarm], `lme4::glmer()` [@citelme4] , `brms::brm()` [@citebrms].
+Built-in options include: `rstanarm::stan_glmer()` and `rstanarm::stan_glm()` [@citestanarm], `lme4::glmer()` [@citelme4], and `brms::brm()` [@citebrms].
 A user could also specify their own function.
-For that purpose there is a `data` argument that accepts a dataframe.
+For that purpose, there is a `data` argument that accepts a dataframe.
 In this example, we fit three models using `rstanarm::stan_glmer()`, `lme4::glmer()`, and `brms:brm()` with `age`, `gender`, and `education` as the predictors.
@@
 2) Considering these post-stratified estimates at a more-aggregated level, for instance state or province level, using `SurveyFit$aggregate`.
-3) Make a table that summarises the posterior mean and standard deviation, along with comparisons with raw estimates and weighted estimates, using the `summary` method.
-4) Visualise the estimates using the `plot` method.
+3) Make a table that summarizes the posterior mean and standard deviation, along with comparisons with raw estimates and weighted estimates, using the `summary` method.
+4) Visualize the estimates using the `plot` method.
 
@@
 The next step is to generate the population estimate or group estimate using the `SurveyFit$aggregate` method.
-This method takes two arguments, namely the post-stratification estimate data and the variable whose level to which the estimated value would be aggregated to.
+This method takes two arguments, namely the post-stratification estimate and the variable whose level to which the estimated value would be aggregated to.
 In this example, we want to aggregate the estimated value by the level of `age` and `state`.
@@
 
-This method then will return a data frame with the variable levels and probability of the outcome, `y`.
+This method then will return a dataframe with the variable levels and probability of the outcome, `y`.
 In this example, `age_estimation` represents the age group and the probability of voting for the Box Party for each age group.
@@
 
-### Summarise and compare
+### Summarize and compare
 
@@
 We can use the `summary` method do to this.
-This method takes the aggregated data frame and produces a dataframe of mean and standard deviation estimates for each group (if in the aggregated frame) or overall in the population.
+This method takes the aggregated dataframe and produces a dataframe of mean and standard deviation estimates for each group (if in the aggregated frame) or overall in the population.
 The standard deviation for the weights is created by wrapping the survey package using the original design specified, and the raw standard deviation is $\sqrt\frac{pq}{n}$.
@@
 Once we have the aggregated estimates, we can visualize them using the `plot` method.
-This method takes the aggregated data frame and additional_stats option (the default is wtd and raw estimates) as its arguments.
+This method takes the aggregated dataframe and additional_stats option (the default is wtd and raw estimates) as its arguments.
 It generates a violin plot of the aggregated estimate by the level of a certain variable.
-When additional statistics are specified (options are raw, wtd and mrp; more than one can be specified as a vector), these are added as points with error bars representing the 95% confidence interval.
-In this example, we display the estimated plot for `age` and `state` and a density plot for the population estimates.
+When additional statistics are specified (options are raw, wtd and mrp; more than one can be specified as a vector), these are added as points with error bars representing the 95 per cent confidence interval.
+In this example, we display the estimated plot for 'age' and 'state' and a density plot for the population estimates.
 
@@
 
+There are a variety of technical notes related to this package and the architectural decisions that were made.
+
+Section \@ref(implementation) discusses some of the implementation issues and technical notes related to the decisions that were made
+
 # Worked example {#example}
+
+In this example we will forecast the 2020 US presidential election using survey data from the Democracy Fund Voter Study Group and use the Integrated Public Use Microdata Series (IPUMS), to access the 2018 American Community Survey (ACS) to use as post-stratification data.
+We are interested in considering the following variables: state, age-group, and gender.
+
+The first step is to gather the datasets.
+For the survey data we are interested in the Nationscape dataset, which can be requested from the Voter Study Group.
+Access is usually granted within hours and is completely free.
+We are going to use the survey from 25 June 2020.
+
+```{r, warning=FALSE, message=FALSE}
+# This code based on that from Alen Mitrovski, Xiaoyan Yang, and Matthew Wankiewicz.
+library(haven)
+library(tidyverse)
+
+raw_nationscape_data <- 
+  read_dta(here::here("paper/ns20200625.dta"))
+
+# The Stata format separates labels so reunite those
+raw_nationscape_data <- 
+  labelled::to_factor(raw_nationscape_data)
+
+# Just keep relevant variables
+nationscape_data <- 
+  raw_nationscape_data %>% 
+  select(vote_2020,
+         gender,
+         state,
+         age)
+
+# For simplicity, remove anyone undecided or planning to vote for someone other than Biden/Trump and make vote a binary variable: 1 for Biden, 0 for Trump.
+nationscape_data <- 
+  nationscape_data %>% 
+  filter(vote_2020 == "Joe Biden" | vote_2020 == "Donald Trump") %>% 
+  mutate(vote_biden = if_else(vote_2020 == "Joe Biden", 1, 0)) %>% 
+  select(-vote_2020)
+
+# Create the dependent variables by grouping the existing variables
+nationscape_data <-
+  nationscape_data %>%
+  mutate(
+    age_group = case_when(
+      age <= 29 ~ 'age_18-29',
+      age <= 44 ~ 'age_30-44',
+      age <= 59 ~ 'age_45-59',
+      age >= 60 ~ 'age_60_or_more',
+      TRUE ~ 'Trouble'
+      )
+    )
+
+# Tidy the class
+nationscape_data <-
+  nationscape_data %>%
+  mutate(across(c(gender, state, age_group, vote_biden), as_factor))
+
+nationscape_data %>% 
+  head()
+```
+
+We also need a post-stratification dataset. In this case we can use the American Community Survey (ACS) accessed through IPUMS.
+
+```{r, warning=FALSE, message=FALSE}
+# Again, closely following code from Alen Mitrovski, Xiaoyan Yang, and Matthew Wankiewicz.
+library(haven)
+library(tidyverse)
+
+raw_poststrat_data <-
+  read_dta(here::here("paper/usa_00009.dta"))
+
+# The Stata format separates labels so reunite those
+raw_poststrat_data <-
+  labelled::to_factor(raw_poststrat_data)
+head(raw_poststrat_data)
+
+raw_poststrat_data$age <- as.numeric(raw_poststrat_data$age)
+raw_poststrat_data$stateicp <-
+  as.character(raw_poststrat_data$stateicp)
+
+poststrat_data <-
+  raw_poststrat_data %>%
+  filter(age >= 18) %>%
+  mutate(
+    age_group = case_when(
+      age <= 29 ~ 'age_18-29',
+      age <= 44 ~ 'age_30-44',
+      age <= 59 ~ 'age_45-59',
+      age >= 60 ~ 'age_60_or_more',
+      TRUE ~ 'Trouble'
+    )
+  ) %>%
+  filter(
+    !stateicp %in% c(
+      "puerto rico",
+      "state groupings (1980 urban/rural sample)",
+      "military/mil. reservations",
+      "state not identified"
+    )
+  )
+
+# Just keep relevant variables
+poststrat_data <-
+  poststrat_data %>%
+  select(sex,
+         age_group,
+         stateicp)
+
+# Tidy the class
+poststrat_data <-
+  poststrat_data %>%
+  mutate(across(c(sex, stateicp, age_group), as_factor))
+```
+
+Now that we have the data, we need to create a `SurveyData` object for the survey dataset.
+
+```{r survey-object-us}
+library(mrpkit)
+# Create a SurveyData object for the survey dataset 
+us_presidential_election <- 
+  SurveyData$new(
+    data = nationscape_data,
+    questions = list(
+      age_group = "Please identify your age",
+      gender = "Please select your gender",
+      vote_biden = "If today is election day, will you vote for Biden?",
+      state = "Which state do you live in?"
+      ),
+    responses = list(
+      age_group = levels(nationscape_data$age_group),
+      gender = levels(nationscape_data$gender),
+      vote_biden = levels(nationscape_data$vote_biden),
+      state = levels(nationscape_data$state)
+      )
+    )
+
+# Display it
+us_presidential_election$print()
+```
+
+And we now need the same for the post-stratification dataset.
+
+```{r survey-object-post-strat}
+# Create a SurveyData object for the survey dataset 
+us_post_strat <- 
+  SurveyData$new(
+    data = poststrat_data,
+    questions = list(
+      age_group = "Please identify your age",
+      sex = "Please select your sex",
+      stateicp = "Which state do you live in?"
+      ),
+    responses = list(
+      age_group = levels(poststrat_data$age_group),
+      sex = levels(poststrat_data$sex),
+      stateicp = levels(poststrat_data$stateicp)
+      )
+    )
+
+# Display it
+us_post_strat$print()
+```
+
+Now we need to start to bring the datasets together.
+
+```{r questions-us}
+# Create QuestionMap$object for the question related to age
+q_age <- QuestionMap$new(
+  name = "age",
+  col_names = c("age_group", "age_group"),
+  values_map = list(
+    "age_18-29" = "age_18-29",
+    "age_30-44" = "age_30-44",
+    "age_45-59" = "age_45-59",
+    "age_60_or_more" = "age_60_or_more"
+  )
+)
+
+# Create QuestionMap$object for the question related to gender
+q_gender <- QuestionMap$new(
+  name = "gender",
+  col_names = c("gender", "sex"),
+  values_map = data.frame(
+    "Female" = "female",
+    "Male" = "male"
+  )
+)
+
+# Create QuestionMap$object for the question related to state
+q_state <- QuestionMap$new(
+  name = "state",
+  col_names = c("state", "stateicp"),
+  values_map = list(
+  "AL" = "alabama",
+  "AK" = "alaska",
+  "AZ" = "arizona",
+  "AR" = "arkansas",
+  "CA" = "california",
+  "CO" = "colorado",
+  "CT" = "connecticut",
+  "DE" = "delaware",
+  "DC" = "district of columbia",
+  "FL" = "florida",
+  "GA" = "georgia",
+  "HI" = "hawaii",
+  "ID" = "idaho",
+  "IL" = "illinois",
+  "IN" = "indiana",
+  "IA" = "iowa",
+  "KS" = "kansas",
+  "KY" = "kentucky",
+  "LA" = "louisiana",
+  "ME" = "maine",
+  "MD" = "maryland",
+  "MA" = "massachusetts",
+  "MI" = "michigan",
+  "MN" = "minnesota",
+  "MS" = "mississippi",
+  "MO" = "missouri",
+  "MT" = "montana",
+  "NE" = "nebraska",
+  "NV" = "nevada",
+  "NH" = "new hampshire",
+  "NJ" = "new jersey",
+  "NM" = "new mexico",
+  "NY" = "new york",
+  "NC" = "north carolina",
+  "ND" = "north dakota",
+  "OH" = "ohio",
+  "OK" = "oklahoma",
+  "OR" = "oregon",
+  "PA" = "pennsylvania",
+  "RI" = "rhode island",
+  "SC" = "south carolina",
+  "SD" = "south dakota",
+  "TN" = "tennessee",
+  "TX" = "texas",
+  "UT" = "utah",
+  "VT" = "vermont",
+  "VA" = "virginia",
+  "WA" = "washington",
+  "WV" = "west virginia",
+  "WI" = "wisconsin",
+  "WY" = "wyoming"
+  )
+)
+```
+
+We can now bring them together.
+
+```{r map-object-us}
+# Create a new SurveyMap object
+us_map <- SurveyMap$new(sample = us_presidential_election, 
+                        population = us_post_strat,
+                        q_state, q_age, q_gender
+                        )
+
+print(us_map)
+```
+
+We now finish the process of bringing the datasets together.
+
+```{r mapping-us}
+us_map$mapping()
+```
+
+If a user wants to include all the variables in the post-stratification step, then a user uses the method `SurveyMap$tabulate` without any arguments.
+This prepares the dataset for post-stratification.
+It is also possible to specify only certain variables in the post-stratification matrix, by including those variables as arguments.
+
+```{r tabulate-us}
+us_map$tabulate()
+```
+
+We can now model the dataset.
+
+```{r  fitting-us}
+# Example of using rstanarm::stan_glmer
+fit1 <- us_map$fit(
+  fun = rstanarm::stan_glmer,
+  formula = vote_biden ~ (1 | age) + (1 | gender) + (1 | state),
+  family = "binomial",
+  refresh = 100,
+  cores = 2)
+```
+
+```{r predict-us}
+# predict the probability of voting for Biden using the fit1 model
+poststrat_est_fit1 <- fit1$population_predict()
+```
+
+```{r}
+# aggregate the predicted value by state
+state_estimation <- fit1$aggregate(poststrat_est_fit1, by = "state")
+
+```
+
+```{r}
+# generate the population estimate
+popn_estimation <- fit1$aggregate(poststrat_est_fit1)
+```
+
+```{r}
+library(ggplot2)
+plot_state <- fit1$plot(state_estimation, additional_stats = "none") +
+  ggtitle("Probability of supporting Biden by state") +
+  theme_classic()
+plot_state
+```
 
@@
 
+One of the defining features of `mrpkit` is the use of R6 objects. This
+
 ## Next steps and cautions

### 2022-01-24 paper/paper.Rmd
Something is better than nothing

@@
 The package helps appropriately set-up the survey and post-stratification datasets and then establish a relationship between different variables in these datasets.
-A substantial amount of the work to do this is specified by the methods in the package, saving time, enhancing reproducibility, and reducing the risk of coding errors.
+A substantial amount of the work to do this is specified by the package's methods, saving time, enhancing reproducibility, and reducing the risk of coding errors.
 The package natively supports multilevel binomial and Bernoulli models fit with `lme4` and `Stan` (via `brms` and `rstanarm`) and allows for the use of custom modeling functions.
-Summaries and visualizations of the resulting post-stratified estimates can be conducted using the methods in the package.
+Summaries and visualizations of the resulting post-stratified estimates can be conducted using the package's methods.
 This implementation provides a detailed workflow when conducting MRP that will be useful for researchers and in teaching.
@@
 
-The process of conducting MRP involves considering some survey outcome as being due to a combination of many different cells that are typically demographic or geographic in nature, estimating this relationship using a multilevel regression model and then aggregating cell-level estimates to match the population [@wang2015forecasting].
+```{r laurenflowchart, echo = FALSE, fig.cap = "Example of typical MRP workflow", out.width = '90%'}
+knitr::include_graphics(here::here("paper/mrpkit_workflow.pdf"))
+```
+
+The process of conducting MRP involves considering some survey outcome as being due to a combination of many different cells that are typically demographic or geographic in nature, estimating this relationship using a multilevel regression model and then aggregating cell-level estimates to match the population Figure \@ref(fig:laurenflowchart).
 The survey of interest would typically be a regular survey, such as a political poll of, say, 1,000 respondents, but it could also be a larger survey, such as the American National Election Studies, or similar.
@@
 The `srvyr` package [@citesrvyr] is a variant of `survey` that adds a syntax similar to `dplyr` [@citedplyr].
-The `DeclareDesign` package [@citeDeclareDesign] is a newer package that is based around a grammar of research design aspects including: data generation, creating treatment and control groups, and sampling.
+The `DeclareDesign` package [@citeDeclareDesign] is a newer package that is based around a grammar of research design.
+These include: data generation, creating treatment and control groups, and sampling.
 In general, these packages are focused on designing, implementing, and simulating from, surveys.
@@
 
-The main alternative approach is for MRP to be conducted on a case-specific basis.
+The main alternative to using `mrpkit` is for MRP to be conducted on a case-specific basis.
 This means that a researcher who is interested in MRP estimates writes the code needed to obtain, clean, prepare, analyze, and ultimately interpret the estimates.
-Again, while there is nothing inherently wrong with this approach, successfully conducting MRP requires dealing with many small issues.
-Each of these is small but getting them wrong can have large effects that are potentially unnoticed on the estimates.
+Again, while there is nothing inherently wrong with this approach, successfully conducting MRP requires dealing with a variety of issues.
+Each of these is small, but getting them wrong can have large effects on the estimates that are potentially unnoticed.
 It is also easy to overlook steps, or to not appropriately document them.
-`mrpkit` is prescriptive about the steps that must be taken, and ensures that all aspects are documented.
+`mrpkit` is prescriptive about the steps that must be taken when conducting MRP, and ensures that all aspects are documented.
 
@@
 - A researcher with domain or survey expertise, but who does not have the resources to ensure that the code has been appropriately tested.
-- A researcher who is experienced with MRP needs to run many different models on the same dataset.
+- A researcher who is experienced with MRP needs to run many different models on the same, or similar, dataset/s.
 
@@
 This means ensuring that it is in a rectangular format such as a CSV and then reading it in as a dataframe.
-The `mrpkit` package does not explicitly contain methods for this, instead a user could use `readr::read_csv()` [@citereadr] or `haven::read_dta()` [@citehaven].
+The `mrpkit` package does not explicitly contain methods for this.
+Instead a user could use `readr::read_csv()` [@citereadr] or `haven::read_dta()` [@citehaven].
 
 A user will need two datasets.
-At this point, a user should be keenly aware of the class of the variables as this becomes important later in the workflow.
-This is because the `mrpkit` package will convert all variables to factor variables.
+At this point, a user should take note of the class of the variables as this becomes important later in the workflow.
+This is because the `mrpkit` package may convert some variables to factors.
 
@@
 2.
-`questions`: The actual questions asked in the survey form or questionnaire for each of the columns in the dataset For instance, a column called 'age' may contain the results of a question 'Please identify your age-group'.
+`questions`: The actual questions asked in the survey form or questionnaire for each of the columns in the dataset.
+For instance, a column called 'age' may contain the results of a question 'Please identify your age-group'.
 3. `responses`: All possible responses for each of the questions. For instance, possible responses in the 'age' column could be: '18-25', '26-35', '36-45', '46-55', '56-65', '66-75', and '76-90'.
@@
 
-To initialize a new `SurveyMap` object a user can use the `SurveyMap$new` method. This takes `SurveyData` and `QuestionMap` objects as argument.
+To initialize a new `SurveyMap` object a user can use the `SurveyMap$new` method. This takes `SurveyData` and `QuestionMap` objects as arguments.
 
@@
 # Create a new SurveyMap object
-ex_map <- SurveyMap$new(sample = box_pref, population = poststrat_obj,
-                        q_age, q_educ, q_gender, q_party_pref
-                        )
+ex_map <- SurveyMap$new(sample = box_pref,
+                        population = poststrat_obj,
+                        q_age,
+                        q_educ,
+                        q_gender,
+                        q_party_pref)
 
@@
 The next step is to generate the population estimate or group estimate using the `SurveyFit$aggregate` method.
-This method takes two arguments, namely the post-stratification estimate and the variable whose level to which the estimated value would be aggregated to.
+This method takes two arguments, namely the post-stratification estimates and the variable whose level to which the estimated value would be aggregated to.
 In this example, we want to aggregate the estimated value by the level of `age` and `state`.
@@
 ```
-
-# Implementation and technical notes {#implementation}
-
-There are a variety of technical notes related to this package and the architectural decisions that were made.
-
-Section \@ref(implementation) discusses some of the implementation issues and technical notes related to the decisions that were made
 

### 2022-01-31 paper/paper.Rmd
Updated motivation in Intro

@@
 The package natively supports multilevel binomial and Bernoulli models fit with `lme4` and `Stan` (via `brms` and `rstanarm`) and allows for the use of custom modeling functions.
-Summaries and visualizations of the resulting post-stratified estimates can be conducted using the package's methods.
-This implementation provides a detailed workflow when conducting MRP that will be useful for researchers and in teaching.
+Summaries and visualizations of the resulting post-stratified estimates can be created using the package's methods.
+This implementation provides a detailed workflow when conducting MRP that will be useful for students, educators and researchers.
 Using `mrpkit` proactively addresses many common issues and makes it possible to conduct MRP in an efficient way.
@@
 
-Multilevel regression with post-stratification (MRP) is a statistical approach in which surveys are related to each other using a statistical model.
+Multilevel regression and post-stratification (MRP [@Gelman97poststratificationinto, @park2004bayesian, @lax2009should]) is a statistical approach used to make population estimates using survey data.
+The method is often characterised in two steps.
+The first uses a multilevel regression to estimate the relationship between some binary outcome variable (that we would like a population estimate for) and demographic groups.
+The second uses this model to predict the probability of the outcome for each combination of demographic variable.
+A weighted mean of probabilities using the number of individuals in each demographic combination of weights is taken to create a population estimate.
+
+In this paper and package, we argue that MRP is not a simple step procedure but rather a many stepped procedure that includes a number of implicit steps and procedures.
+Our hope in proposing the package is that these steps and procedures will become explicit, which will provide the opportunity for both more reproducible and defensible code.
+In doing so, we are proposing an explicit and opinionated workflow for MRP (see Figure \@ref(fig:laurenflowchart)).
+
+To achieve this, we begin by first conceptualising MRP as a relationship between two surveys.
+We argue that one of the key implicit steps of MRP is a mapping demographics measurements between a survey and a population.
+Rarely will these two datasets contain the exact same wording with the exact same demographic categories.
+Even if they do, the two variables will likely have different names in the survey and population datasets.
+It is from this mapping that the power of MRP exists, but it can be difficult to establish this mapping in a reproducible way that fits into a typical applied statistics workflow.
+It can be difficult even for those experienced with MRP to ensure there are no mistakes in this mapping.
+
+<!-- The first is the survey that measures the outcome variable in which surveys are related to each other using a statistical model.
 This is useful because known biases in one survey can be adjusted by another in a statistically reasonable way.
@@
 The package provides a workflow whereby users specify two datasets (a sample dataset and a post-stratification dataset), establish a correspondence between them, and then conduct analysis.
+-->
 
@@
 
-The process of conducting MRP involves considering some survey outcome as being due to a combination of many different cells that are typically demographic or geographic in nature, estimating this relationship using a multilevel regression model and then aggregating cell-level estimates to match the population Figure \@ref(fig:laurenflowchart).
+We also note that in the majority of MRP analyses, the post-stratification matrix is not known but is estimated using another, high quality (large, probability with high response rate) survey such as the American Community Survey (CITE).
+The reason for this approximation is generally due to the scarcity of information collected at the full population.
+In addition, as MRP has been applied in non-probability surveys (CITE XBOX) and to estimate small areas in the population (CITE DEEP INTERACTIONS), the number of desired demographics to adjust by has grown.
+This reinforces the concept that MRP is a mapping between two surveys - one which estimates the population well, and one that includes some variable that we wish to estimate in the population.
+
+<!-- The process of conducting MRP involves considering some survey outcome as being due to a combination of many different cells that are typically demographic or geographic in nature, estimating this relationship using a multilevel regression model and then aggregating cell-level estimates to match the population Figure \@ref(fig:laurenflowchart).
 The survey of interest would typically be a regular survey, such as a political poll of, say, 1,000 respondents, but it could also be a larger survey, such as the American National Election Studies, or similar.
@@
 The model would typically be a Bayesian hierarchical model, but in principle any variety could be used.
+-->
 
-At its core, MRP is a mapping between a survey object and a population object.
-It is from this mapping that the power of MRP exists, but it can be difficult to establish this mapping in a reproducible way that fits into a typical applied statistics workflow.
-Making this implementation easier and more reproducible is important because interest in, and the use of, MRP has considerably increased.
-It can be difficult even for those experienced with MRP to ensure there are no mistakes in this mapping.
-The `mrpkit` package creates a well-defined workflow for MRP that represents an important contribution for enhancing the reproducibility of MRP analysis.
-
-Our package complements existing packages such as `survey` [@lumleyone; @lumleytwo; @lumleythree], and `DeclareDesign` [@citeDeclareDesign].
-The `survey` package is well-established and allows a user to specify a survey design, such as clustered, stratified, etc, generate summary statistics, and conduct sub-population analysis.
-The `srvyr` package [@citesrvyr] is a variant of `survey` that adds a syntax similar to `dplyr` [@citedplyr].
-The `DeclareDesign` package [@citeDeclareDesign] is a newer package that is based around a grammar of research design.
-These include: data generation, creating treatment and control groups, and sampling.
-In general, these packages are focused on designing, implementing, and simulating from, surveys.
-Ours is focused on what is needed for MRP, and we draw on their packages where possible.
+With the proliferation of recent advances to the model component of MRP (such as MRSP [@gao2021improving], SRP [@ornstein2020stacked], BARP [@bisbee2019barp]) it seems likely that multiple models will be fit in a single MRP analysis so that that modelling decisions can be compared by their impact on the population estimate.
+Many existing packages with an MRP focus (e.g., DGO) tend to implement an MRP workflow for their specific modelling method, but it is not easy to compare modelling methods between then without duplication of code.
+The least restrictive of these is (autoMrP; @broniecki2021automrp), which uses a stacked method but still does not allow for novel methods to be included/compared easily.
+Our hope with this package is provide a general scaffolding on which advances in MRP research can be laid upon in future.
 
@@
 This can be difficult.
-In particular, preparing and matching different levels between surveys can be time consuming and potentially introduce undocumented errors.
-The package `ccesMRPprep` [@shirospackage] is focused on MRP using the Cooperative Congressional Election Study (CCES).
-More recently the package `tidymrp` [@jospackage] is being developed, which focuses on conducting analysis in a tidy way [@citetidyverse].
+In particular, preparing and matching different levels between surveys can be time consuming and potentially introduce undocumented decisions if not errors.
+Efforts like the `ccesMRPprep` [@shirospackage] package focus on the full MRP including data cleaning using the Cooperative Congressional Election Study (CCES), but this code is specific to the CCCES rather than a general solution.
+It does however, highlight how important the preproccessing of surveys for MRP really is.
+<!-- More recently the package `tidymrp` [@jospackage] is being developed, which focuses on conducting analysis in a tidy way [@citetidyverse]. -->
+
+Our package complements existing packages such as `survey` [@lumleyone; @lumleytwo; @lumleythree], and `DeclareDesign` [@citeDeclareDesign].
+The `survey` package is well-established and allows a user to create and use weights for a given survey design (e.g., clustered, stratified, etc).
+The `srvyr` package [@citesrvyr] is a variant of `survey` that adds a syntax similar to `dplyr` [@citedplyr].
+Our proposed package differs from the survey package because MRP is an alternate method to weights based analysis, and because of the difference in the two methods it is not possible to add an MRP method.
+The `DeclareDesign` package [@citeDeclareDesign] is a newer package that is based around a grammar of research design.
+These include: data generation, creating treatment and control groups, and sampling.
+While useful, DeclareDesign has a greater focus on simulation and understanding different designs.
 

### 2023-06-29 R/SurveyData.R
minor doc edits

@@ -3,12 +3,13 @@
 #' @name SurveyData
 #' @export
 #' @description
-#' A `SurveyData` object represents a survey and its metadata. The survey itself
-#' is a data frame containing all data from the survey. The survey data object
-#' also includes the survey questions and responses (if left empty these will
-#' just be the column and factor level names). To enable weighted comparisons,
-#' survey weights and a survey design can be specified, with the survey design
-#' specified using \pkg{survey} package notation.
+#' An [R6][R6::R6Class] `SurveyData` object represents a survey and its
+#' metadata. The survey itself is a data frame containing all data from the
+#' survey. The `SurveyData` object also includes the survey questions and
+#' responses (if left empty these will just be the column and factor level
+#' names). To enable weighted comparisons, survey weights and a survey design
+#' can be specified, with the survey design specified using \pkg{survey} package
+#' notation.
 #'
 #' @examples
 #'
@@ -44,7 +45,7 @@
 #'     gender = "Gender?",
 #'     vote_pref = "Which party do you prefer to vote for?"
 #'   ),
-#'   # order doesn't matter (gender before age2 here) because
+#'   # order doesn't matter (gender before age here) because
 #'   # the list has the names of the variables
 #'   responses = list(
 #'     gender = levels(approx_voters_popn$gender),
@@ -95,7 +96,7 @@ SurveyData <- R6::R6Class(
     #'     age_group = "Which age group are you?",
     #'     gender = "Gender?"
     #'   ),
-    #'   # order doesn't matter (gender before age2 here) because
+    #'   # order doesn't matter (gender before age here) because
     #'   # the list has the names of the variables
     #'   responses = list(
     #'     gender = levels(approx_voters_popn$gender),

### 2023-06-29 R/SurveyFit.R

@@ -4,8 +4,8 @@
 #' @export
 #' @description An [R6][R6::R6Class] `SurveyFit` object stores a fitted model
 #'   object and provides methods for generating predicted probabilities for all
-#'   poststrat cells, generating population and group estimates, and visualizing
-#'   results.
+#'   poststratification cells, generating population and group estimates, and
+#'   visualizing results.
 #' @inherit SurveyMap examples
 #'
 SurveyFit <- R6::R6Class(
@@ -121,7 +121,7 @@ SurveyFit <- R6::R6Class(
     },
 
     #' @description Aggregate estimates to the population level or by level of a grouping variable
-    #' @param poststrat_estimates The object returned by `population_predict`.
+    #' @param poststrat_estimates The object returned by the `population_predict` method.
     #' @param by Optionally a string specifying a grouping variable. If
     #'   specified the aggregation will happen by level of the named variable.
     #'   If not specified population-level estimates will be computed.
@@ -157,7 +157,7 @@ SurveyFit <- R6::R6Class(
     },
     #' @description Creates a set of summary statistics of the mrp estimate,
     #' and corresponding weighted and raw data estimates
-    #' @param aggregated_estimates The data frame returned by the `aggregate() method`.
+    #' @param aggregated_estimates The data frame returned by the `aggregate` method.
     #' @return A data frame that consists of a minimum three rows with the raw, MRP
     #' and weighted estimates, plus an estimate of standard error. If the aggregated estimates
     #' were specified with a `by` argument (indicating sub population or small area estimates),
@@ -218,18 +218,19 @@ SurveyFit <- R6::R6Class(
 
     #' @description Plot takes the aggregated MRP estimates and produces a quick
     #' visualization of total and sub-population estimates.
-    #' @param aggregated_estimates The data frame returned by `aggregate`.
+    #' @param aggregated_estimates The data frame returned by the `aggregate` method.
     #' @param additional_stats A vector that specifies which of three additional
-    #'   stats ("wtd", "raw", "mrp", "none") should be included on the plot. The
-    #'   default value is to include the weighted estimate (wtd) and raw data
-    #'   mean (raw), but an analogous bar for MRP (mrp) can be added using the
+    #'   stats ("wtd", "raw", "mrp", "none") should be overlayed on the plot.
+    #'   The default is to include the weighted estimate (wtd) and raw data mean
+    #'   (raw), but an analogous bar for MRP (mrp) can be added using the
[…]

### 2023-06-29 R/SurveyMap.R

@@ -13,9 +13,11 @@
 #' they should be listed in that order, either descending or ascending.
 #'
 #' @examples
-#' library(dplyr)
 #'
+#' # Some fake survey data for demonstration
 #' head(shape_survey)
+#'
+#' # Create SurveyData object for the sample
 #' box_prefs <- SurveyData$new(
 #'   data = shape_survey,
 #'   questions = list(
@@ -39,7 +41,11 @@
 #' box_prefs$print()
 #' box_prefs$n_questions()
 #'
+#'
+#' # Some fake population data for demonstration
 #' head(approx_voters_popn)
+#'
+#' # Create SurveyData object for the population
 #' popn_obj <- SurveyData$new(
 #'   data = approx_voters_popn,
 #'   questions = list(
@@ -47,7 +53,7 @@
 #'     gender = "Gender?",
 #'     vote_pref = "Which party do you prefer to vote for?"
 #'   ),
-#'   # order doesn't matter (gender before age2 here) because
+#'   # order doesn't matter (gender before age here) because
 #'   # the list has the names of the variables
 #'   responses = list(
 #'     gender = levels(approx_voters_popn$gender),
@@ -58,6 +64,9 @@
 #' )
 #' popn_obj$print()
 #'
+#'
+#' # Create the QuestionMap objects mapping each question between the
+#' # survey and population dataset
 #' q_age <- QuestionMap$new(
 #'   name = "age",
 #'   col_names = c("age","age_group"),
@@ -79,6 +88,7 @@
 #'   values_map = list("male" = "m","female" = "f", "nonbinary" = "nb")
 #' )
 #'
+#'
 #' # Create SurveyMap object adding all questions at once
 #' ex_map <- SurveyMap$new(
 #'   sample = box_prefs,
@@ -99,6 +109,7 @@
 #' ex_map$add(q_gender)
 #' print(ex_map)
 #'
+#'
 #' # Create the mapping between sample and population
 #' ex_map$mapping()
 #'
@@ -106,24 +117,27 @@
 #' # (alternatively, can specify particular variables, e.g. tabulate("age"))
 #' ex_map$tabulate()
 #'
-#' # Example rstanarm usage
-#' # Returns a SurveyFit object
+#' # Take a peak at the poststrat data frame
+#' head(ex_map$poststrat_data())
+#'
 #' \dontrun{
+#' # Fit regression model using rstanarm (returns a SurveyFit object)
[…]

### 2023-06-29 README.Rmd
use readme.Rmd to generate readme.md
this allows us to execute code in the readme for demonstration

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mrpkit

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/mrpkit)](https://CRAN.R-project.org/package=mrpkit)
[![R-CMD-check](https://github.com/lauken13/mrpkit/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/lauken13/mrpkit/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/lauken13/mrpkit/branch/master/graph/badge.svg)](https://app.codecov.io/gh/lauken13/mrpkit?branch=master)
<!-- badges: end -->

**NOTE: This package is still a work in progress and is yet not released or officially supported**

For students and researchers who are comfortable with at least `glm()` and want
to conduct multilevel regression with post-stratification (MRP), `mrpkit`
provides a reproducible, opinionated, and highly structured workflow. Unlike
writing all the code yourself, using `mrpkit` proactively addresses many common
issues, and makes it possible for folks who are new to MRP to quickly conduct
their first analysis.

The package assists in setting up the survey data and relationships between
different variables in the sample and the population. From there, a substantial
amount of data cleaning is automated, saving time and reducing the risk of
coding errors. The package has native support for multilevel binomial and
Bernoulli models fit with lme4 and Stan (via brms and rstanarm) and also allows
for the use of custom modeling functions. Summaries and simple visualizations of
the resulting post-stratified estimates are provided.

## Installation

You can install the development version of `mrpkit` from GitHub:

``` r
# install.packages("remotes")
remotes::install_github("lauken13/mrpkit")
```

### License

`mrpkit` is licensed under an MIT license. See the `LICENSE.md` file.

## Example

```{r example}
library(mrpkit)

# Some fake survey data for demonstration
head(shape_survey)

# Create SurveyData object for the sample
box_prefs <- SurveyData$new(
  data = shape_survey,
  questions = list(
    age = "Please identify your age group",
    gender = "Please select your gender",
    vote_for = "Which party did you vote for in the 2018 election?",
    y = "If today is the election day, would you vote for the Box Party?"
  ),
  responses = list(
    age = levels(shape_survey$age),
    gender = levels(shape_survey$gender),
    # Here we use a dataframe for the responses because the levels in the data are abridged versions of the actual responses
    # This can be useful when surveys have brief/non descriptive responses.
    vote_for = data.frame(data = levels(shape_survey$vote_for),
    asked = c("Box Party Faction A", "Box Party Faction B", "Circle Party Coalition", "Circle Party")),
    y = c("no","yes")
  ),
  weights = "wt",
  design = list(ids =~1)
)
box_prefs$print()
box_prefs$n_questions()

# Some fake population data for demonstration
head(approx_voters_popn)

# Create SurveyData object for the population
popn_obj <- SurveyData$new(
  data = approx_voters_popn,
  questions = list(
    age_group = "Which age group are you?",
    gender = "Gender?",
    vote_pref = "Which party do you prefer to vote for?"
  ),
  # order doesn't matter (gender before age here) because
  # the list has the names of the variables
  responses = list(
    gender = levels(approx_voters_popn$gender),
    age_group = levels(approx_voters_popn$age_group),
    vote_pref = levels(approx_voters_popn$vote_pref)
  ),
  weights = "wt"
)
popn_obj$print()

# Create the QuestionMap objects mapping each question between the
# survey and population dataset
q_age <- QuestionMap$new(
  name = "age",
  col_names = c("age","age_group"),
  values_map = list(
    "18-25" = "18-35", "26-35" = "18-35","36-45" = "36-55",
    "46-55" = "36-55", "56-65" = "56-65", "66-75" = "66+", "76-90" = "66+"
  )
)
print(q_age)

q_party_pref <- QuestionMap$new(
  name = "party_pref",
  col_names = c("vote_for","vote_pref"),
  values_map = list("Box Party" = "BP",  "BP" = "BP","Circle Party" = "CP", "CP" = "CP")
)
q_gender <- QuestionMap$new(
  name = "gender",
  col_names = c("gender", "gender"),
  values_map = list("male" = "m","female" = "f", "nonbinary" = "nb")
)

# Create SurveyMap object adding all questions at once
ex_map <- SurveyMap$new(
  sample = box_prefs,
  population = popn_obj,
  q_age,
  q_party_pref,
  q_gender
)
print(ex_map) # or ex_map$print()

# Or can add questions incrementally
ex_map <- SurveyMap$new(sample = box_prefs, population = popn_obj)
print(ex_map)

ex_map$add(q_age, q_party_pref)
print(ex_map)

ex_map$add(q_gender)
print(ex_map)

# Create the mapping between sample and population
ex_map$mapping()

# Create the poststratification data frame using all variables in the mapping
# (alternatively, can specify particular variables, e.g. tabulate("age"))
ex_map$tabulate()

# Take a peak at the poststrat data frame
head(ex_map$poststrat_data())

# Fit regression model using rstanarm (returns a SurveyFit object)
fit_1 <- ex_map$fit(
  fun = rstanarm::stan_glmer,
  formula = y ~ (1|age) + (1|gender),
  family = "binomial",
  seed = 1111,
  chains = 1, # just to keep the example fast and small
  refresh = 0 # suppress printed sampling iteration updates
)

# To use lme4 or brms instead of rstanarm you would use: 
# Example lme4 usage
# fit_2 <- ex_map$fit(
#   fun = lme4::glmer,
#   formula = y ~ (1|age) + (1|gender),
#   family = "binomial"
# )

# Example brms usage
# fit_3 <- ex_map$fit(
#   fun = brms::brm,
#   formula = y ~ (1|age) + (1|gender),
#   family = "bernoulli",
#   seed = 1111
# )

# Predicted probabilities
# returns matrix with rows for poststrat cells, cols for posterior draws
poststrat_estimates <- fit_1$population_predict()

# Compute and summarize estimates by age level and party preference
estimates_by_age <- fit_1$aggregate(poststrat_estimates, by = "age")
estimates_by_party <- fit_1$aggregate(poststrat_estimates, by = "party_pref")

fit_1$summary(estimates_by_age)
fit_1$summary(estimates_by_party)

# Plot estimates
fit_1$plot(estimates_by_party)

fit_1$plot(estimates_by_age)

fit_1$plot(estimates_by_age, additional_stats = "none")
fit_1$plot(estimates_by_age, additional_stats = "wtd")
fit_1$plot(estimates_by_age, additional_stats = "raw")
fit_1$plot(estimates_by_age, additional_stats = c("wtd","raw","mrp"))

# Compute and summarize the population estimate
estimates_popn <- fit_1$aggregate(poststrat_estimates)
fit_1$summary(estimates_popn)

# Plot population estimate
fit_1$plot(estimates_popn)
fit_1$plot(estimates_popn, additional_stats = "none")
fit_1$plot(estimates_popn, additional_stats = "wtd")
fit_1$plot(estimates_popn, additional_stats = "raw")
fit_1$plot(estimates_popn, additional_stats = c("wtd","raw","mrp"))
```

### 2023-06-29 R/SurveyData.R
fix r cmd check note about line length in doc

@@ -25,7 +25,8 @@
 #'   responses = list(
 #'     age = levels(shape_survey$age),
 #'     gender = levels(shape_survey$gender),
-#'     # Here we use a dataframe for the responses because the levels in the data are abridged versions of the actual responses
+#'     # Here we use a dataframe for the responses because the levels
+#'     # in the data are abridged versions of the actual responses.
 #'     # This can be useful when surveys have brief/non descriptive responses.
 #'     vote_for = data.frame(data = levels(shape_survey$vote_for),
 #'     asked = c("Box Party Faction A", "Box Party Faction B", "Circle Party Coalition", "Circle Party")),

### 2023-06-29 R/SurveyData.R
more R cmd check line length fixes

@@ -29,7 +29,8 @@
 #'     # in the data are abridged versions of the actual responses.
 #'     # This can be useful when surveys have brief/non descriptive responses.
 #'     vote_for = data.frame(data = levels(shape_survey$vote_for),
-#'     asked = c("Box Party Faction A", "Box Party Faction B", "Circle Party Coalition", "Circle Party")),
+#'     asked = c("Box Party Faction A", "Box Party Faction B",
+#'               "Circle Party Coalition", "Circle Party")),
 #'     y = c("no","yes")
 #'   ),
 #'   weights = "wt",

### 2023-06-29 R/SurveyMap.R

@@ -29,10 +29,12 @@
 #'   responses = list(
 #'     age = levels(shape_survey$age),
 #'     gender = levels(shape_survey$gender),
-#'     # Here we use a dataframe for the responses because the levels in the data are abridged versions of the actual responses
+#'     # Here we use a dataframe for the responses because the levels
+#'     # in the data are abridged versions of the actual responses.
 #'     # This can be useful when surveys have brief/non descriptive responses.
 #'     vote_for = data.frame(data = levels(shape_survey$vote_for),
-#'     asked = c("Box Party Faction A", "Box Party Faction B", "Circle Party Coalition", "Circle Party")),
+#'     asked = c("Box Party Faction A", "Box Party Faction B",
+#'               "Circle Party Coalition", "Circle Party")),
 #'     y = c("no","yes")
 #'   ),
 #'   weights = "wt",

### 2023-06-29 R/SurveyData.R
Don't duplicate examples in SurveyData doc

@@ -11,53 +11,6 @@
 #' can be specified, with the survey design specified using \pkg{survey} package
 #' notation.
 #'
-#' @examples
-#'
-#' head(shape_survey)
-#' box_prefs <- SurveyData$new(
-#'   data = shape_survey,
-#'   questions = list(
-#'     age = "Please identify your age group",
-#'     gender = "Please select your gender",
-#'     vote_for = "Which party did you vote for in the 2018 election?",
-#'     y = "If today is the election day, would you vote for the Box Party?"
-#'   ),
-#'   responses = list(
-#'     age = levels(shape_survey$age),
-#'     gender = levels(shape_survey$gender),
-#'     # Here we use a dataframe for the responses because the levels
-#'     # in the data are abridged versions of the actual responses.
-#'     # This can be useful when surveys have brief/non descriptive responses.
-#'     vote_for = data.frame(data = levels(shape_survey$vote_for),
-#'     asked = c("Box Party Faction A", "Box Party Faction B",
-#'               "Circle Party Coalition", "Circle Party")),
-#'     y = c("no","yes")
-#'   ),
-#'   weights = "wt",
-#'   design = list(ids =~1)
-#' )
-#' box_prefs$print()
-#' box_prefs$n_questions()
-#'
-#' head(approx_voters_popn)
-#' popn_obj <- SurveyData$new(
-#'   data = approx_voters_popn,
-#'   questions = list(
-#'     age_group = "Which age group are you?",
-#'     gender = "Gender?",
-#'     vote_pref = "Which party do you prefer to vote for?"
-#'   ),
-#'   # order doesn't matter (gender before age here) because
-#'   # the list has the names of the variables
-#'   responses = list(
-#'     gender = levels(approx_voters_popn$gender),
-#'     age_group = levels(approx_voters_popn$age_group),
-#'     vote_pref = levels(approx_voters_popn$vote_pref)
-#'   ),
-#'   weights = "wt"
-#' )
-#' popn_obj$print()
-#'
 SurveyData <- R6::R6Class(
   classname = "SurveyData",
   private = list(
@@ -68,15 +21,15 @@ SurveyData <- R6::R6Class(
     design_ = list()
   ),
   public = list(
-    #' @description Create a new SurveyData object using an existing data frame and other
-    #' survey information.
-    #' This method is used to create the objects for both the sample and the population data.
-    #' If a population is approximated from a large survey (like the ACS or DHS),
[…]

### 2023-06-29 R/QuestionMap.R
more doc cleanup (including doc for more return types)

@@ -3,33 +3,8 @@
 #' @name QuestionMap
 #' @export
 #'
-#' @description An [R6][R6::R6Class] `QuestionMap` object holds the mapping
-#'   for one question or demographic between the survey and population dataset.
-#'
-#' @examples
-#' q_age <- QuestionMap$new(
-#'   name = "age",
-#'   col_names = c("age","age_group"),
-#'   values_map = list(
-#'     "18-25" = "18-35", "26-35" = "18-35","36-45" = "36-55",
-#'     "46-55" = "36-55", "56-65" = "56-65", "66-75" = "66+", "76-90" = "66+"
-#'   )
-#' )
-#' print(q_age)
-#'
-#' q_party_pref <- QuestionMap$new(
-#'   name = "party_pref",
-#'   col_names = c("vote_for","vote_pref"),
-#'   values_map = list("Box Party" = "The BP",
-#'                     "The BP" = "The BP",
-#'                     "The Circle Party" = "The Circles",
-#'                     "The Circles" = "The Circles")
-#' )
-#' q_gender <- QuestionMap$new(
-#'   name = "gender",
-#'   col_names = c("gender", "gender"),
-#'   values_map = list("male" = "m","female" = "f", "nonbinary" = "nb")
-#' )
+#' @description An [R6][R6::R6Class] `QuestionMap` object holds the mapping for
+#'   one question or demographic between the sample data and population data.
 #'
 QuestionMap <- R6::R6Class(
   classname = "QuestionMap",
@@ -41,21 +16,62 @@ QuestionMap <- R6::R6Class(
   public = list(
     #' @description Create a new `QuestionMap`
     #' @param name The name of the underlying construct. For example if the
-    #'   sample data uses the name `age1` and the population data uses
-    #'   the name `age2` then `name` could be `age`.
+    #'   sample data uses the name `"age_bracket"` and the population data uses
+    #'   the name `"age_group"` then `name` could be `"age"`.
     #' @param col_names A character vector of length two giving the column names
     #'   of the question in the sample data (first element) and the population
     #'   data (second element).
     #' @param values_map A named list where the names correspond to the
     #'   responses to the question in the sample data and the values correspond
-    #'   to the responses in the population data. If there is a meaningful
-    #'   ordering over the values, they should be listed in that order, either
-    #'   descending or ascending. Note that if responses were given with both the
-    #'   values in the data and the actual responses given (useful as these will
[…]

### 2023-06-29 R/SurveyData.R

@@ -33,16 +33,20 @@ SurveyData <- R6::R6Class(
     #' @param data A data frame containing the survey data.
     #' @param questions,responses Named lists containing the text of the survey
     #'   questions and the allowed responses, respectively. The names must
-    #'   correspond to the names of variables in `data`. See **Examples**. If
-    #'   these aren't provided then they will be created internally using all
-    #'   factor, character, and binary variables in `data`. Responses can also be
-    #'   provided as a dataframe with column names "data" and "asked", which
-    #'   reflect the responses as coded in the data ("data") and their
-    #'   corresponding actual survey responses ("asked").
+    #'   correspond to the names of variables in `data`. If these aren't
+    #'   provided then they will be created internally using all factor,
+    #'   character, and binary variables in `data`. Responses can also be
+    #'   provided as a data frame with column names `"data"` and `"asked"`,
+    #'   which reflect the responses as coded in the data (`"data"`) and their
+    #'   corresponding actual survey responses (`"asked"`). See **Examples**.
     #' @param weights Optionally, the name of a variable in `data` containing
     #'   survey weights.
     #' @param design Optionally, a named list of arguments (except `weights` and
     #'   `data`) to pass to `survey::svydesign()` to specify the survey design.
+    #'
+    #' @return A `SurveyData` object that can be used in the creation of a
+    #'   [`SurveyMap`] object.
+    #'
     #' @examples
     #' # Example sample data
     #' head(shape_survey)
@@ -59,12 +63,14 @@ SurveyData <- R6::R6Class(
     #'   responses = list(
     #'     age = levels(shape_survey$age),
     #'     gender = levels(shape_survey$gender),
-    #'     # Here we use a dataframe for the responses because the levels
+    #'     # Here we use a data frame for the responses because the levels
     #'     # in the data are abridged versions of the actual responses.
     #'     # This can be useful when surveys have brief/non descriptive responses.
-    #'     vote_for = data.frame(data = levels(shape_survey$vote_for),
-    #'     asked = c("Box Party Faction A", "Box Party Faction B",
-    #'               "Circle Party Coalition", "Circle Party")),
+    #'     vote_for = data.frame(
+    #'       data = levels(shape_survey$vote_for),
+    #'       asked = c("Box Party Faction A", "Box Party Faction B",
+    #'                 "Circle Party Coalition", "Circle Party")
+    #'     ),
     #'     y = c("no","yes")
     #'   ),
     #'   weights = "wt",
[…]

### 2023-06-29 R/SurveyMap.R

@@ -29,13 +29,15 @@
 #'   responses = list(
 #'     age = levels(shape_survey$age),
 #'     gender = levels(shape_survey$gender),
-#'     # Here we use a dataframe for the responses because the levels
+#'     # Here we use a data frame for the responses because the levels
 #'     # in the data are abridged versions of the actual responses.
 #'     # This can be useful when surveys have brief/non descriptive responses.
-#'     vote_for = data.frame(data = levels(shape_survey$vote_for),
-#'     asked = c("Box Party Faction A", "Box Party Faction B",
-#'               "Circle Party Coalition", "Circle Party")),
-#'     y = c("no","yes")
+#'     vote_for = data.frame(
+#'       data = levels(shape_survey$vote_for),
+#'       asked = c("Box Party Faction A", "Box Party Faction B",
+#'                 "Circle Party Coalition", "Circle Party")
+#'     ),
+#'     y = c("no", "yes")
 #'   ),
 #'   weights = "wt",
 #'   design = list(ids =~1)
@@ -199,6 +201,7 @@ SurveyMap <- R6::R6Class(
     #' @param sample The [SurveyData] object corresponding to the sample data.
     #' @param population The [SurveyData] object corresponding to the population data.
     #' @param ... [QuestionMap] objects.
+    #' @return A `SurveyMap` object.
     initialize = function(sample, population, ...) {
       if (!inherits(sample, "SurveyData")) {
         stop("'sample' must be a SurveyData object.", call. = FALSE)
@@ -222,6 +225,7 @@ SurveyMap <- R6::R6Class(
 
     #' @description Print a summary of the mapping.
     #' @param ... Currently ignored.
+    #' @return The `SurveyMap` object, invisibly.
     print = function(...) {
       if (length(private$item_map_) > 0) {
         for (i in 1:length(private$item_map_)) {
@@ -243,6 +247,7 @@ SurveyMap <- R6::R6Class(
 
     #' @description Add new [QuestionMap]s.
     #' @param ... The [QuestionMap]s to add.
+    #' @return The `SurveyMap` object, invisibly.
     add = function(...) {
       dots <- list(...)
       for (i in 1:length(dots)) {
@@ -260,6 +265,7 @@ SurveyMap <- R6::R6Class(
 
     #' @description Delete [QuestionMap]s.
     #' @param ... The [QuestionMap]s to delete.
+    #' @return The `SurveyMap` object, invisibly.
     delete = function(...) {
[…]

### 2023-06-29 R/SurveyFit.R
document return types of SurveyFit methods

@@ -18,11 +18,12 @@ SurveyFit <- R6::R6Class(
   public = list(
 
     #' @description Create a new `SurveyFit` object. This method is called
-    #'   internally by the `$fit()` method of the [`SurveyMap`] object and does
+    #'   internally by the `fit` method of the [`SurveyMap`] object and does
     #'   not need to be called directly by the user.
     #' @param fit A fitted model object.
     #' @param map A [`SurveyMap`] object.
-    #' @param formula A formula object representing the model ran.
+    #' @param formula A formula object for the model that was fit.
+    #' @return A `SurveyFit` object.
     initialize = function(fit, map, formula) {
       private$fit_ <- fit
       private$map_ <- map
@@ -30,29 +31,38 @@ SurveyFit <- R6::R6Class(
       invisible(self)
     },
 
-    #' @description Access the fitted model object
+    #' @description Access the fitted model object.
+    #' @return The fitted model object created by the modeling function called
+    #'   by the `fit` method of the [`SurveyMap`] object. For example, if using
+    #'   `rstanarm::stan_glmer()` then a `stanreg` object from \pkg{rstanarm} is
+    #'   returned.
     fit = function() {
       private$fit_
     },
 
-    #' @description Access the SurveyMap object
+    #' @description Access the [`SurveyMap`] object.
+    #' @return The [`SurveyMap`] associated with the `SurveyFit` object.
     map = function() {
       private$map_
     },
 
-    #' @description Access the SurveyMap object
+    #' @description Access the model formula.
+    #' @return The model formula used when fitting the model.
     formula = function() {
       private$formula_
     },
 
-    #' @description Call the fitted model object's print method
+    #' @description Call the fitted model object's print method. The console
+    #'   output from this method depends on the model fitting function used.
     #' @param ... Optional arguments to pass the print method.
+    #' @return The `SurveyFit` object, invisibly.
     print = function(...) {
       print(private$fit_, ...)
       invisible(self)
     },
 
-    #' @description Use fitted model to add predicted probabilities to post-stratification dataset.
+    #' @description Use fitted model to add predicted probabilities to
+    #'   post-stratification dataset.
     #' @param fun The function to use to generate the predicted probabilities.
     #'   This should only be specified if you used a model fitting function
     #'   not natively supported by \pkg{mrpkit}.
@@ -120,7 +130,8 @@ SurveyFit <- R6::R6Class(
       }
     },
 
-    #' @description Aggregate estimates to the population level or by level of a grouping variable
[…]

### 2023-06-29 R/SurveyMap.R

@@ -592,11 +592,11 @@ SurveyMap <- R6::R6Class(
     #' @return A named list of [`QuestionMap`]s.
     item_map = function() private$item_map_,
 
-    #' @description Access the [SurveyData] object containing the sample data.
+    #' @description Access the [`SurveyData`] object containing the sample data.
     #' @return A [`SurveyData`] object.
     sample = function() private$sample_,
 
-    #' @description Access the [SurveyData] object containing the population data.
+    #' @description Access the [`SurveyData`] object containing the population data.
     #' @return A [`SurveyData`] object.
     population = function() private$population_,
 

### 2023-06-29 R/SurveyFit.R
fixes to allow custom functions

@@ -116,7 +116,7 @@ SurveyFit <- R6::R6Class(
         }
       } else {
         fun <- match.fun(fun)
-        fun(fitted_model, poststrat, ...)
+        fun(private$fit_, poststrat, ...)
       }
     },
 

### 2023-06-29 R/SurveyMap.R

@@ -544,7 +544,8 @@ SurveyMap <- R6::R6Class(
                 call. = FALSE
         )
       }
-      if (!any(getNamespaceName(environment(fun)) %in% c("lme4","brms","rstanarm"))) {
+      if (isNamespace(environment(fun))
+          && !any(getNamespaceName(environment(fun)) %in% c("lme4","brms","rstanarm"))) {
         warning("Only rstanarm, brms and lme4 are supported natively. ",
                 "Other modeling tools will need a custom population_predict() method.",
                 call. = FALSE)

### 2023-06-29 tests/testthat/test-SurveyFit.R

@@ -470,3 +470,26 @@ test_that("force factor works appropriately",{
   x <- c(1,1, NA)
   expect_equal(force_factor(x), c(1,1, NA))
 })
+
+
+test_that("custom prediction function works", {
+  myfit <- function(data, formula, ...) {
+    rstanarm::stan_glm(formula = formula, data = data, ...)
+  }
+  mypred <- function(mod, ps, ...) {
+    t(suppressMessages(rstanarm::posterior_linpred(
+      mod,
+      newdata = ps,
+      transform = TRUE
+    )))
+  }
+  fit1 <- ex_map$fit(
+    fun = myfit,
+    formula = y ~ age + gender,
+    family = "binomial",
+    iter = 1000,
+    refresh = 0
+  )
+  p <- fit1$population_predict(fun = mypred)
+  expect_equal(dim(p), c(nrow(ex_map$poststrat_data()), 2000))
+})

### 2023-06-29 R/SurveyFit.R
avoid one of the warnings on r-devel

@@ -181,7 +181,7 @@ SurveyFit <- R6::R6Class(
           dplyr::mutate(method = "mrp")
 
           lhs_binary <- self$map()$sample()$survey_data() %>%
-          dplyr::select(lhs_var, .key)%>%
+          dplyr::select(dplyr::all_of(lhs_var), .key)%>%
           dplyr::mutate(lhs_binary = force_factor(.data[[lhs_var]]))
 
         raw_data <- self$map()$mapped_sample_data() %>%

### 2023-06-30 R/QuestionMap.R
more details in doc for model fitting and mapping

@@ -17,7 +17,8 @@ QuestionMap <- R6::R6Class(
     #' @description Create a new `QuestionMap`
     #' @param name The name of the underlying construct. For example if the
     #'   sample data uses the name `"age_bracket"` and the population data uses
-    #'   the name `"age_group"` then `name` could be `"age"`.
+    #'   the name `"age_group"` then `name` could be `"age"`. This name will be
+    #'   used later in the workflow when creating a data set with unified names.
     #' @param col_names A character vector of length two giving the column names
     #'   of the question in the sample data (first element) and the population
     #'   data (second element).

### 2023-06-30 R/SurveyMap.R

@@ -381,18 +381,24 @@ SurveyMap <- R6::R6Class(
       invisible(self)
     },
 
-    #' @description The mapping method uses the given maps between questions to
-    #'   create new sample and population data that has unified variable names
-    #'   (i.e., if the underlying construct is called `age`, both sample and
-    #'   population will have an `age` column, even if in the the raw data both
-    #'   had different variable names). The method also unifies the levels of
-    #'   each variable in the sample and population so that the maximum set of
-    #'   consistent levels is created. Names of these new levels are made
-    #'   according the the sample level names. If multiple levels are combined,
-    #'   the new name will be the existing levels separated by a ` + ` (e.g.
-    #'   if age groups "18-25" and "26-30" are combined the new name will be
-    #'   "18-25 + 26-30").
+    #' @description The `mapping` method uses the given maps between questions
+    #'   to create new sample and population data frames that have unified
+    #'   variable names (e.g., if the underlying construct is called `age`, both
+    #'   sample and population will have an `age` column, even if in the the raw
+    #'   data both had different variable names).
+    #'
+    #'   This method also unifies the levels of each variable in the sample and
+    #'   population so that the maximum set of consistent levels is created.
+    #'   Names of these new levels are made according the the sample level
+    #'   names. If multiple levels are combined, the new name will be the
+    #'   existing levels separated by a ` + ` (e.g. if age groups `"18-25"` and
+    #'   `"26-30"` are combined the new name will be `"18-25 + 26-30"`).
+    #'
+    #'   Use the `mapped_sample_data` and `mapped_population_data` methods to
+    #'   access the resulting data frames.
+    #'
     #' @return The `SurveyMap` object, invisibly.
+    #'
     mapping  = function() {
       for (j in 1:length(private$item_map_)) {
         samp_mapnames <- private$item_map_[[j]]$col_names()[1]
@@ -491,7 +497,7 @@ SurveyMap <- R6::R6Class(
     },
 
     #' @description Prepare the poststratification table. The resulting data
-    #'   frame is available via the `SurveyMap$poststrat_data()` method. See
+    #'   frame is available via the `poststrat_data` method. See
     #'   **Examples**.
     #' @param ... The names of the variables to include as strings.
     #' @return The `SurveyMap` object, invisibly.
@@ -514,15 +520,21 @@ SurveyMap <- R6::R6Class(
       invisible(self)
     },
 
[…]

### 2023-06-30 R/SurveyFit.R
clarify plot method doc

@@ -227,21 +227,29 @@ SurveyFit <- R6::R6Class(
       as.data.frame(out)
     },
 
-    #' @description Plot takes the aggregated MRP estimates and produces a quick
-    #' visualization of total and sub-population estimates.
+    #' @description Visualize population or sub-population estimates.
+    #'
+    #'   When passed the data frame containing the posterior distribution of the
+    #'   population MRP estimate a density plot is generated.  If visualizing
+    #'   sub-populations it generates a violin plot of the posterior
+    #'   distribution of the aggregated MRP estimates for each level of the
+    #'   grouping variable. The `additional_stats` argument controls which
+    #'   other information is overlaid on the plot.
+    #'
     #' @param aggregated_estimates The data frame returned by the `aggregate` method.
     #' @param additional_stats A vector that specifies which of three additional
-    #'   stats ("wtd", "raw", "mrp", "none") should be overlayed on the plot.
-    #'   The default is to include the weighted estimate (wtd) and raw data mean
-    #'   (raw), but an analogous bar for MRP (mrp) can be added using the
-    #'   posterior mean and sd. The sd for the weighted estimate uses the survey
-    #'   design and the \pkg{survey} package, whilst the raw estimate is a
-    #'   direct mean and binomial sd of the binary responses. Uncertainty is
+    #'   stats (`"wtd"`, `"raw"`, `"mrp"`, `"none"`) should be overlaid on the
+    #'   plot. The default is to overlay intervals for the weighted and raw
+    #'   estimates on top of the density plot representing the MRP estimates.
+    #'   The weighted estimates are computed by passing the optional survey
+    #'   weights and design specified in the [`SurveyData`] to the \pkg{survey}
+    #'   package. The raw estimate is a direct mean and binomial sd of the
+    #'   binary responses. Uncertainty estimates for the `additional_stats` are
     #'   included on violin plots but not on density plots. Intervals are 95%
     #'   CI.
-    #' @return A ggplot object that is either a violin plot (if aggregated
-    #'   estimates is at a small area level) or density plot (if aggregated
-    #'   estimates is at the population level).
+    #' @return A ggplot object that is either a violin plot if showing small
+    #'   area level (sub-population) estimates, or a density plot if showing
+    #'   population estimates.
     plot = function(aggregated_estimates, additional_stats = c("wtd","raw")) {
[…]

### 2023-07-01 README.Rmd
a few more edits to readme

@@
 For students and researchers who are comfortable with at least `glm()` and want
-to conduct multilevel regression with post-stratification (MRP), `mrpkit`
-provides a reproducible, opinionated, and highly structured workflow. Unlike
-writing all the code yourself, using `mrpkit` proactively addresses many common
-issues, and makes it possible for folks who are new to MRP to quickly conduct
-their first analysis.
+to conduct multilevel regression with post-stratification (MRP), the `mrpkit` R
+package provides a reproducible, opinionated, and highly structured workflow.
+Unlike writing all the code yourself, using `mrpkit` proactively addresses many
+common issues, and makes it possible for people who are new to MRP to quickly
+conduct their first analysis.
 
-The package assists in setting up the survey data and relationships between
-different variables in the sample and the population. From there, a substantial
-amount of data cleaning is automated, saving time and reducing the risk of
-coding errors. The package has native support for multilevel binomial and
+The package first assists in setting up the survey data and relationships
+between different variables in the sample and the population. From there, a
+substantial amount of data cleaning is automated, saving time and reducing the
+risk of coding errors. `mrpkit` has native support for multilevel binomial and
 Bernoulli models fit with lme4 and Stan (via brms and rstanarm) and also allows
-for the use of custom modeling functions. Summaries and simple visualizations of
-the resulting post-stratified estimates are provided.
+for the use of custom modeling functions. After model fitting, `mrpkit` handles
+the post-stratification step, producing population and sub-population estimates.
+Summary statistics and simple visualizations of the resulting MRP estimates are
+provided.
 
